---
title: "interrow weed biomass"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r message=FALSE, warning=FALSE}
#Set work directory
setwd("/Users/ey239/Github/Mowtivation/rmarkdowns")

#Load packages 
library(tidyverse) ##install.packages("tidyverse")
library(knitr)
library(patchwork) ##install.packages("patchwork")
library(skimr)     ##install.packages("skimr")
library(readxl)
library(janitor) ##install.packages("janitor")
library(MASS)
library(kableExtra) ##install.packages("kableExtra")
library(webshot) ##install.packages("webshot")
webshot::install_phantomjs()
library(viridis) ##install.packages("viridis")
library(lme4) ##install.packages("lme4")
library(lmerTest) ##install.packages("lmerTest")
library(emmeans) ##install.packages("emmeans")
library(rstatix) ##install.packages("rstatix")
#library(Matrix) ##install.packages("Matrix")
library(multcomp) ##install.packages("multcomp")
library(multcompView) ##install.packages("multcompView")
library(ggResidpanel) ##install.packages("ggResidpanel")
#library(car)
#library(TMB)  ##install.packages("TMB")
library(glmmTMB)  ##install.packages("glmmTMB")
library(DHARMa)  ##install.packages("DHARMa")
library(performance) ##install.packages("performance")

#Load Functions
MeanPlusSe<-function(x) mean(x)+plotrix::std.error(x)

find_logw0=function(x){c=trunc(log(min(x[x>0],na.rm=T)))
d=exp(c)
return(d)}
```
<br>

# Load and Clean Data

## Load data

``` {r}
combined_raw <- read_excel("~/Github/Mowtivation/raw-data/All Treatments/combined_raw.xlsx")
kable(head(combined_raw))

```

## Clean data
```{r}


#Standardaze column names, convert to factors, check for outliers of variable**
clean_combined <- clean_names(combined_raw) |>  
  rename ('weed_control'= treatment) |> 
  mutate(across(c(weed_control, block, plot, location, year), as.factor)) #|> 
  #mutate(is_outlier = totwbm < (quantile(totwbm, 0.25) - 1.5 * IQR(totwbm)) |
                       #wbm > (quantile(totwbm, 0.75) + 1.5 * IQR(totwbm)))

#select and convert data for wbm analysis
interrow_weed_biomass_clean <- clean_combined |>
  mutate(
    interrow_weed_biomass_grams_meter = interrow_weed_biomass / 0.5,
    interrow_weed_biomass_kg_ha = interrow_weed_biomass_grams_meter * 10,
    interrow_weed_biomass_lbs_ac = interrow_weed_biomass_kg_ha * 0.892179,
  )

kable(head(interrow_weed_biomass_clean))


```

Tyler is under the impression that block should always be random and that post-hoc comparisons should use TUKEY rather the Fischer. Fisher is bogus apparently. 

#Model testing Location fixed
##Glmm with Tweedie Regression 
###Metric
```{r}

model_tweedie <- glmmTMB(
interrow_weed_biomass_kg_ha ~  weed_control*location +
  (1|location:block), 
  data = interrow_weed_biomass_clean, 
  family = tweedie(link = "log")

)


### Two checks specifically for a generalize linear approach
simulateResiduals(model_tweedie,plot = TRUE) # Residuals and normality look good
check_model(model_tweedie) 
summary(model_tweedie )
VarCorr(model_tweedie)


```
### Joint test (anova)
```{r}
model_tweedie |> 
  joint_tests() |> 
  kable()  
```
#### Anova

```{r}
interrow_weed_biomass_model_tweedie.anova <- Anova(model_tweedie, type = 3)
print(interrow_weed_biomass_model_tweedie.anova)

```
<br>

## Tukey means comparisons with fisher cld
### Weed-control (significant)

```{r}
#model predicted means
tukey_weed_control <- emmeans(model_tweedie, pairwise ~ weed_control, adjust = "tukey", type = "response")
tukey_weed_control

#raw means
raw_weed_control <- aggregate(interrow_weed_biomass_kg_ha ~ weed_control, data = interrow_weed_biomass_clean, mean)
print(raw_weed_control)


cld_weed_control_fisher <-cld(emmeans(model_tweedie, ~  weed_control, type = "response"), Letters = letters,adjust = "none", sort = TRUE, reversed=TRUE)
cld_weed_control_fisher 
```

### Location (significant)
```{r}
#model predicted means
tukey_location <- emmeans(model_tweedie, pairwise ~ location, adjust = "tukey", type = "response")
tukey_location

#raw means
raw_location <- aggregate(interrow_weed_biomass_kg_ha ~ location, data = interrow_weed_biomass_clean, mean)
print(raw_location)


cld_location_fisher <-cld(emmeans(model_tweedie, ~  location, type = "response"), Letters = letters,adjust = "none", sort = TRUE, reversed=TRUE)
cld_location_fisher
```
### Fisher compact letter display
#### Weed_control|location (significant)

```{r}
#model predicted means
tukey_weed_control_location <- emmeans(model_tweedie, pairwise ~ weed_control|location, adjust = "tukey", type = "response")
tukey_weed_control_location

#raw means
raw_weed_control_location <- aggregate(interrow_weed_biomass_kg_ha ~ weed_control+location, data = interrow_weed_biomass_clean, mean)
print(raw_weed_control_location)

# Perform pairwise comparisons and generate compact letter displays
cld_weed_control_location_fisher <- cld(
  emmeans(model_tweedie, ~ weed_control | location, type = "response"),
  Letters = letters,
  adjust = "none",
  sort = TRUE,
  reversed = TRUE
)

# Display the results
print(cld_weed_control_location_fisher)


```

# Figures
##lbs/a
```{r message=FALSE}
interrow_weed_biomass_clean |> 
  left_join(cld_weed_control_fisher_zero) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")), y = response, fill = weed_control)) +
  facet_wrap( ~location, labeller = labeller(
    location = c("field O2 east" = "Field O2 East", "field O2 west" = "Field O2 West","field v" = "Field V" )))+
  #stat_summary(geom = "bar", fun = mean, width = 0.7) +  # Ensuring mean calculation
  #stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2) +  # Explicit mean_se
  #stat_summary(geom = "text", fun = mean, aes(label = trimws(.group)), 
               #size = 6.5, vjust = -0.5) +  # Adjusting labels) 
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), width=.2,
                 position=position_dodge(.9))+
geom_text(aes(label = trimws(.group), y = response + (SE + 20)), size = 7) +
  labs(
    x = "",
    y = expression("Weed biomass" ~ (kg~ha^{-1})),
    #title = str_c("Influence of interrow weed control on interrow weed biomass"),
    subtitle = expression(italic("P < 0.005"))) +
  
 scale_x_discrete(labels = c("Rolled,\nno additional\nweed control",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nhigh-residue\ncultivation",
                              "Tilled,\ninterrow\nmowing",
                          "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 20),
    axis.title = element_text(size = 20),  # Increase font size of axis titles
    axis.text = element_text(size = 20),   # Increase font size of axis labels
    plot.title = element_text(size = 24, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 24, face = "italic")  # Increase font size of subtitle
  )
ggsave("interrow_weed_biomass_weed_control_location_kg_ha_mr.png", width = 24, height = 8, dpi = 300)
```

###Eric conversion
```{r eval=FALSE, message=FALSE, include=FALSE}
interrow_weed_biomass_clean |> 
  left_join(cld_weed_control_fisher) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")), y = response, fill = weed_control)) +
  #stat_summary(geom = "bar", fun = mean, width = 0.7) +  # Ensuring mean calculation
  #stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2) +  # Explicit mean_se
  #stat_summary(geom = "text", fun = mean, aes(label = trimws(.group)), 
               #size = 6.5, vjust = -0.5) +  # Adjusting labels) 
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), width=.2,
                 position=position_dodge(.9))+
geom_text(aes(label = trimws(.group), y = response + (SE + 20)), size = 7) +
  labs(
    x = "",
    y = expression("Interrow weed biomass" ~ (lbs * "/" * a)),
    #title = str_c("Influence of interrow weed control on interrow weed biomass"),
    subtitle = expression(italic("P < 0.005"))) +
  
 scale_x_discrete(labels = c("Rolled,\nno additional\nweed control",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nhigh-residue\ncultivation",
                              "Tilled,\ninterrow\nmowing",
                          "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12),
    axis.title = element_text(size = 24),  # Increase font size of axis titles
    axis.text = element_text(size = 24),   # Increase font size of axis labels
    plot.title = element_text(size = 24, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 24, face = "italic")  # Increase font size of subtitle
  )
ggsave("interrow_weed_biomass_weed_control_lbac_ey.png", width = 10, height = 8, dpi = 300)
```
##Metric
###Matt conversion
```{r message=FALSE}
interrow_weed_biomass_clean |> 
  left_join(cld_weed_control_location_fisher) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")), y = response, fill = weed_control)) +
  facet_wrap( ~location, labeller = labeller(
    location = c("field O2 east" = "Field O2 East", "field O2 west" = "Field O2 West","field v" = "Field V" )))+
  #stat_summary(geom = "bar", fun = mean, width = 0.7) +  # Ensuring mean calculation
  #stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2) +  # Explicit mean_se
  #stat_summary(geom = "text", fun = mean, aes(label = trimws(.group)), 
               #size = 6.5, vjust = -0.5) +  # Adjusting labels) 
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), width=.2,
                 position=position_dodge(.9))+
geom_text(aes(label = trimws(.group), y = response + (SE + 25)), size = 7) +
  labs(
    x = "",
    y = expression("Weed biomass" ~ (kg~ha^{-1})),
    #title = str_c("Influence of interrow weed control on interrow weed biomass"),
    subtitle = expression(italic("P < 0.001"))) +
  
 scale_x_discrete(labels = c("Rolled,\nno additional\nweed control",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nhigh-residue\ncultivation",
                              "Tilled,\ninterrow\nmowing",
                          "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 20),
    axis.title = element_text(size = 20),  # Increase font size of axis titles
    axis.text = element_text(size = 20),   # Increase font size of axis labels
    plot.title = element_text(size = 24, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 24, face = "italic")  # Increase font size of subtitle
  )
ggsave("interrow_weed_biomass_weed_control_location_kg_ha.png", width = 24, height = 8, dpi = 300)
```

###Eric conversion
```{r eval=FALSE, message=FALSE, include=FALSE}
interrow_weed_biomass_clean |> 
  left_join(cld_weed_control_fisher) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")), y = response, fill = weed_control)) +
  #stat_summary(geom = "bar", fun = mean, width = 0.7) +  # Ensuring mean calculation
  #stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2) +  # Explicit mean_se
  #stat_summary(geom = "text", fun = mean, aes(label = trimws(.group)), 
               #size = 6.5, vjust = -0.5) +  # Adjusting labels) 
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), width=.2,
                 position=position_dodge(.9))+
geom_text(aes(label = trimws(.group), y = response + (SE + 20)), size = 7) +
  labs(
    x = "",
    y = expression("Interrow weed biomass" ~ ~ (kg ~ ha^{-1})),
    #title = str_c("Influence of interrow weed control on interrow weed biomass"),
    subtitle = expression(italic("P < 0.005"))) +
  
 scale_x_discrete(labels = c("Rolled,\nno additional\nweed control",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nhigh-residue\ncultivation",
                              "Tilled,\ninterrow\nmowing",
                          "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12),
    axis.title = element_text(size = 24),  # Increase font size of axis titles
    axis.text = element_text(size = 24),   # Increase font size of axis labels
    plot.title = element_text(size = 24, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 24, face = "italic")  # Increase font size of subtitle
  )
ggsave("interrow_weed_biomass_weed_control_kgha_ey.png", width = 10, height = 8, dpi = 300)
```
