---
title: "Weed biomass"

output:
  github_document:
    toc: true
always_allow_html: true


---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.path = "analysis/figs/weed_biomass-"
)
```

# Packages
```{r}
# Packages
library(tidyverse)    # includes dplyr, ggplot2, readr, tibble, etc.
library(janitor)
library(readxl)
library(glmmTMB)
library(DHARMa)
library(emmeans)
library(multcomp)
library(car)
library(kableExtra)
library(here)
library(conflicted)
library(lme4)
library(WrensBookshelf)
library(writexl)

# Handle conflicts
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::recode)

# Treatment level order (use everywhere)
mow_levels <- c(
  "Rolled, no control",
  "Rolled, mowing",
  "Rolled, high-residue cultivation",
  "Tilled, mowing",
  "Tilled, cultivation"
)

# One consistent CVD-safe color palette for all figures (WrensBookshelf)
fill_cols <- WB_brewer(
  name = "WhatWellBuild",
  n    = length(mow_levels),
  type = "discrete"
) |>
  setNames(mow_levels)

# x-axis label helpers ---------------------------------------------------

# Break on spaces (if you ever want every word on its own line)
label_break_spaces <- function(x) {
  stringr::str_replace_all(x, " ", "\n")
}

# Break after the comma: "Rolled,\nno control", etc.
label_break_comma <- function(x) {
  stringr::str_replace_all(x, ", ", ",\n")
}

# Break after comma AND split "high-residue cultivation"
# -> "Rolled,\nhigh-residue\ncultivation"
label_break_comma_cult <- function(x) {
  x |>
    stringr::str_replace("high-residue cultivation",
                         "high-residue\ncultivation") |>
    stringr::str_replace_all(", ", ",\n")
}

# Helper: tidy emmeans output regardless of CI column names --------------
# (works directly on an emmeans object)

tidy_emm <- function(emm, ref_levels = NULL) {
  emm_df <- as.data.frame(emm)

  lcl_col <- intersect(c("lower.CL", "asymp.LCL"), names(emm_df))[1]
  ucl_col <- intersect(c("upper.CL", "asymp.UCL"), names(emm_df))[1]

  if (is.na(lcl_col) || is.na(ucl_col)) {
    stop("Could not find CI columns in emmeans output.")
  }

  out <- emm_df |>
    dplyr::mutate(
      ci_low  = .data[[lcl_col]],
      ci_high = .data[[ucl_col]]
    )

  if (!is.null(ref_levels) && "weed_trt" %in% names(out)) {
    out <- out |>
      dplyr::mutate(weed_trt = factor(weed_trt, levels = ref_levels))
  }

  out
}

```

# Data import and preparation
``` {r}
# 1) Read + clean master (weed biomass) ---------------------------------

weed_biomass_clean <- read_excel(
  here("data", "raw", "All Treatments", "combined_raw.xlsx")
) |>
  clean_names() |>
  rename(weed_trt = treatment) |>
  mutate(
    year      = factor(year),
    location  = factor(location),
    site_year = factor(interaction(year, location, drop = TRUE)),
    block     = factor(block),
    weed_trt  = recode(
      weed_trt,
      "RNO" = "Rolled, no control",
      "RIM" = "Rolled, mowing",
      "RIC" = "Rolled, high-residue cultivation",
      "TIM" = "Tilled, mowing",
      "TIC" = "Tilled, cultivation"
    ),
    weed_trt = factor(weed_trt, levels = mow_levels)
  ) |>
  # keep only rows with non-missing total weed biomass
  filter(!is.na(weed_biomass)) |>
  # per-area metrics, assuming all biomass vars are g per 0.5 m² quadrat
  mutate(
    weed_biomass_g_m2           = weed_biomass          * 2,   # g/m²
    weed_biomass_kg_ha          = weed_biomass          * 20,  # kg/ha
    inrow_weed_biomass_g_m2     = inrow_weed_biomass    * 2,
    inrow_weed_biomass_kg_ha    = inrow_weed_biomass    * 20,
    interrow_weed_biomass_g_m2  = interrow_weed_biomass * 2,
    interrow_weed_biomass_kg_ha = interrow_weed_biomass * 20
  )

# 2) 2023 Field V subset -----------------------------------------------

weed_biomass_field_v_2023 <- weed_biomass_clean |>
  filter(year == 2023, location == "field v")

# Quick check
kable(
  head(weed_biomass_clean),
  caption = "All site-years, cleaned (weed biomass)"
)

kable(
  head(weed_biomass_field_v_2023),
  caption = "Field V only, 2023 (weed biomass)"
)


```

# Model testing and diagnostics
## Exploratory checks
### Total weed biomass (kg ha⁻¹) by site-year × treatment
```{r}
## Exploratory data visualisation
### Total weed biomass (kg ha^-1) by site-year × treatment

# 1) Summary table: total weed biomass by site-year × treatment ---------
weed_biomass_clean |>
  group_by(site_year, weed_trt) |>
  summarise(
    n      = n(),
    mean   = mean(weed_biomass_kg_ha,   na.rm = TRUE),
    median = median(weed_biomass_kg_ha, na.rm = TRUE),
    sd     = sd(weed_biomass_kg_ha,     na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(site_year, weed_trt) |>
  kable(
    digits  = 1,
    caption = "Weed biomass (kg ha^-1) by site-year × treatment"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

# Nice facet labels for site_year ---------------------------------------
site_year_labs <- c(
  "2024.field O2 east" = "2024 \u2013 Field O2 East",
  "2024.field O2 west" = "2024 \u2013 Field O2 West",
  "2023.field v"       = "2023 \u2013 Field V"
)

# 2) Faceted boxplot: all site-years -----------------------------------
wbm_box_total <- weed_biomass_clean |>
  ggplot(aes(x = weed_trt, y = weed_biomass_kg_ha, fill = weed_trt)) +
  geom_boxplot(
    outlier.shape = NA,
    width  = 0.55,
    color  = "black"
  ) +
  geom_jitter(
    width  = 0.12,
    height = 0,
    alpha  = 0.4,
    size   = 1.8,
    color  = "grey30"
  ) +
  facet_wrap(
    ~ site_year,
    nrow     = 1,
    labeller = labeller(site_year = site_year_labs)
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x     = NULL,
    y     = expression(Weed~biomass~"(kg"~ha^{-1}*")"),
    title = "Weed biomass by treatment across site-years"
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 10),
    strip.text  = element_text(face = "bold")
  )

# Print to document
wbm_box_total



# 3) Save figure --------------------------------------------------------
fig_dir_total <- here("analysis", "figs", "weed-biomass", "total")
dir.create(fig_dir_total, showWarnings = FALSE, recursive = TRUE)

ggsave(
  filename = file.path(fig_dir_total,
                       "fig_weed_biomass_total_box_by-site-year_kg_ha.png"),
  plot     = wbm_box_total,
  width    = 12,
  height   = 5.5,
  dpi      = 300
)


```

### In-row weed biomass (kg ha⁻¹) by site-year × treatment
```{r}
## In-row weed biomass (kg ha^-1) by site-year × treatment --------------

# 1) Summary table: in-row weed biomass by site-year × treatment -------
weed_biomass_clean |>
  group_by(site_year, weed_trt) |>
  summarise(
    n      = n(),
    mean   = mean(inrow_weed_biomass_kg_ha,   na.rm = TRUE),
    median = median(inrow_weed_biomass_kg_ha, na.rm = TRUE),
    sd     = sd(inrow_weed_biomass_kg_ha,     na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(site_year, weed_trt) |>
  kable(
    digits  = 1,
    caption = "In-row weed biomass (kg ha^-1) by site-year × treatment"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

# (Assumes site_year_labs already defined as in the total biomass chunk)
# site_year_labs <- c(
#   "2024.field O2 east" = "2024 \u2013 Field O2 East",
#   "2024.field O2 west" = "2024 \u2013 Field O2 West",
#   "2023.field v"       = "2023 \u2013 Field V"
# )

# 2) Faceted boxplot: all site-years -----------------------------------

inrow_box_total <- weed_biomass_clean |>
  ggplot(aes(x = weed_trt, y = inrow_weed_biomass_kg_ha, fill = weed_trt)) +
  geom_boxplot(
    outlier.shape = NA,
    width  = 0.55,
    color  = "black"
  ) +
  geom_jitter(
    width  = 0.12,
    height = 0,
    alpha  = 0.4,
    size   = 1.8,
    color  = "grey30"
  ) +
  facet_wrap(
    ~ site_year,
    nrow     = 1,
    labeller = labeller(site_year = site_year_labs)
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x     = NULL,
    y     = expression(In-row~weed~biomass~"(kg"~ha^{-1}*")"),
    title = "In-row weed biomass by treatment across site-years"
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 10),
    strip.text  = element_text(face = "bold")
  )

# Print to document
inrow_box_total

# 3) Save figure --------------------------------------------------------
fig_dir_total <- here("analysis", "figs", "weed-biomass", "in-row")
dir.create(fig_dir_total, showWarnings = FALSE, recursive = TRUE)

ggsave(
  filename = file.path(fig_dir_total,
                       "fig_weed_biomass_inrow_box_by-site-year_kg_ha.png"),
  plot     = wbm_box_total,
  width    = 12,
  height   = 5.5,
  dpi      = 300
)



```

### Inter-row weed biomass (kg ha⁻¹) by site-year × treatment
```{r}
## Inter-row weed biomass (kg ha^-1) by site-year × treatment -----------

# 1) Summary table: inter-row weed biomass by site-year × treatment ----
weed_biomass_clean |>
  group_by(site_year, weed_trt) |>
  summarise(
    n      = n(),
    mean   = mean(interrow_weed_biomass_kg_ha,   na.rm = TRUE),
    median = median(interrow_weed_biomass_kg_ha, na.rm = TRUE),
    sd     = sd(interrow_weed_biomass_kg_ha,     na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(site_year, weed_trt) |>
  kable(
    digits  = 1,
    caption = "Inter-row weed biomass (kg ha^-1) by site-year × treatment"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

# Assumes site_year_labs already defined earlier, e.g.:
# site_year_labs <- c(
#   "2024.field O2 east" = "2024 \u2013 Field O2 East",
#   "2024.field O2 west" = "2024 \u2013 Field O2 West",
#   "2023.field v"       = "2023 \u2013 Field V"
# )

# 2) Faceted boxplot: all site-years -----------------------------------

interrow_box_total <- weed_biomass_clean |>
  ggplot(aes(x = weed_trt, y = interrow_weed_biomass_kg_ha, fill = weed_trt)) +
  geom_boxplot(
    outlier.shape = NA,
    width  = 0.55,
    color  = "black"
  ) +
  geom_jitter(
    width  = 0.12,
    height = 0,
    alpha  = 0.4,
    size   = 1.8,
    color  = "grey30"
  ) +
  facet_wrap(
    ~ site_year,
    nrow     = 1,
    labeller = labeller(site_year = site_year_labs)
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x     = NULL,
    y     = expression(Inter-row~weed~biomass~"(kg"~ha^{-1}*")"),
    title = "Inter-row weed biomass by treatment across site-years"
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 10),
    strip.text  = element_text(face = "bold")
  )

# Print to document
interrow_box_total

# 3) Save figure --------------------------------------------------------
fig_dir_total <- here("analysis", "figs", "weed-biomass", "inter-row")
dir.create(fig_dir_total, showWarnings = FALSE, recursive = TRUE)

ggsave(
  filename = file.path(fig_dir_total,
                       "fig_weed_biomass_interrow_box_by-site-year_kg_ha.png"),
  plot     = wbm_box_total,
  width    = 12,
  height   = 5.5,
  dpi      = 300
)

```
## Model selection
```{r}
# Model selection rule for weed biomass responses:
# - Fit a full GLMM with weed_trt × site_year.
# - Compare to an additive model (weed_trt + site_year) using LRT + ΔAIC:
#     * LRT p < 0.01 or ΔAIC ≥ 4 → keep the interaction model.
#     * LRT p > 0.20 and |ΔAIC| < 2 → treat the interaction as negligible and use the additive model.
#     * Intermediate cases (0.01 ≤ p ≤ 0.20 and |ΔAIC| < 2) = ambiguous evidence:
#         use the additive model for parsimony, but visualize site-year patterns
#         and note the marginal interaction in the text/figure captions.


```


### Total weed biomass
```{r}
### Model testing / selection for total weed biomass (kg ha^-1)

options(contrasts = c("contr.sum", "contr.poly"))

# Base fixed/random structure used for family selection: full interaction
form_all <- weed_biomass_kg_ha ~ weed_trt * site_year + (1 | site_year:block)

## 1) Candidate families / structures (all with interaction) -------------

m_all_tw <- glmmTMB(
  formula = form_all,
  family  = tweedie(link = "log"),
  data    = weed_biomass_clean
)

m_all_nb <- glmmTMB(
  formula = form_all,
  family  = nbinom2(link = "log"),
  data    = weed_biomass_clean
)

m_all_tw_zi <- glmmTMB(
  formula   = form_all,
  family    = tweedie(link = "log"),
  ziformula = ~ 1,
  data      = weed_biomass_clean
)

m_all_tw_disp <- glmmTMB(
  formula     = form_all,
  family      = tweedie(link = "log"),
  ziformula   = ~ 1,
  dispformula = ~ site_year,
  data        = weed_biomass_clean
)

# Safe AIC helper --------------------------------------------------------
safe_aic <- function(mod) {
  out <- try(AIC(mod), silent = TRUE)
  if (inherits(out, "try-error")) return(NA_real_)
  out
}

aic_all <- tibble(
  model = c(
    "ALL: Tweedie",
    "ALL: NB2",
    "ALL: Tweedie + ZI",
    "ALL: Tweedie + ZI + disp(site_year)"
  ),
  AIC = c(
    safe_aic(m_all_tw),
    safe_aic(m_all_nb),
    safe_aic(m_all_tw_zi),
    safe_aic(m_all_tw_disp)
  )
)

kable(
  aic_all,
  digits  = 1,
  caption = "Pooled (all site-years) weed biomass (kg ha^-1): candidate distributional families"
) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

## 2) Choose best family (NA-safe, require > 2 AIC units better) --------

best_all   <- m_all_tw
best_all_n <- "ALL: Tweedie"
best_aic   <- safe_aic(m_all_tw)

cands <- list(
  "ALL: NB2"                            = m_all_nb,
  "ALL: Tweedie + ZI"                   = m_all_tw_zi,
  "ALL: Tweedie + ZI + disp(site_year)" = m_all_tw_disp
)

for (nm in names(cands)) {
  this_aic <- safe_aic(cands[[nm]])
  if (is.finite(this_aic) && (this_aic + 2 < best_aic)) {
    best_all   <- cands[[nm]]
    best_all_n <- nm
    best_aic   <- this_aic
  }
}

cat("Selected family/structure for weed biomass (all site-years):", best_all_n, "\n")

## 3) Within chosen family: additive vs interaction ----------------------

# Interaction model in chosen family (already fit above)
wbm_int <- best_all

# Additive model: same family/ZI/dispersion, simpler fixed effects
wbm_add <- update(
  best_all,
  . ~ weed_trt + site_year + (1 | site_year:block)
)

# Compare additive vs interaction
lrt_wbm   <- anova(wbm_add, wbm_int)   # LRT for interaction
p_int_wbm <- lrt_wbm$`Pr(>Chisq)`[2]

AIC_add  <- AIC(wbm_add)
AIC_int  <- AIC(wbm_int)
deltaAIC <- AIC_add - AIC_int         # > 0 means interaction has lower AIC

# Helper thresholds
p_strong    <- 0.01   # strong evidence for interaction
p_none      <- 0.20   # essentially no evidence for interaction
dAIC_strong <- 4      # "substantial" AIC improvement

# Classify evidence for interaction
interaction_class <- dplyr::case_when(
  p_int_wbm < p_strong | deltaAIC >= dAIC_strong ~ "interaction",
  p_int_wbm > p_none  & abs(deltaAIC) < 2        ~ "additive",
  TRUE                                           ~ "gray_zone"
)

# Choose primary model (default to additive in gray zone)
primary_model_name_wbm <- dplyr::case_when(
  interaction_class == "interaction" ~ "Interaction: weed_trt * site_year",
  TRUE                               ~ "Additive: weed_trt + site_year"
)

weed_glmm <- if (primary_model_name_wbm == "Interaction: weed_trt * site_year") {
  wbm_int
} else {
  wbm_add
}

# AIC table for reporting
aic_wbm_out <- tibble::tibble(
  model = c(
    "Additive: weed_trt + site_year",
    "Interaction: weed_trt * site_year"
  ),
  AIC = c(AIC_add, AIC_int)
) |>
  dplyr::mutate(
    deltaAIC                = AIC - min(AIC),
    Selected                = dplyr::if_else(model == primary_model_name_wbm, "Yes", ""),
    Evidence_for_interaction = interaction_class
  )

kable(
  aic_wbm_out,
  digits  = 2,
  caption = "Weed biomass (kg ha^-1): additive vs interaction (within chosen family)"
) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected primary model for weed biomass (used in emmeans/plots):\n  ",
  primary_model_name_wbm,
  sprintf(
    "  [LRT p = %.3f; ΔAIC (add - int) = %.2f; class = %s]\n",
    p_int_wbm, deltaAIC, interaction_class
  )
)



```
#### Diagnostics (residuals, dispersion, zero inflation)
```{r}
### Diagnostics for total weed biomass model

set.seed(123)
res_wbm <- DHARMa::simulateResiduals(weed_glmm)

plot(res_wbm)
DHARMa::testDispersion(weed_glmm)
DHARMa::testZeroInflation(weed_glmm)

anova_wbm <- car::Anova(weed_glmm, type = 3)
anova_wbm

```


### In-row weed biomass

```{r}
#### In-row pooled (all site-years): model testing / selection ----------

options(contrasts = c("contr.sum", "contr.poly"))

# Model formula: in-row instead of total (full interaction)
form_all_inrow <- inrow_weed_biomass_kg_ha ~ weed_trt * site_year + (1 | site_year:block)

## 1) Candidate families / structures (all with interaction) ------------

m_all_inrow_tw <- glmmTMB(
  formula = form_all_inrow,
  family  = tweedie(link = "log"),
  data    = weed_biomass_clean
)

m_all_inrow_nb <- glmmTMB(
  formula = form_all_inrow,
  family  = nbinom2(link = "log"),
  data    = weed_biomass_clean
)

m_all_inrow_tw_zi <- glmmTMB(
  formula   = form_all_inrow,
  family    = tweedie(link = "log"),
  ziformula = ~ 1,
  data      = weed_biomass_clean
)

m_all_inrow_tw_disp <- glmmTMB(
  formula     = form_all_inrow,
  family      = tweedie(link = "log"),
  ziformula   = ~ 1,
  dispformula = ~ site_year,
  data        = weed_biomass_clean
)

# NA-safe AIC helper (if already defined earlier, you can omit this) ----
safe_aic <- function(mod) {
  out <- try(AIC(mod), silent = TRUE)
  if (inherits(out, "try-error")) return(NA_real_)
  out
}

aic_all_inrow <- tibble(
  model = c(
    "INROW: Tweedie",
    "INROW: NB2",
    "INROW: Tweedie + ZI",
    "INROW: Tweedie + ZI + disp(site_year)"
  ),
  AIC = c(
    safe_aic(m_all_inrow_tw),
    safe_aic(m_all_inrow_nb),
    safe_aic(m_all_inrow_tw_zi),
    safe_aic(m_all_inrow_tw_disp)
  )
)

kable(
  aic_all_inrow,
  digits  = 2,
  caption = "In-row weed biomass (kg ha^-1): candidate distributional families"
) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

## 2) Choose best family (require > 2 AIC units better) -----------------

best_all_inrow   <- m_all_inrow_tw
best_all_inrow_n <- "INROW: Tweedie"
best_aic_inrow   <- safe_aic(m_all_inrow_tw)

cands_inrow <- list(
  "INROW: NB2"                            = m_all_inrow_nb,
  "INROW: Tweedie + ZI"                   = m_all_inrow_tw_zi,
  "INROW: Tweedie + ZI + disp(site_year)" = m_all_inrow_tw_disp
)

for (nm in names(cands_inrow)) {
  this_aic <- safe_aic(cands_inrow[[nm]])
  if (is.finite(this_aic) && (this_aic + 2 < best_aic_inrow)) {
    best_all_inrow   <- cands_inrow[[nm]]
    best_all_inrow_n <- nm
    best_aic_inrow   <- this_aic
  }
}

cat(
  "Selected family/structure for in-row weed biomass (all site-years):",
  best_all_inrow_n, "\n"
)

## 3) Within chosen family: additive vs interaction ---------------------

# Interaction model in chosen family
inrow_int <- best_all_inrow

# Additive model: same family/ZI/dispersion, simpler fixed effects
inrow_add <- update(
  best_all_inrow,
  . ~ weed_trt + site_year + (1 | site_year:block)
)

# Compare additive vs interaction
lrt_inrow   <- anova(inrow_add, inrow_int)
p_int_inrow <- lrt_inrow$`Pr(>Chisq)`[2]

AIC_add_inrow <- AIC(inrow_add)
AIC_int_inrow <- AIC(inrow_int)
deltaAIC_inrow <- AIC_add_inrow - AIC_int_inrow  # > 0 => interaction has lower AIC

# Same decision rule as total biomass -----------------------------------

p_strong_inrow    <- 0.01  # strong evidence
p_none_inrow      <- 0.20  # essentially none
dAIC_strong_inrow <- 4     # substantial AIC drop

interaction_class_inrow <- dplyr::case_when(
  p_int_inrow < p_strong_inrow | deltaAIC_inrow >= dAIC_strong_inrow ~ "interaction",
  p_int_inrow > p_none_inrow  & abs(deltaAIC_inrow) < 2              ~ "additive",
  TRUE                                                               ~ "gray_zone"
)

primary_model_name_inrow <- dplyr::case_when(
  interaction_class_inrow == "interaction" ~ "Interaction: weed_trt * site_year",
  TRUE                                     ~ "Additive: weed_trt + site_year"
)

weed_inrow_glmm <- if (primary_model_name_inrow == "Interaction: weed_trt * site_year") {
  inrow_int
} else {
  inrow_add
}

# AIC table for reporting
aic_inrow_out <- tibble(
  model = c(
    "Additive: weed_trt + site_year",
    "Interaction: weed_trt * site_year"
  ),
  AIC = c(AIC_add_inrow, AIC_int_inrow)
) |>
  mutate(
    deltaAIC                = AIC - min(AIC),
    Selected                = if_else(model == primary_model_name_inrow, "Yes", ""),
    Evidence_for_interaction = interaction_class_inrow
  )

kable(
  aic_inrow_out,
  digits  = 2,
  caption = "In-row weed biomass (kg ha^-1): additive vs interaction (within chosen family)"
) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected primary model for in-row weed biomass (used in emmeans/plots):\n  ",
  primary_model_name_inrow,
  sprintf(
    "  [LRT p = %.3f; ΔAIC (add - int) = %.2f; class = %s]\n",
    p_int_inrow, deltaAIC_inrow, interaction_class_inrow
  )
)


```
#### Diagnostics (residuals, dispersion, zero inflation)
```{r}
## Diagnostics + Type-III tests for chosen in-row model --------------

set.seed(123)
res_inrow <- DHARMa::simulateResiduals(weed_inrow_glmm)
plot(res_inrow)
DHARMa::testDispersion(weed_inrow_glmm)
DHARMa::testZeroInflation(weed_inrow_glmm)

car::Anova(weed_inrow_glmm, type = 3)
```









### Inter-row weed biomass

```{r}
#### Inter-row pooled (all site-years): model testing / selection ----------

options(contrasts = c("contr.sum", "contr.poly"))

# Model formula: inter-row instead of total (full interaction)
form_all_interrow <- interrow_weed_biomass_kg_ha ~ weed_trt * site_year + (1 | site_year:block)

## 1) Candidate families / structures (all with interaction) -------------

m_all_interrow_tw <- glmmTMB(
  formula = form_all_interrow,
  family  = tweedie(link = "log"),
  data    = weed_biomass_clean
)

m_all_interrow_nb <- glmmTMB(
  formula = form_all_interrow,
  family  = nbinom2(link = "log"),
  data    = weed_biomass_clean
)

m_all_interrow_tw_zi <- glmmTMB(
  formula   = form_all_interrow,
  family    = tweedie(link = "log"),
  ziformula = ~ 1,
  data      = weed_biomass_clean
)

m_all_interrow_tw_disp <- glmmTMB(
  formula     = form_all_interrow,
  family      = tweedie(link = "log"),
  ziformula   = ~ 1,
  dispformula = ~ site_year,
  data        = weed_biomass_clean
)

# NA-safe AIC helper (omit if already defined earlier) -------------------
safe_aic <- function(mod) {
  out <- try(AIC(mod), silent = TRUE)
  if (inherits(out, "try-error")) return(NA_real_)
  out
}

aic_all_interrow <- tibble(
  model = c(
    "INTERROW: Tweedie",
    "INTERROW: NB2",
    "INTERROW: Tweedie + ZI",
    "INTERROW: Tweedie + ZI + disp(site_year)"
  ),
  AIC = c(
    safe_aic(m_all_interrow_tw),
    safe_aic(m_all_interrow_nb),
    safe_aic(m_all_interrow_tw_zi),
    safe_aic(m_all_interrow_tw_disp)
  )
)

kable(
  aic_all_interrow,
  digits  = 2,
  caption = "Inter-row weed biomass (kg ha^-1): candidate distributional families"
) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

## 2) Choose best family (require > 2 AIC units better) ------------------

best_all_interrow   <- m_all_interrow_tw
best_all_interrow_n <- "INTERROW: Tweedie"
best_aic_interrow   <- safe_aic(m_all_interrow_tw)

cands_interrow <- list(
  "INTERROW: NB2"                            = m_all_interrow_nb,
  "INTERROW: Tweedie + ZI"                   = m_all_interrow_tw_zi,
  "INTERROW: Tweedie + ZI + disp(site_year)" = m_all_interrow_tw_disp
)

for (nm in names(cands_interrow)) {
  this_aic <- safe_aic(cands_interrow[[nm]])
  if (is.finite(this_aic) && (this_aic + 2 < best_aic_interrow)) {
    best_all_interrow   <- cands_interrow[[nm]]
    best_all_interrow_n <- nm
    best_aic_interrow   <- this_aic
  }
}

cat(
  "Selected family/structure for inter-row weed biomass (all site-years):",
  best_all_interrow_n, "\n"
)

## 3) Within chosen family: additive vs interaction ----------------------

# Interaction model in chosen family
interrow_int <- best_all_interrow

# Additive model: same family/ZI/dispersion, simpler fixed effects
interrow_add <- update(
  best_all_interrow,
  . ~ weed_trt + site_year + (1 | site_year:block)
)

# Compare additive vs interaction
lrt_interrow    <- anova(interrow_add, interrow_int)
p_int_interrow  <- lrt_interrow$`Pr(>Chisq)`[2]

AIC_add_interrow   <- AIC(interrow_add)
AIC_int_interrow   <- AIC(interrow_int)
deltaAIC_interrow  <- AIC_add_interrow - AIC_int_interrow  # > 0 => interaction has lower AIC

# Same decision rule as total & in-row -----------------------------------

p_strong_interrow    <- 0.01  # strong evidence
p_none_interrow      <- 0.20  # essentially none
dAIC_strong_interrow <- 4     # substantial AIC drop

interaction_class_interrow <- dplyr::case_when(
  p_int_interrow < p_strong_interrow | deltaAIC_interrow >= dAIC_strong_interrow ~ "interaction",
  p_int_interrow > p_none_interrow  & abs(deltaAIC_interrow) < 2                 ~ "additive",
  TRUE                                                                          ~ "gray_zone"
)

primary_model_name_interrow <- dplyr::case_when(
  interaction_class_interrow == "interaction" ~ "Interaction: weed_trt * site_year",
  TRUE                                        ~ "Additive: weed_trt + site_year"
)

weed_interrow_glmm <- if (primary_model_name_interrow == "Interaction: weed_trt * site_year") {
  interrow_int
} else {
  interrow_add
}

# AIC table for reporting
aic_interrow_out <- tibble(
  model = c(
    "Additive: weed_trt + site_year",
    "Interaction: weed_trt * site_year"
  ),
  AIC = c(AIC_add_interrow, AIC_int_interrow)
) |>
  mutate(
    deltaAIC                 = AIC - min(AIC),
    Selected                 = if_else(model == primary_model_name_interrow, "Yes", ""),
    Evidence_for_interaction = interaction_class_interrow
  )

kable(
  aic_interrow_out,
  digits  = 2,
  caption = "Inter-row weed biomass (kg ha^-1): additive vs interaction (within chosen family)"
) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected primary model for inter-row weed biomass (used in emmeans/plots):\n  ",
  primary_model_name_interrow,
  sprintf(
    "  [LRT p = %.3f; ΔAIC (add - int) = %.2f; class = %s]\n",
    p_int_interrow, deltaAIC_interrow, interaction_class_interrow
  )
)

```
#### Diagnostics (residuals, dispersion, zero inflation)
```{r}
##  Diagnostics + Type-III tests for chosen inter-row model -----------

set.seed(123)
res_interrow <- DHARMa::simulateResiduals(weed_interrow_glmm)
plot(res_interrow)
DHARMa::testDispersion(weed_interrow_glmm)
DHARMa::testZeroInflation(weed_interrow_glmm)

car::Anova(weed_interrow_glmm, type = 3)

```

# Summary tables

## Total weed biomass
### Post-hoc treatment means and CLDs
```{r}
##### Post-hoc summary table

### Total weed biomass (kg ha^-1) with Fisher's LSD CLDs

# Estimated marginal means for weed_trt on the response scale (kg ha^-1)
emm_wbm <- emmeans(
  weed_glmm,
  ~ weed_trt,
  type = "response"   # back-transform from log link
)

# Tidy emmeans for reporting (rename columns to match other RMDs)
emm_wbm_df <- as_tibble(emm_wbm) |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels)
  ) |>
  rename(
    emmean  = response,   # mean on response scale
    ci_low  = asymp.LCL,  # lower CI
    ci_high = asymp.UCL   # upper CI
  )

# Compact letter display (Fisher's LSD, no adjustment; "a" = highest)
cld_wbm <- cld(
  emm_wbm,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE   # so "a" = highest biomass group(s)
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    .group   = str_trim(.group)
  ) |>
  select(weed_trt, .group)

# Join emmeans + CLDs and format for reporting
emm_wbm_df |>
  left_join(cld_wbm, by = "weed_trt") |>
  select(weed_trt, emmean, SE, ci_low, ci_high, .group) |>
  mutate(
    across(c(emmean, SE, ci_low, ci_high), ~ round(.x, 1))
  ) |>
  kable(
    caption   = "Estimated total weed biomass (kg ha^-1) with 95% CI and Fisher's LSD group letters",
    col.names = c("Treatment", "Mean", "SE", "Lower CI", "Upper CI", "Group")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

emm_wbm_out <- emm_wbm_df |>
  left_join(cld_wbm, by = "weed_trt") |>
  select(weed_trt, emmean, SE, ci_low, ci_high, .group) |>
  mutate(
    across(c(emmean, SE, ci_low, ci_high), ~ round(.x, 1))
  )

```

### Global response summary (ANOVA, AIC, model info)
```{r}
## 0) Directory for all weed-biomass tables ------------------------------

tab_dir_wbm <- here("analysis", "tables", "weed-biomass", "total")
dir.create(tab_dir_wbm, showWarnings = FALSE, recursive = TRUE)


# 1) P-value summary (Location, Treatment, Interaction) ------------------

anova_wbm <- Anova(weed_glmm, type = 3)

anova_wbm_df <- anova_wbm |>
  as.data.frame() |>
  tibble::rownames_to_column("Effect")

p_site <- anova_wbm_df$`Pr(>Chisq)`[anova_wbm_df$Effect == "site_year"]
p_trt  <- anova_wbm_df$`Pr(>Chisq)`[anova_wbm_df$Effect == "weed_trt"]
p_int  <- anova_wbm_df$`Pr(>Chisq)`[anova_wbm_df$Effect == "weed_trt:site_year"]

pvals_wbm <- tibble(
  Effect = c(
    "Location (site_year)",
    "Treatment (weed_trt)"
  ),
  p_raw = c(p_site, p_trt)
)

# Only include the interaction row if the *selected* primary model uses the interaction
if (primary_model_name_wbm == "Interaction: weed_trt * site_year") {
  pvals_wbm <- bind_rows(
    pvals_wbm,
    tibble(
      Effect = "Location × Treatment",
      p_raw  = p_int
    )
  )
}

pvals_wbm <- pvals_wbm |>
  mutate(
    `P-value` = case_when(
      p_raw < 0.001 ~ "<0.001",
      p_raw < 0.01  ~ "<0.01",
      TRUE          ~ sprintf("%.3f", p_raw)
    )
  ) |>
  select(Effect, `P-value`)

# Save ANOVA p-value summary
readr::write_csv(
  pvals_wbm,
  file.path(tab_dir_wbm, "tab_total-wbm_Anova_pvals.csv")
)


## 1b) Likelihood-ratio test: additive vs interaction --------------------

lrt_table_wbm <- tibble(
  Test  = "LRT (additive vs interaction)",
  p_raw = lrt_wbm$`Pr(>Chisq)`[2]
) |>
  mutate(
    `P-value` = case_when(
      p_raw < 0.001 ~ "<0.001",
      p_raw < 0.01  ~ "<0.01",
      TRUE          ~ sprintf("%.3f", p_raw)
    )
  ) |>
  select(Test, `P-value`)

# Save LRT summary
readr::write_csv(
  lrt_table_wbm,
  file.path(tab_dir_wbm, "tab_total-wbm_LRT_add-vs-int.csv")
)

lrt_table_wbm |>
  kable(
    caption   = "Weed biomass (kg ha^-1): Likelihood-ratio test comparing additive vs interaction models",
    col.names = c("Test", "P-value")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 2) Location block: site-year means (model + raw) ----------------------

emm_loc_wbm <- emmeans(
  weed_glmm,
  ~ site_year,
  type = "response"
)

emm_loc_wbm_df <- as_tibble(emm_loc_wbm) |>
  mutate(
    site_year  = as.factor(site_year),
    model_mean = response   # kg ha^-1
  ) |>
  select(site_year, model_mean)

cld_loc_wbm <- cld(
  emm_loc_wbm,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE
) |>
  as_tibble() |>
  mutate(
    site_year = as.factor(site_year),
    loc_CLD   = str_trim(.group)
  ) |>
  select(site_year, loc_CLD)

raw_loc_wbm <- weed_biomass_clean |>
  group_by(site_year) |>
  summarise(
    raw_mean = mean(weed_biomass_kg_ha, na.rm = TRUE),
    .groups  = "drop"
  ) |>
  mutate(site_year = as.factor(site_year))

loc_summary_wbm <- emm_loc_wbm_df |>
  left_join(cld_loc_wbm, by = "site_year") |>
  left_join(raw_loc_wbm, by = "site_year") |>
  mutate(
    model_mean = round(model_mean, 1),
    raw_mean   = round(raw_mean, 1),
    raw_CLD    = loc_CLD
  ) |>
  arrange(site_year)

# Save location summary
readr::write_csv(
  loc_summary_wbm,
  file.path(tab_dir_wbm, "tab_total-wbm_location_means_CLD.csv")
)

loc_summary_wbm |>
  kable(
    caption   = "Weed biomass (kg ha^-1): location (site-year) means with CLDs",
    col.names = c("Site-year", "Model mean", "Model CLD", "Raw mean", "Raw CLD")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 3) Treatment block: means (model + raw) -------------------------------

emm_wbm <- emmeans(
  weed_glmm,
  ~ weed_trt,
  type = "response"
)

emm_trt_wbm_df <- as_tibble(emm_wbm) |>
  mutate(
    weed_trt   = factor(weed_trt, levels = mow_levels),
    model_mean = response   # kg ha^-1
  ) |>
  select(weed_trt, model_mean)

cld_wbm <- cld(
  emm_wbm,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    trt_CLD  = str_trim(.group)
  ) |>
  select(weed_trt, trt_CLD)

raw_trt_wbm <- weed_biomass_clean |>
  group_by(weed_trt) |>
  summarise(
    raw_mean = mean(weed_biomass_kg_ha, na.rm = TRUE),
    .groups  = "drop"
  ) |>
  mutate(weed_trt = factor(weed_trt, levels = mow_levels))

trt_summary_wbm <- emm_trt_wbm_df |>
  left_join(cld_wbm,     by = "weed_trt") |>
  left_join(raw_trt_wbm, by = "weed_trt") |>
  mutate(
    model_mean = round(model_mean, 1),
    raw_mean   = round(raw_mean, 1),
    raw_CLD    = trt_CLD
  ) |>
  arrange(weed_trt)

# Save treatment summary
readr::write_csv(
  trt_summary_wbm,
  file.path(tab_dir_wbm, "tab_total-wbm_treatment_means_CLD.csv")
)

trt_summary_wbm |>
  kable(
    caption   = "Weed biomass (kg ha^-1): treatment means with CLDs",
    col.names = c("Treatment", "Model mean", "Model CLD", "Raw mean", "Raw CLD")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 4) Interaction block: site-year × treatment means ---------------------

emm_sy_wbm <- emmeans(
  weed_glmm,
  ~ weed_trt | site_year,
  type = "response"
)

emm_sy_wbm_df <- as_tibble(emm_sy_wbm) |>
  mutate(
    weed_trt   = factor(weed_trt, levels = mow_levels),
    site_year  = as.factor(site_year),
    model_mean = response   # kg ha^-1
  ) |>
  select(site_year, weed_trt, model_mean)

cld_sy_wbm <- cld(
  emm_sy_wbm,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE
) |>
  as_tibble() |>
  mutate(
    weed_trt  = factor(weed_trt, levels = mow_levels),
    site_year = as.factor(site_year),
    int_CLD   = str_trim(.group)
  ) |>
  select(site_year, weed_trt, int_CLD)

raw_sy_wbm <- weed_biomass_clean |>
  group_by(site_year, weed_trt) |>
  summarise(
    raw_mean = mean(weed_biomass_kg_ha, na.rm = TRUE),
    .groups  = "drop"
  ) |>
  mutate(
    site_year = as.factor(site_year),
    weed_trt  = factor(weed_trt, levels = mow_levels)
  )

int_summary_wbm <- emm_sy_wbm_df |>
  left_join(cld_sy_wbm, by = c("site_year", "weed_trt")) |>
  left_join(raw_sy_wbm, by = c("site_year", "weed_trt")) |>
  mutate(
    model_mean = round(model_mean, 1),
    raw_mean   = round(raw_mean, 1),
    raw_CLD    = int_CLD
  ) |>
  arrange(site_year, weed_trt)

# Save interaction summary
readr::write_csv(
  int_summary_wbm,
  file.path(tab_dir_wbm, "tab_total-wbm_site-year_treatment_means_CLD.csv")
)

int_summary_wbm |>
  kable(
    caption   = "Weed biomass (kg ha^-1): site-year × treatment means with CLDs",
    col.names = c(
      "Site-year", "Treatment",
      "Model mean", "Model CLD",
      "Raw mean",   "Raw CLD"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 5) Model-info row for global response summary -------------------------

model_info_wbm <- tibble::tibble(
  response_label    = "Weed biomass (kg ha^-1), model-predicted means*",
  family_structure  = best_all_n,
  fixed_effects     = primary_model_name_wbm,
  random_effects    = "(1 | site_year:block)",
  AIC_additive      = round(AIC_add, 2),
  AIC_interaction   = round(AIC_int, 2),
  deltaAIC_add_int  = round(deltaAIC, 2),
  LRT_p_int_raw     = p_int_wbm,
  LRT_p_int_label   = dplyr::case_when(
    p_int_wbm < 0.001 ~ "<0.001",
    p_int_wbm < 0.01  ~ "<0.01",
    TRUE              ~ sprintf("%.3f", p_int_wbm)
  ),
  interaction_class = interaction_class
)

readr::write_csv(
  model_info_wbm,
  file.path(tab_dir_wbm, "tab_total-wbm_model-info.csv")
)

## 6) OPTIONAL: Combine all TOTAL weed-biomass tables into one workbook --
## (requires: library(writexl))

wbm_total_tables <- list(
  Anova_pvals         = pvals_wbm,
  LRT_add_vs_int      = lrt_table_wbm,
  Location_means_CLD  = loc_summary_wbm,
  Treatment_means_CLD = trt_summary_wbm,
  SiteYear_trt_means  = int_summary_wbm,
  Model_info          = model_info_wbm
)

write_xlsx(
  wbm_total_tables,
  path = file.path(tab_dir_wbm, "total-weed-biomass_all-tables.xlsx")
)


```

## In-row weed biomass
### Post-hoc treatment means and CLDs
```{r}
#### Post-hoc summary table

### In-row weed biomass (kg ha^-1) with Fisher's LSD CLDs

# Estimated marginal means for weed_trt on the response scale (kg ha^-1)
emm_inrow <- emmeans(
  weed_inrow_glmm,
  ~ weed_trt,
  type = "response"   # back-transform from log link
)

# Tidy emmeans for reporting (rename columns to match other RMDs)
emm_inrow_df <- as_tibble(emm_inrow) |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels)
  ) |>
  rename(
    emmean  = response,   # mean on response scale
    ci_low  = asymp.LCL,  # lower CI
    ci_high = asymp.UCL   # upper CI
  )

# Compact letter display (Fisher's LSD, no adjustment; "a" = highest)
cld_inrow <- cld(
  emm_inrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE   # so "a" = highest biomass group(s)
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    .group   = str_trim(.group)
  ) |>
  select(weed_trt, .group)

# Join emmeans + CLDs and format for reporting
emm_inrow_df |>
  left_join(cld_inrow, by = "weed_trt") |>
  select(weed_trt, emmean, SE, ci_low, ci_high, .group) |>
  mutate(
    across(c(emmean, SE, ci_low, ci_high), ~ round(.x, 1))
  ) |>
  kable(
    caption   = "Estimated in-row weed biomass (kg ha^-1) with 95% CI and Fisher's LSD group letters",
    col.names = c("Treatment", "Mean", "SE", "Lower CI", "Upper CI", "Group")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

emm_inrow_out <- emm_inrow_df |>
  left_join(cld_inrow, by = "weed_trt") |>
  select(weed_trt, emmean, SE, ci_low, ci_high, .group) |>
  mutate(
    across(c(emmean, SE, ci_low, ci_high), ~ round(.x, 1))
  )

```

### Global response summary (ANOVA, AIC, model info)
```{r}
## 0) Directory for all *in-row* weed-biomass tables ----------------------

tab_dir_inrow <- here("analysis", "tables", "weed-biomass", "in-row")
dir.create(tab_dir_inrow, showWarnings = FALSE, recursive = TRUE)


# 1) P-value summary (Location, Treatment, Interaction) ------------------

anova_inrow <- Anova(weed_inrow_glmm, type = 3)

anova_inrow_df <- anova_inrow |>
  as.data.frame() |>
  tibble::rownames_to_column("Effect")

p_site_inrow <- anova_inrow_df$`Pr(>Chisq)`[anova_inrow_df$Effect == "site_year"]
p_trt_inrow  <- anova_inrow_df$`Pr(>Chisq)`[anova_inrow_df$Effect == "weed_trt"]
p_int_inrow3 <- anova_inrow_df$`Pr(>Chisq)`[anova_inrow_df$Effect == "weed_trt:site_year"]

pvals_inrow <- tibble(
  Effect = c(
    "Location (site_year)",
    "Treatment (weed_trt)"
  ),
  p_raw = c(p_site_inrow, p_trt_inrow)
)

# Only include the interaction row if the *selected* primary model uses the interaction
if (primary_model_name_inrow == "Interaction: weed_trt * site_year") {
  pvals_inrow <- bind_rows(
    pvals_inrow,
    tibble(
      Effect = "Location × Treatment",
      p_raw  = p_int_inrow3
    )
  )
}

pvals_inrow <- pvals_inrow |>
  mutate(
    `P-value` = case_when(
      p_raw < 0.001 ~ "<0.001",
      p_raw < 0.01  ~ "<0.01",
      TRUE          ~ sprintf("%.3f", p_raw)
    )
  ) |>
  select(Effect, `P-value`)

# Save ANOVA p-value summary
readr::write_csv(
  pvals_inrow,
  file.path(tab_dir_inrow, "tab_inrow-wbm_Anova_pvals.csv")
)


## 1b) Likelihood-ratio test: additive vs interaction --------------------

lrt_table_inrow <- tibble(
  Test  = "LRT (additive vs interaction)",
  p_raw = p_int_inrow   # from your in-row model selection chunk
) |>
  mutate(
    `P-value` = case_when(
      p_raw < 0.001 ~ "<0.001",
      p_raw < 0.01  ~ "<0.01",
      TRUE          ~ sprintf("%.3f", p_raw)
    )
  ) |>
  select(Test, `P-value`)

# Save LRT summary
readr::write_csv(
  lrt_table_inrow,
  file.path(tab_dir_inrow, "tab_inrow-wbm_LRT_add-vs-int.csv")
)

lrt_table_inrow |>
  kable(
    caption   = "In-row weed biomass (kg ha^-1): Likelihood-ratio test comparing additive vs interaction models",
    col.names = c("Test", "P-value")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 2) Location block: site-year means (model + raw) ----------------------

emm_loc_inrow <- emmeans(
  weed_inrow_glmm,
  ~ site_year,
  type = "response"
)

emm_loc_inrow_df <- as_tibble(emm_loc_inrow) |>
  mutate(
    site_year  = as.factor(site_year),
    model_mean = response   # kg ha^-1
  ) |>
  select(site_year, model_mean)

cld_loc_inrow <- cld(
  emm_loc_inrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE
) |>
  as_tibble() |>
  mutate(
    site_year = as.factor(site_year),
    loc_CLD   = str_trim(.group)
  ) |>
  select(site_year, loc_CLD)

raw_loc_inrow <- weed_biomass_clean |>
  group_by(site_year) |>
  summarise(
    raw_mean = mean(inrow_weed_biomass_kg_ha, na.rm = TRUE),
    .groups  = "drop"
  ) |>
  mutate(site_year = as.factor(site_year))

loc_summary_inrow <- emm_loc_inrow_df |>
  left_join(cld_loc_inrow, by = "site_year") |>
  left_join(raw_loc_inrow, by = "site_year") |>
  mutate(
    model_mean = round(model_mean, 1),
    raw_mean   = round(raw_mean, 1),
    raw_CLD    = loc_CLD
  ) |>
  arrange(site_year)

# Save location summary
readr::write_csv(
  loc_summary_inrow,
  file.path(tab_dir_inrow, "tab_inrow-wbm_location_means_CLD.csv")
)

loc_summary_inrow |>
  kable(
    caption   = "In-row weed biomass (kg ha^-1): location (site-year) means with CLDs",
    col.names = c("Site-year", "Model mean", "Model CLD", "Raw mean", "Raw CLD")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 3) Treatment block: means (model + raw) -------------------------------

emm_inrow <- emmeans(
  weed_inrow_glmm,
  ~ weed_trt,
  type = "response"
)

emm_trt_inrow_df <- as_tibble(emm_inrow) |>
  mutate(
    weed_trt   = factor(weed_trt, levels = mow_levels),
    model_mean = response   # kg ha^-1
  ) |>
  select(weed_trt, model_mean)

cld_inrow_trt <- cld(
  emm_inrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    trt_CLD  = str_trim(.group)
  ) |>
  select(weed_trt, trt_CLD)

raw_trt_inrow <- weed_biomass_clean |>
  group_by(weed_trt) |>
  summarise(
    raw_mean = mean(inrow_weed_biomass_kg_ha, na.rm = TRUE),
    .groups  = "drop"
  ) |>
  mutate(weed_trt = factor(weed_trt, levels = mow_levels))

trt_summary_inrow <- emm_trt_inrow_df |>
  left_join(cld_inrow_trt, by = "weed_trt") |>
  left_join(raw_trt_inrow, by = "weed_trt") |>
  mutate(
    model_mean = round(model_mean, 1),
    raw_mean   = round(raw_mean, 1),
    raw_CLD    = trt_CLD
  ) |>
  arrange(weed_trt)

# Save treatment summary
readr::write_csv(
  trt_summary_inrow,
  file.path(tab_dir_inrow, "tab_inrow-wbm_treatment_means_CLD.csv")
)

trt_summary_inrow |>
  kable(
    caption   = "In-row weed biomass (kg ha^-1): treatment means with CLDs",
    col.names = c("Treatment", "Model mean", "Model CLD", "Raw mean", "Raw CLD")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 4) Interaction block: site-year × treatment means ---------------------

emm_sy_inrow <- emmeans(
  weed_inrow_glmm,
  ~ weed_trt | site_year,
  type = "response"
)

emm_sy_inrow_df <- as_tibble(emm_sy_inrow) |>
  mutate(
    weed_trt   = factor(weed_trt, levels = mow_levels),
    site_year  = as.factor(site_year),
    model_mean = response   # kg ha^-1
  ) |>
  select(site_year, weed_trt, model_mean)

cld_sy_inrow <- cld(
  emm_sy_inrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE
) |>
  as_tibble() |>
  mutate(
    weed_trt  = factor(weed_trt, levels = mow_levels),
    site_year = as.factor(site_year),
    int_CLD   = str_trim(.group)
  ) |>
  select(site_year, weed_trt, int_CLD)

raw_sy_inrow <- weed_biomass_clean |>
  group_by(site_year, weed_trt) |>
  summarise(
    raw_mean = mean(inrow_weed_biomass_kg_ha, na.rm = TRUE),
    .groups  = "drop"
  ) |>
  mutate(
    site_year = as.factor(site_year),
    weed_trt  = factor(weed_trt, levels = mow_levels)
  )

int_summary_inrow <- emm_sy_inrow_df |>
  left_join(cld_sy_inrow, by = c("site_year", "weed_trt")) |>
  left_join(raw_sy_inrow, by = c("site_year", "weed_trt")) |>
  mutate(
    model_mean = round(model_mean, 1),
    raw_mean   = round(raw_mean, 1),
    raw_CLD    = int_CLD
  ) |>
  arrange(site_year, weed_trt)

# Save interaction summary
readr::write_csv(
  int_summary_inrow,
  file.path(tab_dir_inrow, "tab_inrow-wbm_site-year_treatment_means_CLD.csv")
)

int_summary_inrow |>
  kable(
    caption   = "In-row weed biomass (kg ha^-1): site-year × treatment means with CLDs",
    col.names = c(
      "Site-year", "Treatment",
      "Model mean", "Model CLD",
      "Raw mean",   "Raw CLD"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 5) Model-info row for global response summary -------------------------

model_info_inrow <- tibble::tibble(
  response_label    = "In-row weed biomass (kg ha^-1), model-predicted means*",
  family_structure  = best_all_inrow_n,
  fixed_effects     = primary_model_name_inrow,
  random_effects    = "(1 | site_year:block)",
  AIC_additive      = round(AIC_add_inrow, 2),
  AIC_interaction   = round(AIC_int_inrow, 2),
  deltaAIC_add_int  = round(deltaAIC_inrow, 2),
  LRT_p_int_raw     = p_int_inrow,
  LRT_p_int_label   = dplyr::case_when(
    p_int_inrow < 0.001 ~ "<0.001",
    p_int_inrow < 0.01  ~ "<0.01",
    TRUE                ~ sprintf("%.3f", p_int_inrow)
  ),
  interaction_class = interaction_class_inrow
)

readr::write_csv(
  model_info_inrow,
  file.path(tab_dir_inrow, "tab_inrow-wbm_model-info.csv")
)

## OPTIONAL: Combine all IN-ROW weed-biomass tables into one workbook ----

wbm_inrow_tables <- list(
  Anova_pvals         = pvals_inrow,
  LRT_add_vs_int      = lrt_table_inrow,
  Location_means_CLD  = loc_summary_inrow,
  Treatment_means_CLD = trt_summary_inrow,
  SiteYear_trt_means  = int_summary_inrow,
  Model_info          = model_info_inrow
)

write_xlsx(
  wbm_inrow_tables,
  path = file.path(tab_dir_inrow, "inrow-weed-biomass_all-tables.xlsx")
)


```

## Inter-row weed biomass
### Post-hoc treatment means and CLDs
```{r}
#### Post-hoc summary table

### Inter-row weed biomass (kg ha^-1) with Fisher's LSD CLDs

# Estimated marginal means for weed_trt on the response scale (kg ha^-1)
emm_interrow <- emmeans(
  weed_interrow_glmm,
  ~ weed_trt,
  type = "response"   # back-transform from log link
)

# Tidy emmeans for reporting (rename columns to match other RMDs)
emm_interrow_df <- as_tibble(emm_interrow) |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels)
  ) |>
  rename(
    emmean  = response,   # mean on response scale
    ci_low  = asymp.LCL,  # lower CI
    ci_high = asymp.UCL   # upper CI
  )

# Compact letter display (Fisher's LSD, no adjustment; "a" = highest)
cld_interrow <- cld(
  emm_interrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE   # so "a" = highest biomass group(s)
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    .group   = str_trim(.group)
  ) |>
  select(weed_trt, .group)

# Join emmeans + CLDs and format for reporting
emm_interrow_df |>
  left_join(cld_interrow, by = "weed_trt") |>
  select(weed_trt, emmean, SE, ci_low, ci_high, .group) |>
  mutate(
    across(c(emmean, SE, ci_low, ci_high), ~ round(.x, 1))
  ) |>
  kable(
    caption   = "Estimated inter-row weed biomass (kg ha^-1) with 95% CI and Fisher's LSD group letters",
    col.names = c("Treatment", "Mean", "SE", "Lower CI", "Upper CI", "Group")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

emm_interrow_out <- emm_interrow_df |>
  left_join(cld_interrow, by = "weed_trt") |>
  select(weed_trt, emmean, SE, ci_low, ci_high, .group) |>
  mutate(
    across(c(emmean, SE, ci_low, ci_high), ~ round(.x, 1))
  )

```

### Global response summary (ANOVA, AIC, model info)
```{r}
## 0) Directory for all *inter-row* weed-biomass tables -------------------

tab_dir_interrow <- here("analysis", "tables", "weed-biomass", "inter-row")
dir.create(tab_dir_interrow, showWarnings = FALSE, recursive = TRUE)


# 1) P-value summary (Location, Treatment, Interaction) ------------------

anova_interrow <- Anova(weed_interrow_glmm, type = 3)

anova_interrow_df <- anova_interrow |>
  as.data.frame() |>
  tibble::rownames_to_column("Effect")

p_site_interrow <- anova_interrow_df$`Pr(>Chisq)`[anova_interrow_df$Effect == "site_year"]
p_trt_interrow  <- anova_interrow_df$`Pr(>Chisq)`[anova_interrow_df$Effect == "weed_trt"]
p_int_interrow3 <- anova_interrow_df$`Pr(>Chisq)`[anova_interrow_df$Effect == "weed_trt:site_year"]

pvals_interrow <- tibble(
  Effect = c(
    "Location (site_year)",
    "Treatment (weed_trt)"
  ),
  p_raw = c(p_site_interrow, p_trt_interrow)
)

# Only include the interaction row if the *selected* primary model uses the interaction
if (primary_model_name_interrow == "Interaction: weed_trt * site_year") {
  pvals_interrow <- bind_rows(
    pvals_interrow,
    tibble(
      Effect = "Location × Treatment",
      p_raw  = p_int_interrow3
    )
  )
}

pvals_interrow <- pvals_interrow |>
  mutate(
    `P-value` = case_when(
      p_raw < 0.001 ~ "<0.001",
      p_raw < 0.01  ~ "<0.01",
      TRUE          ~ sprintf("%.3f", p_raw)
    )
  ) |>
  select(Effect, `P-value`)

# Save ANOVA p-value summary
readr::write_csv(
  pvals_interrow,
  file.path(tab_dir_interrow, "tab_interrow-wbm_Anova_pvals.csv")
)


## 1b) Likelihood-ratio test: additive vs interaction --------------------

lrt_table_interrow <- tibble(
  Test  = "LRT (additive vs interaction)",
  p_raw = p_int_interrow   # from your inter-row model selection chunk
) |>
  mutate(
    `P-value` = case_when(
      p_raw < 0.001 ~ "<0.001",
      p_raw < 0.01  ~ "<0.01",
      TRUE          ~ sprintf("%.3f", p_raw)
    )
  ) |>
  select(Test, `P-value`)

# Save LRT summary
readr::write_csv(
  lrt_table_interrow,
  file.path(tab_dir_interrow, "tab_interrow-wbm_LRT_add-vs-int.csv")
)

lrt_table_interrow |>
  kable(
    caption   = "Inter-row weed biomass (kg ha^-1): Likelihood-ratio test comparing additive vs interaction models",
    col.names = c("Test", "P-value")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 2) Location block: site-year means (model + raw) ----------------------

emm_loc_interrow <- emmeans(
  weed_interrow_glmm,
  ~ site_year,
  type = "response"
)

emm_loc_interrow_df <- as_tibble(emm_loc_interrow) |>
  mutate(
    site_year  = as.factor(site_year),
    model_mean = response   # kg ha^-1
  ) |>
  select(site_year, model_mean)

cld_loc_interrow <- cld(
  emm_loc_interrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE
) |>
  as_tibble() |>
  mutate(
    site_year = as.factor(site_year),
    loc_CLD   = str_trim(.group)
  ) |>
  select(site_year, loc_CLD)

raw_loc_interrow <- weed_biomass_clean |>
  group_by(site_year) |>
  summarise(
    raw_mean = mean(interrow_weed_biomass_kg_ha, na.rm = TRUE),
    .groups  = "drop"
  ) |>
  mutate(site_year = as.factor(site_year))

loc_summary_interrow <- emm_loc_interrow_df |>
  left_join(cld_loc_interrow, by = "site_year") |>
  left_join(raw_loc_interrow, by = "site_year") |>
  mutate(
    model_mean = round(model_mean, 1),
    raw_mean   = round(raw_mean, 1),
    raw_CLD    = loc_CLD
  ) |>
  arrange(site_year)

# Save location summary
readr::write_csv(
  loc_summary_interrow,
  file.path(tab_dir_interrow, "tab_interrow-wbm_location_means_CLD.csv")
)

loc_summary_interrow |>
  kable(
    caption   = "Inter-row weed biomass (kg ha^-1): location (site-year) means with CLDs",
    col.names = c("Site-year", "Model mean", "Model CLD", "Raw mean", "Raw CLD")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 3) Treatment block: means (model + raw) -------------------------------

emm_interrow <- emmeans(
  weed_interrow_glmm,
  ~ weed_trt,
  type = "response"
)

emm_trt_interrow_df <- as_tibble(emm_interrow) |>
  mutate(
    weed_trt   = factor(weed_trt, levels = mow_levels),
    model_mean = response   # kg ha^-1
  ) |>
  select(weed_trt, model_mean)

cld_interrow_trt <- cld(
  emm_interrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    trt_CLD  = str_trim(.group)
  ) |>
  select(weed_trt, trt_CLD)

raw_trt_interrow <- weed_biomass_clean |>
  group_by(weed_trt) |>
  summarise(
    raw_mean = mean(interrow_weed_biomass_kg_ha, na.rm = TRUE),
    .groups  = "drop"
  ) |>
  mutate(weed_trt = factor(weed_trt, levels = mow_levels))

trt_summary_interrow <- emm_trt_interrow_df |>
  left_join(cld_interrow_trt, by = "weed_trt") |>
  left_join(raw_trt_interrow, by = "weed_trt") |>
  mutate(
    model_mean = round(model_mean, 1),
    raw_mean   = round(raw_mean, 1),
    raw_CLD    = trt_CLD
  ) |>
  arrange(weed_trt)

# Save treatment summary
readr::write_csv(
  trt_summary_interrow,
  file.path(tab_dir_interrow, "tab_interrow-wbm_treatment_means_CLD.csv")
)

trt_summary_interrow |>
  kable(
    caption   = "Inter-row weed biomass (kg ha^-1): treatment means with CLDs",
    col.names = c("Treatment", "Model mean", "Model CLD", "Raw mean", "Raw CLD")
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 4) Interaction block: site-year × treatment means ---------------------

emm_sy_interrow <- emmeans(
  weed_interrow_glmm,
  ~ weed_trt | site_year,
  type = "response"
)

emm_sy_interrow_df <- as_tibble(emm_sy_interrow) |>
  mutate(
    weed_trt   = factor(weed_trt, levels = mow_levels),
    site_year  = as.factor(site_year),
    model_mean = response   # kg ha^-1
  ) |>
  select(site_year, weed_trt, model_mean)

cld_sy_interrow <- cld(
  emm_sy_interrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE
) |>
  as_tibble() |>
  mutate(
    weed_trt  = factor(weed_trt, levels = mow_levels),
    site_year = as.factor(site_year),
    int_CLD   = str_trim(.group)
  ) |>
  select(site_year, weed_trt, int_CLD)

raw_sy_interrow <- weed_biomass_clean |>
  group_by(site_year, weed_trt) |>
  summarise(
    raw_mean = mean(interrow_weed_biomass_kg_ha, na.rm = TRUE),
    .groups  = "drop"
  ) |>
  mutate(
    site_year = as.factor(site_year),
    weed_trt  = factor(weed_trt, levels = mow_levels)
  )

int_summary_interrow <- emm_sy_interrow_df |>
  left_join(cld_sy_interrow, by = c("site_year", "weed_trt")) |>
  left_join(raw_sy_interrow, by = c("site_year", "weed_trt")) |>
  mutate(
    model_mean = round(model_mean, 1),
    raw_mean   = round(raw_mean, 1),
    raw_CLD    = int_CLD
  ) |>
  arrange(site_year, weed_trt)

# Save interaction summary
readr::write_csv(
  int_summary_interrow,
  file.path(tab_dir_interrow, "tab_interrow-wbm_site-year_treatment_means_CLD.csv")
)

int_summary_interrow |>
  kable(
    caption   = "Inter-row weed biomass (kg ha^-1): site-year × treatment means with CLDs",
    col.names = c(
      "Site-year", "Treatment",
      "Model mean", "Model CLD",
      "Raw mean",   "Raw CLD"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


## 5) Model-info row for global response summary -------------------------

model_info_interrow <- tibble::tibble(
  response_label    = "Inter-row weed biomass (kg ha^-1), model-predicted means*",
  family_structure  = best_all_interrow_n,
  fixed_effects     = primary_model_name_interrow,
  random_effects    = "(1 | site_year:block)",
  AIC_additive      = round(AIC_add_interrow, 2),
  AIC_interaction   = round(AIC_int_interrow, 2),
  deltaAIC_add_int  = round(deltaAIC_interrow, 2),
  LRT_p_int_raw     = p_int_interrow,
  LRT_p_int_label   = dplyr::case_when(
    p_int_interrow < 0.001 ~ "<0.001",
    p_int_interrow < 0.01  ~ "<0.01",
    TRUE                   ~ sprintf("%.3f", p_int_interrow)
  ),
  interaction_class = interaction_class_interrow
)

readr::write_csv(
  model_info_interrow,
  file.path(tab_dir_interrow, "tab_interrow-wbm_model-info.csv")
)

## OPTIONAL: Combine all INTER-ROW weed-biomass tables into one workbook ----

wbm_inter_tables <- list(
  Anova_pvals         = pvals_interrow,
  LRT_add_vs_int      = lrt_table_interrow,
  Location_means_CLD  = loc_summary_interrow,
  Treatment_means_CLD = trt_summary_interrow,
  SiteYear_trt_means  = int_summary_interrow,
  Model_info          = model_info_interrow
)

write_xlsx(
  wbm_inter_tables,
  path = file.path(tab_dir_interrow, "interrow-weed-biomass_all-tables.xlsx")
)
```



# Figures
## Total weed biomass
### Model-predicted means (kg ha⁻¹)
```{r}
## 0) Directory for total weed-biomass figures ---------------------------
fig_dir_total <- here("analysis", "figs", "weed-biomass", "total")
dir.create(fig_dir_total, showWarnings = FALSE, recursive = TRUE)

## Figure: Weed biomass by weed management treatment 
## (model-predicted means ± SE, model CLDs)

# 1) Model-based emmeans by treatment -----------------------------------
emm_wbm <- emmeans(
  weed_glmm,
  ~ weed_trt,
  type = "response"   # back-transform from link scale (kg ha^-1)
)

# Tidy emmeans and prepare SE-based error bars
emm_wbm_df <- as_tibble(emm_wbm) |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    mean     = response,
    ymin     = pmax(mean - SE, 0),
    ymax     = mean + SE
  )

# 2) Model-based CLDs for treatment main effect -------------------------
cld_wbm <- cld(
  emm_wbm,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE   # so "a" = highest biomass group(s)
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    .group   = str_trim(.group)
  ) |>
  select(weed_trt, .group)

# 3) Join model-predicted means + CLDs ----------------------------------
plot_df_wbm_model <- emm_wbm_df |>
  left_join(cld_wbm, by = "weed_trt")

# 4) Plot ---------------------------------------------------------------
fig_wbm_total_model <- ggplot(
  plot_df_wbm_model,
  aes(x = weed_trt, y = mean, fill = weed_trt)
) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 6
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x       = NULL,
    y       = expression(Weed~biomass~"(kg"~ha^{-1}*")"),
    title   = "Weed biomass by weed management",
    caption = "Bars show GLMM-predicted marginal means (kg"~ha^{-1}*") ± SE; letters denote Fisher’s LSD groups (α = 0.05)."
  ) +
  theme_classic(base_size = 18) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_wbm_total_model  # print in the Rmd

# 5) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_total, "fig_weed_biomass_total_model_kg_ha.png"),
  plot     = fig_wbm_total_model,
  width    = 9,
  height   = 5.5,
  dpi      = 300
)

```
### Raw means (kg ha⁻¹)
```{r}
## 0) Directory for total weed-biomass figures ---------------------------
fig_dir_total <- here("analysis", "figs", "weed-biomass", "total")
dir.create(fig_dir_total, showWarnings = FALSE, recursive = TRUE)

## Figure: Weed biomass by weed management treatment (raw means ± SE, model CLDs)

# 1) Raw means and SE by treatment --------------------------------------
raw_wbm_summary <- weed_biomass_clean |>
  group_by(weed_trt) |>
  summarise(
    n    = n(),
    mean = mean(weed_biomass_kg_ha, na.rm = TRUE),
    sd   = sd(weed_biomass_kg_ha, na.rm = TRUE),
    se   = sd / sqrt(n),
    .groups = "drop"
  ) |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    ymin     = pmax(mean - se, 0),
    ymax     = mean + se
  )

# 2) Model-based CLDs for treatment main effect -------------------------
emm_wbm <- emmeans(
  weed_glmm,
  ~ weed_trt,
  type = "response"   # back-transform from link scale
)

cld_wbm <- cld(
  emm_wbm,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE   # so "a" = highest biomass group(s)
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    .group   = str_trim(.group)
  ) |>
  select(weed_trt, .group)

# 3) Join raw means + model CLDs ----------------------------------------
plot_df_wbm_raw <- raw_wbm_summary |>
  left_join(cld_wbm, by = "weed_trt")

# 4) Plot ---------------------------------------------------------------
fig_wbm_total_raw <- ggplot(plot_df_wbm_raw, aes(x = weed_trt, y = mean, fill = weed_trt)) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 6
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x       = NULL,
    y       = expression(Weed~biomass~"(kg"~ha^{-1}*")"),
    title   = "Weed biomass by weed management",
    caption = "Bars show raw means (kg"~ha^{-1}*") ± SE; similar letters indicate no significant difference (Fisher’s LSD test, α = 0.05)."
  ) +
  theme_classic(base_size = 18) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_wbm_total_raw  # print in the Rmd

# 5) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_total, "fig_weed_biomass_total_raw_kg_ha.png"),
  plot     = fig_wbm_total_raw,
  width    = 9,
  height   = 5.5,
  dpi      = 300
)

```

### Extension figures (lb ac⁻¹)
#### Model-predicted means (lb ac⁻¹)

```{r}
### Model-predicted means (lb ac^-1)
## 0) Directory for total weed-biomass figures ---------------------------
fig_dir_total <- here("analysis", "figs", "weed-biomass", "total")
dir.create(fig_dir_total, showWarnings = FALSE, recursive = TRUE)

# kg/ha → lb/ac conversion factor
kg_ha_to_lb_ac <- 0.892179

## Figure: Weed biomass by weed management treatment 
## (model-predicted means ± SE, model CLDs; lb ac^-1)

# 1) Model-based emmeans by treatment (still on kg ha^-1 scale) ---------
emm_wbm <- emmeans(
  weed_glmm,
  ~ weed_trt,
  type = "response"   # back-transform from link scale (kg ha^-1)
)

# 2) Tidy emmeans and convert to lb ac^-1 -------------------------------
emm_wbm_df <- as_tibble(emm_wbm) |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    mean_kg  = response,
    se_kg    = SE,
    # convert to lb ac^-1
    mean     = mean_kg * kg_ha_to_lb_ac,
    SE       = se_kg   * kg_ha_to_lb_ac,
    ymin     = pmax(mean - SE, 0),
    ymax     = mean + SE
  )

# 3) Model-based CLDs for treatment main effect -------------------------
cld_wbm <- cld(
  emm_wbm,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE   # so "a" = highest biomass group(s)
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    .group   = str_trim(.group)
  ) |>
  select(weed_trt, .group)

# 4) Join model-predicted means (lb ac^-1) + CLDs -----------------------
plot_df_wbm_model_lb <- emm_wbm_df |>
  left_join(cld_wbm, by = "weed_trt")

# 5) Plot ---------------------------------------------------------------
fig_wbm_total_model_lb <- ggplot(
  plot_df_wbm_model_lb,
  aes(x = weed_trt, y = mean, fill = weed_trt)
) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 6
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x       = NULL,
    y       = expression(Weed~biomass~"(lb"~ac^{-1}*")"),
    title   = "Weed biomass by weed management",
    caption = "Bars show GLMM-predicted marginal means (lb"~ac^{-1}*") ± SE; letters denote Fisher’s LSD groups (α = 0.05)."
  ) +
  theme_classic(base_size = 18) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_wbm_total_model_lb  # print in the Rmd

# 6) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_total, "fig_weed_biomass_total_model_lb_ac.png"),
  plot     = fig_wbm_total_model_lb,
  width    = 9,
  height   = 5.5,
  dpi      = 300
)

```


#### Raw means (lb ha⁻¹)
```{r}
### Figure: Weed biomass by weed management treatment (raw means ± SE, model CLDs)
### Units converted from kg ha^-1 to lb ac^-1 for farmer meetings

# kg/ha → lb/ac conversion factor
kg_ha_to_lb_ac <- 0.892179

# 1) Raw means and SE by treatment (in lb ac^-1) -------------------------
raw_wbm_summary_lb <- weed_biomass_clean |>
  group_by(weed_trt) |>
  summarise(
    n    = n(),
    mean = mean(weed_biomass_kg_ha, na.rm = TRUE),
    sd   = sd(weed_biomass_kg_ha, na.rm = TRUE),
    se   = sd / sqrt(n),
    .groups = "drop"
  ) |>
  mutate(
    # convert to lb ac^-1
    mean = mean * kg_ha_to_lb_ac,
    sd   = sd   * kg_ha_to_lb_ac,
    se   = se   * kg_ha_to_lb_ac,
    weed_trt = factor(weed_trt, levels = mow_levels),
    ymin     = pmax(mean - se, 0),
    ymax     = mean + se
  )

# 2) Model-based CLDs for treatment main effect -------------------------
emm_wbm <- emmeans(
  weed_glmm,
  ~ weed_trt,
  type = "response"   # back-transform from link scale (still in kg/ha)
)

cld_wbm <- cld(
  emm_wbm,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE   # so "a" = highest biomass group(s)
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    .group   = str_trim(.group)
  ) |>
  select(weed_trt, .group)

# 3) Join raw means + model CLDs ----------------------------------------
plot_df_wbm_raw_lb <- raw_wbm_summary_lb |>
  left_join(cld_wbm, by = "weed_trt")

# 4) Plot ---------------------------------------------------------------
fig_wbm_total_raw_lb <- ggplot(plot_df_wbm_raw_lb, aes(x = weed_trt, y = mean, fill = weed_trt)) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 6
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x       = NULL,
    y       = expression(Weed~biomass~"(lb"~ac^{-1}*")"),
    title   = "Weed biomass by weed management",
    caption = "Bars show raw means (lb ac⁻¹) ± SE; similar letters indicate no significant difference (Fisher’s LSD test, α = 0.05)."
  ) +
  theme_classic(base_size = 18) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_wbm_total_raw_lb  # print in the Rmd

# 5) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_total, "fig_weed_biomass_total_raw_lb_ac.png"),
  plot     = fig_wbm_total_raw_lb,
  width    = 9,
  height   = 5.5,
  dpi      = 300
)

```



## In-row weed biomass
### Model-prediected means (kg ha⁻¹)
```{r}
#### In-row weed biomass (kg ha^-1): model-predicted means ± SE with model CLDs

## 0) Directory for in-row weed-biomass figures --------------------------
fig_dir_inrow <- here("analysis", "figs", "weed-biomass", "in-row")
dir.create(fig_dir_inrow, showWarnings = FALSE, recursive = TRUE)

## Figure: In-row weed biomass by weed management treatment 
## (GLMM-predicted marginal means ± SE, model CLDs)

# 1) Model-based emmeans by treatment -----------------------------------
emm_inrow <- emmeans(
  weed_inrow_glmm,
  ~ weed_trt,
  type = "response"   # back-transform from link scale (kg ha^-1)
)

# Tidy emmeans and prepare SE-based error bars
emm_inrow_df <- as_tibble(emm_inrow) |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    mean     = response,
    ymin     = pmax(mean - SE, 0),
    ymax     = mean + SE
  )

# 2) Model-based CLDs for treatment main effect -------------------------
cld_inrow <- cld(
  emm_inrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE   # so "a" = highest biomass group(s)
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    .group   = str_trim(.group)
  ) |>
  select(weed_trt, .group)

# 3) Join model-predicted means + CLDs ----------------------------------
plot_df_inrow_model <- emm_inrow_df |>
  left_join(cld_inrow, by = "weed_trt")

# 4) Plot ---------------------------------------------------------------
fig_inrow_model <- ggplot(
  plot_df_inrow_model,
  aes(x = weed_trt, y = mean, fill = weed_trt)
) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 6
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x       = NULL,
    y       = expression(In-row~weed~biomass~"(kg"~ha^{-1}*")"),
    title   = "In-row weed biomass by weed management",
    caption = "Bars show GLMM-predicted marginal means (kg"~ha^{-1}*") ± SE; letters denote Fisher’s LSD groups (α = 0.05)."
  ) +
  theme_classic(base_size = 18) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_inrow_model  # print in the Rmd

# 5) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_inrow, "fig_inrow-weed_biomass_model_kg_ha.png"),
  plot     = fig_inrow_model,
  width    = 9,
  height   = 5.5,
  dpi      = 300
)

```
### Raw means (kg ha⁻¹)
```{r}
#### In-row weed biomass (kg ha^-1): raw means ± SE with model CLDs

## 0) Directory for in-row weed-biomass figures --------------------------
fig_dir_inrow <- here("analysis", "figs", "weed-biomass", "in-row")
dir.create(fig_dir_inrow, showWarnings = FALSE, recursive = TRUE)

## Figure: In-row weed biomass by weed management treatment (raw means ± SE, model CLDs)

# 1) Raw means and SE by treatment --------------------------------------
raw_inrow_summary <- weed_biomass_clean |>
  group_by(weed_trt) |>
  summarise(
    n    = n(),
    mean = mean(inrow_weed_biomass_kg_ha, na.rm = TRUE),
    sd   = sd(inrow_weed_biomass_kg_ha, na.rm = TRUE),
    se   = sd / sqrt(n),
    .groups = "drop"
  ) |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    ymin     = pmax(mean - se, 0),
    ymax     = mean + se
  )

# 2) Model-based CLDs for treatment main effect -------------------------
emm_inrow <- emmeans(
  weed_inrow_glmm,
  ~ weed_trt,
  type = "response"   # back-transform from link scale
)

cld_inrow <- cld(
  emm_inrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE   # so "a" = highest biomass group(s)
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    .group   = str_trim(.group)
  ) |>
  select(weed_trt, .group)

# 3) Join raw means + model CLDs ----------------------------------------
plot_df_inrow_raw <- raw_inrow_summary |>
  left_join(cld_inrow, by = "weed_trt")

# 4) Plot ---------------------------------------------------------------
fig_inrow_raw <- ggplot(plot_df_inrow_raw, aes(x = weed_trt, y = mean, fill = weed_trt)) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 6
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x       = NULL,
    y       = expression(In-row~weed~biomass~"(kg"~ha^{-1}*")"),
    title   = "In-row weed biomass by weed management",
     caption = "Bars show raw means (kg ha⁻¹) ± SE; similar letters indicate no significant difference (Fisher’s LSD test, α = 0.05)."
  ) +
  theme_classic(base_size = 18) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_inrow_raw  # print in the Rmd

# 5) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_inrow, "fig_inrow-weed_biomass_raw_kg_ha.png"),
  plot     = fig_inrow_raw,
  width    = 9,
  height   = 5.5,
  dpi      = 300
)

```

### Extension figures (lb ac⁻¹)
#### Model-prediected means (lb ac⁻¹)
```{r}
### Figure: In-row weed biomass by weed management 
### (GLMM-predicted marginal means ± SE, model CLDs, lb ac^-1)

# kg/ha → lb/ac conversion factor
kg_ha_to_lb_ac <- 0.892179

# 1) Model-based emmeans by treatment (still fit in kg ha^-1) -----------
emm_inrow <- emmeans(
  weed_inrow_glmm,
  ~ weed_trt,
  type = "response"   # back-transform from link scale (kg ha^-1)
)

# Tidy emmeans and convert to lb ac^-1
emm_inrow_lb_df <- as_tibble(emm_inrow) |>
  mutate(
    weed_trt   = factor(weed_trt, levels = mow_levels),
    mean_lb    = response * kg_ha_to_lb_ac,
    SE_lb      = SE       * kg_ha_to_lb_ac,
    ymin       = pmax(mean_lb - SE_lb, 0),
    ymax       = mean_lb + SE_lb
  )

# 2) Model-based CLDs for treatment main effect -------------------------
cld_inrow <- cld(
  emm_inrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE   # so "a" = highest biomass group(s)
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    .group   = str_trim(.group)
  ) |>
  select(weed_trt, .group)

# 3) Join model-predicted means (lb/ac) + CLDs --------------------------
plot_df_inrow_model_lb <- emm_inrow_lb_df |>
  left_join(cld_inrow, by = "weed_trt")

# 4) Plot ---------------------------------------------------------------
fig_inrow_model_lb <- ggplot(
  plot_df_inrow_model_lb,
  aes(x = weed_trt, y = mean_lb, fill = weed_trt)
) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 6
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x       = NULL,
    y       = expression(In-row~weed~biomass~"(lb"~ac^{-1}*")"),
    title   = "In-row weed biomass by weed management",
    caption = "Bars show GLMM-predicted marginal means (lb"~ac^{-1}*") ± SE; letters denote Fisher’s LSD groups (α = 0.05)."
  ) +
  theme_classic(base_size = 18) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_inrow_model_lb  # print in the Rmd

# 5) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_inrow, "fig_inrow-weed_biomass_model_lb_ac.png"),
  plot     = fig_inrow_model_lb,
  width    = 9,
  height   = 5.5,
  dpi      = 300
)

```


#### Raw means (lb ha⁻¹)
```{r}
### Figure: In-row weed biomass by weed management (raw means ± SE, model CLDs)
### Units converted from kg ha^-1 to lb ac^-1 for farmer meetings

# kg/ha → lb/ac conversion factor
kg_ha_to_lb_ac <- 0.892179

# 1) Raw means and SE by treatment (in lb ac^-1) -------------------------
raw_inrow_summary_lb <- weed_biomass_clean |>
  group_by(weed_trt) |>
  summarise(
    n    = n(),
    mean = mean(inrow_weed_biomass_kg_ha, na.rm = TRUE),
    sd   = sd(inrow_weed_biomass_kg_ha, na.rm = TRUE),
    se   = sd / sqrt(n),
    .groups = "drop"
  ) |>
  mutate(
    # convert to lb ac^-1
    mean = mean * kg_ha_to_lb_ac,
    sd   = sd   * kg_ha_to_lb_ac,
    se   = se   * kg_ha_to_lb_ac,
    weed_trt = factor(weed_trt, levels = mow_levels),
    ymin     = pmax(mean - se, 0),
    ymax     = mean + se
  )

# 2) Model-based CLDs for treatment main effect (still on kg/ha scale) ---
emm_inrow <- emmeans(
  weed_inrow_glmm,
  ~ weed_trt,
  type = "response"   # back-transform from link scale
)

cld_inrow <- cld(
  emm_inrow,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE   # so "a" = highest biomass group(s)
) |>
  as_tibble() |>
  mutate(
    weed_trt = factor(weed_trt, levels = mow_levels),
    .group   = str_trim(.group)
  ) |>
  select(weed_trt, .group)

# 3) Join raw means + model CLDs ----------------------------------------
plot_df_inrow_raw_lb <- raw_inrow_summary_lb |>
  left_join(cld_inrow, by = "weed_trt")

# 4) Plot ---------------------------------------------------------------
fig_inrow_raw_lb <- ggplot(plot_df_inrow_raw_lb, aes(x = weed_trt, y = mean, fill = weed_trt)) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 6
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x       = NULL,
    y       = expression(In-row~weed~biomass~"(lb"~ac^{-1}*")"),
    title   = "In-row weed biomass by weed management",
    caption = "Bars show raw means (lb ac^-1) ± SE; similar letters indicate no significant difference (Fisher’s LSD test, P > 0.05)."
  ) +
  theme_classic(base_size = 18) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_inrow_raw_lb  # print in the Rmd

# 5) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_inrow, "fig_inrow-weed_biomass_raw_lb_ac.png"),
  plot     = fig_inrow_raw_lb,
  width    = 9,
  height   = 5.5,
  dpi      = 300
)

```
## Interrow weed biomass
### Model-prediected means (kg ha⁻¹)

```{r}
## 0) Directory for inter-row weed-biomass figures -----------------------
fig_dir_interrow <- here("analysis", "figs", "weed-biomass", "inter-row")
dir.create(fig_dir_interrow, showWarnings = FALSE, recursive = TRUE)

## Optional: nice facet labels for site_year (reuse if already defined) --
site_year_labs <- c(
  "2024.field O2 east" = "2024 \u2013 Field O2 East",
  "2024.field O2 west" = "2024 \u2013 Field O2 West",
  "2023.field v"       = "2023 \u2013 Field V"
)

## Figure: Inter-row weed biomass by weed management 
## (GLMM-predicted means ± SE, site-year CLDs)

# 1) Model-based emmeans by site-year × treatment -----------------------
emm_interrow_sy <- emmeans(
  weed_interrow_glmm,
  ~ weed_trt | site_year,
  type = "response"   # back-transform from link scale (kg ha^-1)
)

emm_interrow_sy_df <- as_tibble(emm_interrow_sy) |>
  mutate(
    site_year = as.factor(site_year),
    weed_trt  = factor(weed_trt, levels = mow_levels),
    mean      = response,
    ymin      = pmax(mean - SE, 0),
    ymax      = mean + SE
  )

# 2) Model-based CLDs for treatment *within each site-year* --------------
cld_interrow_sy <- cld(
  emm_interrow_sy,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE      # so "a" = highest biomass group(s) within each site-year
) |>
  as_tibble() |>
  mutate(
    site_year = as.factor(site_year),
    weed_trt  = factor(weed_trt, levels = mow_levels),
    .group    = str_trim(.group)
  ) |>
  select(site_year, weed_trt, .group)

# 3) Join model means + site-year CLDs ----------------------------------
plot_df_interrow_sy_model <- emm_interrow_sy_df |>
  left_join(cld_interrow_sy, by = c("site_year", "weed_trt"))

# 4) Plot ---------------------------------------------------------------
fig_interrow_model_kg_sy <- ggplot(
  plot_df_interrow_sy_model,
  aes(x = weed_trt, y = mean, fill = weed_trt)
) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 4
  ) +
  facet_wrap(
    ~ site_year,
    nrow     = 1,
    labeller = labeller(site_year = site_year_labs)
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x     = NULL,
    y     = expression(Inter-row~weed~biomass~"(kg"~ha^{-1}*")"),
    title = "Inter-row weed biomass by weed management across site-years",
    caption = "Bars show GLMM-predicted marginal means (kg"~ha^{-1}*") ± SE; letters denote Fisher’s LSD groups within each site-year (α = 0.05)."
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    strip.text   = element_text(face = "bold"),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_interrow_model_kg_sy  # print in the Rmd

# 5) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_interrow, "fig_interrow-weed_biomass_model_by-site-year_kg_ha.png"),
  plot     = fig_interrow_model_kg_sy,
  width    = 12,
  height   = 5.5,
  dpi      = 300
)

```
### Raw means (kg ha⁻¹)

```{r}
## 0) Directory for inter-row weed-biomass figures -----------------------
fig_dir_interrow <- here("analysis", "figs", "weed-biomass", "inter-row")
dir.create(fig_dir_interrow, showWarnings = FALSE, recursive = TRUE)

## Optional: nice facet labels for site_year (reuse if already defined) --
site_year_labs <- c(
  "2024.field O2 east" = "2024 \u2013 Field O2 East",
  "2024.field O2 west" = "2024 \u2013 Field O2 West",
  "2023.field v"       = "2023 \u2013 Field V"
)

## Figure: Inter-row weed biomass by weed management (raw means ± SE, site-year CLDs)

# 1) Raw means and SE by site-year × treatment ---------------------------
raw_interrow_summary <- weed_biomass_clean |>
  group_by(site_year, weed_trt) |>
  summarise(
    n    = n(),
    mean = mean(interrow_weed_biomass_kg_ha, na.rm = TRUE),
    sd   = sd(interrow_weed_biomass_kg_ha, na.rm = TRUE),
    se   = sd / sqrt(n),
    .groups = "drop"
  ) |>
  mutate(
    site_year = as.factor(site_year),
    weed_trt  = factor(weed_trt, levels = mow_levels),
    ymin      = pmax(mean - se, 0),
    ymax      = mean + se
  )

# 2) Model-based CLDs for treatment *within each site-year* --------------
emm_interrow_sy <- emmeans(
  weed_interrow_glmm,
  ~ weed_trt | site_year,
  type = "response"   # back-transform from link scale
)

cld_interrow_sy <- cld(
  emm_interrow_sy,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE      # so "a" = highest biomass group(s) within each site-year
) |>
  as_tibble() |>
  mutate(
    site_year = as.factor(site_year),
    weed_trt  = factor(weed_trt, levels = mow_levels),
    .group    = str_trim(.group)
  ) |>
  select(site_year, weed_trt, .group)

# 3) Join raw means + site-year CLDs ------------------------------------
plot_df_interrow_sy <- raw_interrow_summary |>
  left_join(cld_interrow_sy, by = c("site_year", "weed_trt"))

# 4) Plot ---------------------------------------------------------------
fig_interrow_raw_kg_sy <- ggplot(
  plot_df_interrow_sy,
  aes(x = weed_trt, y = mean, fill = weed_trt)
) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 4
  ) +
  facet_wrap(
    ~ site_year,
    nrow     = 1,
    labeller = labeller(site_year = site_year_labs)
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x     = NULL,
    y     = expression(Inter-row~weed~biomass~"(kg"~ha^{-1}*")"),
    title = "Inter-row weed biomass by weed management across site-years",
     caption = "Bars show raw means (kg"~ha^{-1}*") ± SE; letters denote model-based CLDs within each site-year (Fisher’s LSD, α = 0.05)."
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    strip.text   = element_text(face = "bold"),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_interrow_raw_kg_sy  # print in the Rmd

# 5) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_interrow, "fig_interrow-weed_biomass_raw_by-site-year_kg_ha.png"),
  plot     = fig_interrow_raw_kg_sy,
  width    = 12,
  height   = 5.5,
  dpi      = 300
)


```
### Extension figures (lb ac^-1)
#### Model-predicted means (lb ac⁻¹)
```{r}
## 0) Directory for inter-row weed-biomass figures -----------------------
fig_dir_interrow <- here("analysis", "figs", "weed-biomass", "inter-row")
dir.create(fig_dir_interrow, showWarnings = FALSE, recursive = TRUE)

## Optional: nice facet labels for site_year (reuse if already defined) --
site_year_labs <- c(
  "2024.field O2 east" = "2024 \u2013 Field O2 East",
  "2024.field O2 west" = "2024 \u2013 Field O2 West",
  "2023.field v"       = "2023 \u2013 Field V"
)

## kg/ha -> lb/ac conversion factor --------------------------------------
kg_ha_to_lb_ac <- 0.892179

## Figure: Inter-row weed biomass by weed management 
## (GLMM-predicted means ± SE, site-year CLDs; lb ac^-1)

# 1) Model-based emmeans by site-year × treatment -----------------------
emm_interrow_sy <- emmeans(
  weed_interrow_glmm,
  ~ weed_trt | site_year,
  type = "response"   # back-transform from link scale (kg ha^-1)
)

emm_interrow_sy_lb <- as_tibble(emm_interrow_sy) |>
  mutate(
    site_year = as.factor(site_year),
    weed_trt  = factor(weed_trt, levels = mow_levels),
    mean      = response * kg_ha_to_lb_ac,
    SE_lb     = SE       * kg_ha_to_lb_ac,
    ymin      = pmax(mean - SE_lb, 0),
    ymax      = mean + SE_lb
  )

# 2) Model-based CLDs for treatment *within each site-year* --------------
cld_interrow_sy <- cld(
  emm_interrow_sy,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE      # "a" = highest biomass group(s) within each site-year
) |>
  as_tibble() |>
  mutate(
    site_year = as.factor(site_year),
    weed_trt  = factor(weed_trt, levels = mow_levels),
    .group    = str_trim(.group)
  ) |>
  select(site_year, weed_trt, .group)

# 3) Join model means (lb/ac) + site-year CLDs --------------------------
plot_df_interrow_sy_lb <- emm_interrow_sy_lb |>
  left_join(cld_interrow_sy, by = c("site_year", "weed_trt"))

# 4) Plot ---------------------------------------------------------------
fig_interrow_model_lb_sy <- ggplot(
  plot_df_interrow_sy_lb,
  aes(x = weed_trt, y = mean, fill = weed_trt)
) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 4
  ) +
  facet_wrap(
    ~ site_year,
    nrow     = 1,
    labeller = labeller(site_year = site_year_labs)
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x       = NULL,
    y       = expression(Inter-row~weed~biomass~"(lb"~ac^{-1}*")"),
    title   = "Inter-row weed biomass by weed management across site-years",
    caption = "Bars show GLMM-predicted marginal means (lb"~ac^{-1}*") ± SE; letters denote Fisher’s LSD groups within each site-year (α = 0.05)."
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    strip.text   = element_text(face = "bold"),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_interrow_model_lb_sy  # print in the Rmd

# 5) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_interrow,
                       "fig_interrow-weed_biomass_model_by-site-year_lb_ac.png"),
  plot     = fig_interrow_model_lb_sy,
  width    = 12,
  height   = 5.5,
  dpi      = 300
)

```

#### Raw means (lb ha⁻¹)
```{r}
## 0) Directory for inter-row weed-biomass figures -----------------------
fig_dir_interrow <- here("analysis", "figs", "weed-biomass", "inter-row")
dir.create(fig_dir_interrow, showWarnings = FALSE, recursive = TRUE)

## Optional: nice facet labels for site_year (reuse if already defined) --
site_year_labs <- c(
  "2024.field O2 east" = "2024 \u2013 Field O2 East",
  "2024.field O2 west" = "2024 \u2013 Field O2 West",
  "2023.field v"       = "2023 \u2013 Field V"
)

## kg/ha -> lb/ac conversion factor --------------------------------------
kg_ha_to_lb_ac <- 0.892179

## Figure: Inter-row weed biomass by weed management (raw means ± SE, site-year CLDs)

# 1) Raw means and SE by site-year × treatment (now in lb ac^-1) --------
raw_interrow_summary <- weed_biomass_clean |>
  group_by(site_year, weed_trt) |>
  summarise(
    n    = n(),
    mean = mean(interrow_weed_biomass_kg_ha, na.rm = TRUE),
    sd   = sd(interrow_weed_biomass_kg_ha, na.rm = TRUE),
    se   = sd / sqrt(n),
    .groups = "drop"
  ) |>
  mutate(
    # convert to lb ac^-1
    mean = mean * kg_ha_to_lb_ac,
    sd   = sd   * kg_ha_to_lb_ac,
    se   = se   * kg_ha_to_lb_ac,
    site_year = as.factor(site_year),
    weed_trt  = factor(weed_trt, levels = mow_levels),
    ymin      = pmax(mean - se, 0),
    ymax      = mean + se
  )

# 2) Model-based CLDs for treatment *within each site-year* --------------
emm_interrow_sy <- emmeans(
  weed_interrow_glmm,
  ~ weed_trt | site_year,
  type = "response"   # back-transform from link scale (still kg/ha internally)
)

cld_interrow_sy <- cld(
  emm_interrow_sy,
  adjust   = "none",
  Letters  = letters,
  sort     = TRUE,
  reversed = TRUE      # so "a" = highest biomass group(s) within each site-year
) |>
  as_tibble() |>
  mutate(
    site_year = as.factor(site_year),
    weed_trt  = factor(weed_trt, levels = mow_levels),
    .group    = str_trim(.group)
  ) |>
  select(site_year, weed_trt, .group)

# 3) Join raw means + site-year CLDs ------------------------------------
plot_df_interrow_sy <- raw_interrow_summary |>
  left_join(cld_interrow_sy, by = c("site_year", "weed_trt"))

# 4) Plot ---------------------------------------------------------------
fig_interrow_raw_lb_sy <- ggplot(
  plot_df_interrow_sy,
  aes(x = weed_trt, y = mean, fill = weed_trt)
) +
  geom_col(width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.14) +
  geom_text(
    aes(y = ymax * 1.08, label = .group),
    vjust    = 0,
    fontface = "bold",
    size     = 4
  ) +
  facet_wrap(
    ~ site_year,
    nrow     = 1,
    labeller = labeller(site_year = site_year_labs)
  ) +
  scale_fill_manual(values = fill_cols, guide = "none") +
  scale_x_discrete(labels = label_break_comma_cult) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x     = NULL,
    y     = expression(Inter-row~weed~biomass~"(lb"~ac^{-1}*")"),
    title = "Inter-row weed biomass by weed management across site-years",
     caption = "Bars show raw means (lb"~ac^{-1}*") ± SE; letters denote model-based CLDs within each site-year (Fisher’s LSD, P > 0.05)."
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.text.x  = element_text(lineheight = 0.95, margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    strip.text   = element_text(face = "bold"),
    plot.title   = element_text(face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0)
  )

fig_interrow_raw_lb_sy  # print in the Rmd

# 5) Save figure --------------------------------------------------------
ggsave(
  filename = file.path(fig_dir_interrow, "fig_interrow-weed_biomass_raw_by-site-year_lb_ac.png"),
  plot     = fig_interrow_raw_lb_sy,
  width    = 12,
  height   = 5.5,
  dpi      = 300
)

```

