---
title: "Weed soy relationships"

output:
  github_document:
    toc: true
always_allow_html: true


---

#Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.path = "figs/analysis/weed_soy-"
)
```

#Packages
```{r}
## Packages -----------------------------------------------------------------

library(tidyverse)    # dplyr, ggplot2, readr, tibble, etc.
library(janitor)
library(readxl)
library(here)

# Mixed models / GLMMs
library(lme4)
library(glmmTMB)
library(DHARMa)
library(emmeans)
library(multcomp)
library(car)

# Tables / reporting
library(kableExtra)
library(broom)        # tidy() / glance() for lm
library(broom.mixed)  # tidy() / glance() for lmer / glmmTMB

# Diagnostics
library(performance)  # check_model(), etc.

# Nonlinear fits
library(minpack.lm)   # nlsLM(), if needed

# Palettes
library(WrensBookshelf)

# Conflict handling ---------------------------------------------------------

library(conflicted)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::recode)

## Treatment level order (used everywhere) ----------------------------------

mow_levels <- c(
  "Rolled, no control",
  "Rolled, mowing",
  "Rolled, high-residue cultivation",
  "Tilled, mowing",
  "Tilled, cultivation"
)

## Color palette for treatments (CVD-safe) ----------------------------------

fill_cols <- WB_brewer(
  name = "WhatWellBuild",
  n    = length(mow_levels),
  type = "discrete"
) |>
  setNames(mow_levels)

## Axis label helpers -------------------------------------------------------

# Break on spaces (every word on its own line)
label_break_spaces <- function(x) {
  stringr::str_replace_all(x, " ", "\n")
}

# Break after the comma: "Rolled,\nno control"
label_break_comma <- function(x) {
  stringr::str_replace_all(x, ", ", ",\n")
}

# Break after comma and split "high-residue cultivation"
# -> "Rolled,\nhigh-residue\ncultivation"
label_break_comma_cult <- function(x) {
  x |>
    stringr::str_replace("high-residue cultivation",
                         "high-residue\ncultivation") |>
    stringr::str_replace_all(", ", ",\n")
}

## Helper: tidy emmeans output regardless of CI column names ---------------

tidy_emm <- function(emm, ref_levels = NULL) {
  emm_df <- as.data.frame(emm)

  lcl_col <- intersect(c("lower.CL", "asymp.LCL"), names(emm_df))[1]
  ucl_col <- intersect(c("upper.CL", "asymp.UCL"), names(emm_df))[1]

  if (is.na(lcl_col) || is.na(ucl_col)) {
    stop("Could not find CI columns in emmeans output.")
  }

  out <- emm_df |>
    dplyr::mutate(
      ci_low  = .data[[lcl_col]],
      ci_high = .data[[ucl_col]]
    )

  if (!is.null(ref_levels) && "weed_trt" %in% names(out)) {
    out <- out |>
      dplyr::mutate(weed_trt = factor(weed_trt, levels = ref_levels))
  }

  out
}

```

# Data import & prep
```{r}
## Data import & prep ---------------------------------------------------------

# convenient conversion factor (kg/ha -> lb/ac)
kg_ha_to_lb_ac <- 0.892179   # 1 kg/ha ≈ 0.892179 lb/ac

# 1) Read + clean master (weed + soybean) -------------------------------------

weed_biomass_clean <- read_excel(
  here("data", "raw", "All Treatments", "combined_raw.xlsx")
) |>
  clean_names() |>
  rename(weed_trt = treatment) |>
  mutate(
    year      = factor(year),
    location  = factor(location),
    site_year = factor(interaction(year, location, drop = TRUE)),
    block     = factor(block),
    weed_trt  = recode(
      weed_trt,
      "RNO" = "Rolled, no control",
      "RIM" = "Rolled, mowing",
      "RIC" = "Rolled, high-residue cultivation",
      "TIM" = "Tilled, mowing",
      "TIC" = "Tilled, cultivation"
    ),
    weed_trt = factor(weed_trt, levels = mow_levels)
  ) |>
  # keep only rows with non-missing total weed biomass
  filter(!is.na(weed_biomass)) |>
  # per-area metrics (biomass in g per 0.5 m² quadrat)
  mutate(
    ## ---- WEEDS: total / in-row / interrow ----------------------------------
    weed_biomass_g_m2           = weed_biomass          * 2,   # g/m²
    weed_biomass_kg_ha          = weed_biomass          * 20,  # kg/ha
    weed_biomass_lb_ac          = weed_biomass_kg_ha    * kg_ha_to_lb_ac,

    inrow_weed_biomass_g_m2     = inrow_weed_biomass    * 2,
    inrow_weed_biomass_kg_ha    = inrow_weed_biomass    * 20,
    inrow_weed_biomass_lb_ac    = inrow_weed_biomass_kg_ha * kg_ha_to_lb_ac,

    interrow_weed_biomass_g_m2  = interrow_weed_biomass * 2,
    interrow_weed_biomass_kg_ha = interrow_weed_biomass * 20,
    interrow_weed_biomass_lb_ac = interrow_weed_biomass_kg_ha * kg_ha_to_lb_ac,

    ## ---- SOYBEAN: biomass --------------------------------------------------
    # bean_biomass assumed g per 0.5 m² quadrat (same as weeds)
    bean_biomass_g_m2    = bean_biomass * 2,           # g/m²
    bean_biomass_kg_ha   = bean_biomass * 20,          # kg/ha
    bean_biomass_lb_ac   = bean_biomass_kg_ha * kg_ha_to_lb_ac,

    ## ---- SOYBEAN: adjusted yield (same formulas as before) -----------------
    bean_yield_adj_bu_acre = (((bean_yield / 454) / (16.4 / 43560)) / 60) *
      ((100 - 0.00001) / (100 - 13)),
    bean_yield_adj_lbs_acre = ((bean_yield / 454) / (16.4 / 43560)) *
      ((100 - 0.00001) / (100 - 13)),
    bean_yield_adj_kg_ha = ((bean_yield / 454) / (16.4 / 43560)) * 1.12085 *
      ((100 - 0.00001) / (100 - 13)),

    ## ---- SOYBEAN: population metrics --------------------------------------
    # bean_population assumed plants per 1 m of row
    bean_population_two_meter = bean_population * 2,
    bean_population_hectare   = (bean_population / 0.762) * 10000,
    bean_population_acre      = bean_population_hectare / 2.471
  )

# 2) 2023 Field V subset ------------------------------------------------------

weed_biomass_field_v_2023 <- weed_biomass_clean |>
  filter(
    year == "2023",          # year is a factor; compare to character label
    location == "field v"
  )

# Quick check tables ----------------------------------------------------------

kable(
  head(weed_biomass_clean),
  caption = "All site-years, cleaned (weed + soybean metrics)"
)

kable(
  head(weed_biomass_field_v_2023),
  caption = "Field V only, 2023 (weed + soybean metrics)"
)

```
# Model testing
## Total weed biomass
### Yield 
#### Dataset

```{r}
# ----------------------------------------------------------------------
# Model testing: total weed biomass vs adjusted soybean yield (kg ha^-1)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_yield <- weed_biomass_clean |>
  filter(
    !is.na(weed_biomass_kg_ha),
    !is.na(bean_yield_adj_kg_ha),
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_yield[, c("site_year", "block",
                  "weed_biomass_kg_ha", "bean_yield_adj_kg_ha")])

```
#### Mixed model selection (site-year effects)
```{r}
## Mixed model selection: additive vs interaction ----------------------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-year (site_year as fixed effect)
yield_add <- lmer(
  bean_yield_adj_kg_ha ~ weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_yield
)

# Interaction: slope allowed to differ by site-year
yield_int <- lmer(
  bean_yield_adj_kg_ha ~ weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_yield
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_yield   <- anova(yield_add, yield_int)
p_int_yield <- lrt_yield$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_yield <- tibble(
  model = c(
    "Additive: weed_biomass_kg_ha + site_year",
    "Interaction: weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(yield_add), AIC(yield_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_yield)
  )

# Selection rule: prefer simpler model unless interaction is clearly better
keep_interaction <- (p_int_yield < 0.05) &&
  (aic_yield$deltaAIC[aic_yield$model ==
                        "Interaction: weed_biomass_kg_ha * site_year"] < 2)

yield_mod_best <- if (keep_interaction) yield_int else yield_add

chosen_model_name_yield <- if (keep_interaction) {
  "Interaction: weed_biomass_kg_ha * site_year"
} else {
  "Additive: weed_biomass_kg_ha + site_year"
}

aic_yield_out <- aic_yield |>
  mutate(
    Selected = if_else(model == chosen_model_name_yield, "Yes", "")
  )

aic_yield_out |>
  kable(
    digits  = 2,
    caption = "Adjusted soybean yield vs total weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for adjusted soybean yield vs total weed biomass:\n  ",
  chosen_model_name_yield,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_yield)
)

```
#### Slope summary via emtrends
```{r}
## Slope summary: emtrends for yield vs total weed biomass --------------

# helper: find the trend column name from emtrends output
get_trend_col <- function(df, var = "weed_biomass_kg_ha") {
  nm <- grep(paste0("^", var, "\\.trend$"), names(df), value = TRUE)
  if (length(nm) != 1) {
    stop("Could not uniquely identify trend column for ", var)
  }
  nm
}

# helper: find t or z statistic column
get_stat_col <- function(df) {
  nm <- intersect(c("t.ratio", "z.ratio"), names(df))
  if (length(nm) == 0) {
    return(NA_character_)
  }
  nm[1]
}

if (chosen_model_name_yield == "Additive: weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope <- emtrends(yield_mod_best, ~ 1, var = "weed_biomass_kg_ha")
  emm_slope_df <- summary(emm_slope, infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col <- get_trend_col(emm_slope_df, "weed_biomass_kg_ha")
  stat_col  <- get_stat_col(emm_slope_df)
  
  has_df <- "df" %in% names(emm_slope_df)
  has_p  <- "p.value" %in% names(emm_slope_df)
  
  slope_tbl <- emm_slope_df |>
    transmute(
      Scope              = "Pooled across site-years",
      slope_kg_per_kg    = .data[[trend_col]],
      slope_kg_per_100kg = .data[[trend_col]] * 100,
      SE,
      df         = if (has_df) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col)) .data[[stat_col]] else NA_real_,
      p_value    = if (has_p) .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope <- emtrends(yield_mod_best, ~ site_year, var = "weed_biomass_kg_ha")
  emm_slope_df <- summary(emm_slope, infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col <- get_trend_col(emm_slope_df, "weed_biomass_kg_ha")
  stat_col  <- get_stat_col(emm_slope_df)
  
  has_df <- "df" %in% names(emm_slope_df)
  has_p  <- "p.value" %in% names(emm_slope_df)
  
  slope_tbl <- emm_slope_df |>
    transmute(
      Scope              = as.character(site_year),
      slope_kg_per_kg    = .data[[trend_col]],
      slope_kg_per_100kg = .data[[trend_col]] * 100,
      SE,
      df         = if (has_df) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col)) .data[[stat_col]] else NA_real_,
      p_value    = if (has_p) .data[["p.value"]] else NA_real_
    )
}

slope_tbl |>
  mutate(
    across(
      c(slope_kg_per_kg, slope_kg_per_100kg, SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of adjusted soybean yield vs total weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (kg yield per kg weed)",
      "Slope (kg yield per 100 kg weed)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```
#### Linear vs rectangular hyperbolic response
```{r}
## Linear vs rectangular hyperbolic response of yield to weed biomass ----

# Use the same plot-level data (no random effects here, just a marginal comparison)
dat_yield <- reg_yield |>
  select(weed_biomass_kg_ha, bean_yield_adj_kg_ha) |>
  rename(
    B = weed_biomass_kg_ha,
    Y = bean_yield_adj_kg_ha
  )

# Quick sanity check (optional)
names(dat_yield)    # should be "B" and "Y"

# 1) Linear model -------------------------------------------------------

lin_mod <- lm(Y ~ B, data = dat_yield)

# 2) Rectangular hyperbolic model:
#    Y = Y0 - (Lmax * B) / (K + B)
#    Y0   = weed-free yield
#    Lmax = maximum yield loss
#    K    = biomass at half of Lmax

start_vals <- list(
  Y0   = max(dat_yield$Y, na.rm = TRUE),  # near weed-free yield
  Lmax = max(dat_yield$Y, na.rm = TRUE) - min(dat_yield$Y, na.rm = TRUE),
  K    = median(dat_yield$B[dat_yield$B > 0], na.rm = TRUE)
)

hyper_mod <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_yield,
  start = start_vals
)

summary(hyper_mod)

# 3) AIC and R2 comparison -----------------------------------------------

# AIC
aic_lin   <- AIC(lin_mod)
aic_hyper <- AIC(hyper_mod)

# R2 for linear model
gl_lin <- broom::glance(lin_mod)
R2_lin     <- gl_lin$r.squared
R2_lin_adj <- gl_lin$adj.r.squared

# R2 for nls (hyperbolic)
get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res  # reconstruct observed response
  
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper <- get_r2_nls(hyper_mod)

# Residual df (more meaningful than n)
df_lin   <- df.residual(lin_mod)
df_hyper <- df.residual(hyper_mod)

mod_comp <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin, df_hyper),
  AIC   = c(aic_lin, aic_hyper),
  R2    = c(R2_lin, R2_hyper)
)

mod_comp |>
  kable(
    digits  = 3,
    caption = "Model comparison for adjusted soybean yield vs total weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```
#### Diagnostics
```{r}
## Diagnostics & visualization ------------------------------------------

library(performance)

# Mixed model diagnostics (chosen model)
check_model(yield_mod_best)

# Simple linear model diagnostics
check_model(lin_mod)

# Optional specific checks for the linear model
check_heteroscedasticity(lin_mod)
check_normality(lin_mod)
check_outliers(lin_mod)

```

#### Visualisation

```{r}
# Reuse treatment palette for line/point colors
line_cols <- unname(fill_cols)[1:4]

```

```{r}
ggplot(dat_yield, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: uses palette color and appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: always black, not tied to color scale
  stat_function(
    fun = function(x) coef(lin_mod)[1] + coef(lin_mod)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: uses palette color and appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("dashed", "dotdash"),
        shape    = c(NA, NA),
        alpha    = c(1, 1)
      )
    )
  ) +
  labs(
    x = bquote("Total weed biomass (kg ha"^-1*")"),
    y = bquote("Adjusted soybean yield (kg ha"^-1*")"),
    title = "Yield vs total weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )


```
### Biomass
#### Dataset

```{r}
# ----------------------------------------------------------------------
# Model testing: total weed biomass vs soybean biomass (kg ha^-1)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_biomass <- weed_biomass_clean |>
  filter(
    !is.na(weed_biomass_kg_ha),
    !is.na(bean_biomass_kg_ha),
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_biomass[, c("site_year", "block",
                    "weed_biomass_kg_ha", "bean_biomass_kg_ha")])

```
#### Mixed model selection (site-year effects)
```{r}
### Mixed model selection (site-year effects): soybean biomass ---------
## Mixed model selection: additive vs interaction ----------------------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-year (site_year as fixed effect)
biomass_add <- lmer(
  bean_biomass_kg_ha ~ weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_biomass
)

# Interaction: slope allowed to differ by site-year
biomass_int <- lmer(
  bean_biomass_kg_ha ~ weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_biomass
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_biomass   <- anova(biomass_add, biomass_int)
p_int_biomass <- lrt_biomass$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_biomass <- tibble(
  model = c(
    "Additive: weed_biomass_kg_ha + site_year",
    "Interaction: weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(biomass_add), AIC(biomass_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_biomass)
  )

# Selection rule: prefer simpler model unless interaction is clearly better
keep_interaction_biomass <- (p_int_biomass < 0.05) &&
  (aic_biomass$deltaAIC[aic_biomass$model ==
                          "Interaction: weed_biomass_kg_ha * site_year"] < 2)

biomass_mod_best <- if (keep_interaction_biomass) biomass_int else biomass_add

chosen_model_name_biomass <- if (keep_interaction_biomass) {
  "Interaction: weed_biomass_kg_ha * site_year"
} else {
  "Additive: weed_biomass_kg_ha + site_year"
}

aic_biomass_out <- aic_biomass |>
  mutate(
    Selected = if_else(model == chosen_model_name_biomass, "Yes", "")
  )

aic_biomass_out |>
  kable(
    digits  = 2,
    caption = "Soybean biomass vs total weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for soybean biomass vs total weed biomass:\n  ",
  chosen_model_name_biomass,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_biomass)
)

```
#### Slope summary via emtrends
```{r}
### Slope summary: soybean biomass vs total weed biomass ---------------

if (chosen_model_name_biomass == "Additive: weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope_biomass <- emtrends(
    biomass_mod_best, ~ 1,
    var = "weed_biomass_kg_ha"
  )
  
  emm_slope_biomass_df <- summary(emm_slope_biomass, infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_biomass <- get_trend_col(emm_slope_biomass_df, "weed_biomass_kg_ha")
  stat_col_biomass  <- get_stat_col(emm_slope_biomass_df)
  
  has_df_biomass <- "df" %in% names(emm_slope_biomass_df)
  has_p_biomass  <- "p.value" %in% names(emm_slope_biomass_df)
  
  slope_tbl_biomass <- emm_slope_biomass_df |>
    transmute(
      Scope                  = "Pooled across site-years",
      slope_kgbean_per_kgweed    = .data[[trend_col_biomass]],
      slope_kgbean_per_100kgweed = .data[[trend_col_biomass]] * 100,
      SE,
      df         = if (has_df_biomass) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_biomass)) .data[[stat_col_biomass]] else NA_real_,
      p_value    = if (has_p_biomass) .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope_biomass <- emtrends(
    biomass_mod_best, ~ site_year,
    var = "weed_biomass_kg_ha"
  )
  
  emm_slope_biomass_df <- summary(emm_slope_biomass, infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_biomass <- get_trend_col(emm_slope_biomass_df, "weed_biomass_kg_ha")
  stat_col_biomass  <- get_stat_col(emm_slope_biomass_df)
  
  has_df_biomass <- "df" %in% names(emm_slope_biomass_df)
  has_p_biomass  <- "p.value" %in% names(emm_slope_biomass_df)
  
  slope_tbl_biomass <- emm_slope_biomass_df |>
    transmute(
      Scope                  = as.character(site_year),
      slope_kgbean_per_kgweed    = .data[[trend_col_biomass]],
      slope_kgbean_per_100kgweed = .data[[trend_col_biomass]] * 100,
      SE,
      df         = if (has_df_biomass) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_biomass)) .data[[stat_col_biomass]] else NA_real_,
      p_value    = if (has_p_biomass) .data[["p.value"]] else NA_real_
    )
}

slope_tbl_biomass |>
  mutate(
    across(
      c(slope_kgbean_per_kgweed,
        slope_kgbean_per_100kgweed,
        SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of soybean biomass vs total weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (kg bean per kg weed)",
      "Slope (kg bean per 100 kg weed)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```
#### Linear vs rectangular hyperbolic response
```{r}

## Linear vs rectangular hyperbolic response of soybean biomass to weed biomass ----

# Use the same plot-level data (no random effects here, just a marginal comparison)
dat_biomass <- reg_biomass |>
  select(weed_biomass_kg_ha, bean_biomass_kg_ha) |>
  rename(
    B = weed_biomass_kg_ha,
    Y = bean_biomass_kg_ha
  )

# Quick sanity check (optional)
names(dat_biomass)    # should be "B" and "Y"

# 1) Linear model -------------------------------------------------------

lin_mod_biomass <- lm(Y ~ B, data = dat_biomass)

# 2) Rectangular hyperbolic model:
#    Y = Y0 - (Lmax * B) / (K + B)
#    Y0   = weed-free soybean biomass
#    Lmax = maximum biomass loss
#    K    = biomass at half of Lmax

start_vals_biomass <- list(
  Y0   = max(dat_biomass$Y, na.rm = TRUE),  # near weed-free biomass
  Lmax = max(dat_biomass$Y, na.rm = TRUE) - min(dat_biomass$Y, na.rm = TRUE),
  K    = median(dat_biomass$B[dat_biomass$B > 0], na.rm = TRUE)
)

hyper_mod_biomass <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_biomass,
  start = start_vals_biomass
)

summary(hyper_mod_biomass)

# 3) AIC and R2 comparison -----------------------------------------------

# AIC
aic_lin_biomass   <- AIC(lin_mod_biomass)
aic_hyper_biomass <- AIC(hyper_mod_biomass)

# R2 for linear model
gl_lin_biomass <- broom::glance(lin_mod_biomass)
R2_lin_biomass     <- gl_lin_biomass$r.squared
R2_lin_biomass_adj <- gl_lin_biomass$adj.r.squared

# R2 for nls (hyperbolic)
get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res  # reconstruct observed response
  
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper_biomass <- get_r2_nls(hyper_mod_biomass)

# Residual df (more meaningful than n)
df_lin_biomass   <- df.residual(lin_mod_biomass)
df_hyper_biomass <- df.residual(hyper_mod_biomass)

mod_comp_biomass <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin_biomass, df_hyper_biomass),
  AIC   = c(aic_lin_biomass, aic_hyper_biomass),
  R2    = c(R2_lin_biomass, R2_hyper_biomass)
)

mod_comp_biomass |>
  kable(
    digits  = 3,
    caption = "Model comparison for soybean biomass vs total weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```
#### Diagnostics
```{r}
# Mixed model diagnostics (chosen biomass model)
check_model(biomass_mod_best)

# Simple linear model diagnostics (biomass)
check_model(lin_mod_biomass)

# Optional specific checks for the biomass linear model
check_heteroscedasticity(lin_mod_biomass)
check_normality(lin_mod_biomass)
check_outliers(lin_mod_biomass)
```
#### Visualisation
```{r}
ggplot(dat_biomass, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: uses palette color and appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: always black, not tied to color scale
  stat_function(
    fun = function(x) coef(lin_mod_biomass)[1] + coef(lin_mod_biomass)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: uses palette color and appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod_biomass)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("dashed", "dotdash"),
        shape    = c(NA, NA),
        alpha    = c(1, 1)
      )
    )
  ) +
  labs(
    x = bquote("Total weed biomass (kg ha"^-1*")"),
    y = bquote("Soybean biomass (kg ha"^-1*")"),
    title = "Soybean biomass vs total weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )

```
### Population
#### Dataset
```{r}
# ----------------------------------------------------------------------
# Model testing: total weed biomass vs bean population (plants ha^-1)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_pop <- weed_biomass_clean |>
  filter(
    !is.na(weed_biomass_kg_ha),
    !is.na(bean_population_hectare),
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_pop[, c("site_year", "block",
                "weed_biomass_kg_ha", "bean_population_hectare")])

```

#### Mixed model selection (site-year effects)
```{r}
### Mixed model selection (site-year effects): bean population ----------
## Mixed model selection: additive vs interaction ----------------------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-year (site_year as fixed effect)
pop_add <- lmer(
  bean_population_hectare ~ weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_pop
)

# Interaction: slope allowed to differ by site-year
pop_int <- lmer(
  bean_population_hectare ~ weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_pop
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_pop   <- anova(pop_add, pop_int)
p_int_pop <- lrt_pop$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_pop <- tibble(
  model = c(
    "Additive: weed_biomass_kg_ha + site_year",
    "Interaction: weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(pop_add), AIC(pop_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_pop)
  )

# Selection rule: prefer simpler model unless interaction is clearly better
keep_interaction_pop <- (p_int_pop < 0.05) &&
  (aic_pop$deltaAIC[aic_pop$model ==
                      "Interaction: weed_biomass_kg_ha * site_year"] < 2)

pop_mod_best <- if (keep_interaction_pop) pop_int else pop_add

chosen_model_name_pop <- if (keep_interaction_pop) {
  "Interaction: weed_biomass_kg_ha * site_year"
} else {
  "Additive: weed_biomass_kg_ha + site_year"
}

aic_pop_out <- aic_pop |>
  mutate(
    Selected = if_else(model == chosen_model_name_pop, "Yes", "")
  )

aic_pop_out |>
  kable(
    digits  = 2,
    caption = "Bean population vs total weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for bean population vs total weed biomass:\n  ",
  chosen_model_name_pop,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_pop)
)

```

#### Slope summary via emtrends
```{r}
### Slope summary: bean population vs total weed biomass ---------------

if (chosen_model_name_pop == "Additive: weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope_pop <- emtrends(
    pop_mod_best, ~ 1,
    var = "weed_biomass_kg_ha"
  )
  
  emm_slope_pop_df <- summary(emm_slope_pop, infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_pop <- get_trend_col(emm_slope_pop_df, "weed_biomass_kg_ha")
  stat_col_pop  <- get_stat_col(emm_slope_pop_df)
  
  has_df_pop <- "df" %in% names(emm_slope_pop_df)
  has_p_pop  <- "p.value" %in% names(emm_slope_pop_df)
  
  slope_tbl_pop <- emm_slope_pop_df |>
    transmute(
      Scope                           = "Pooled across site-years",
      slope_plants_per_kgweed         = .data[[trend_col_pop]],
      slope_plants_per_100kgweed      = .data[[trend_col_pop]] * 100,
      SE,
      df         = if (has_df_pop) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_pop)) .data[[stat_col_pop]] else NA_real_,
      p_value    = if (has_p_pop) .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope_pop <- emtrends(
    pop_mod_best, ~ site_year,
    var = "weed_biomass_kg_ha"
  )
  
  emm_slope_pop_df <- summary(emm_slope_pop, infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_pop <- get_trend_col(emm_slope_pop_df, "weed_biomass_kg_ha")
  stat_col_pop  <- get_stat_col(emm_slope_pop_df)
  
  has_df_pop <- "df" %in% names(emm_slope_pop_df)
  has_p_pop  <- "p.value" %in% names(emm_slope_pop_df)
  
  slope_tbl_pop <- emm_slope_pop_df |>
    transmute(
      Scope                           = as.character(site_year),
      slope_plants_per_kgweed         = .data[[trend_col_pop]],
      slope_plants_per_100kgweed      = .data[[trend_col_pop]] * 100,
      SE,
      df         = if (has_df_pop) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_pop)) .data[[stat_col_pop]] else NA_real_,
      p_value    = if (has_p_pop) .data[["p.value"]] else NA_real_
    )
}

slope_tbl_pop |>
  mutate(
    across(
      c(slope_plants_per_kgweed,
        slope_plants_per_100kgweed,
        SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of bean population vs total weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (plants ha^-1 per kg weed ha^-1)",
      "Slope (plants ha^-1 per 100 kg weed ha^-1)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

####Linear vs rectangular hyperbolic response
```{r}
## Linear vs rectangular hyperbolic response of bean population to weed biomass ----

# Use the same plot-level data (no random effects here, just a marginal comparison)
dat_pop <- reg_pop |>
  select(weed_biomass_kg_ha, bean_population_hectare) |>
  rename(
    B = weed_biomass_kg_ha,
    Y = bean_population_hectare
  )

# Quick sanity check (optional)
names(dat_pop)    # should be "B" and "Y"

# 1) Linear model -------------------------------------------------------

lin_mod_pop <- lm(Y ~ B, data = dat_pop)

# 2) Rectangular hyperbolic model:
#    Y = Y0 - (Lmax * B) / (K + B)
#    Y0   = weed-free bean population (plants ha^-1)
#    Lmax = maximum population loss
#    K    = weed biomass at half of Lmax

start_vals_pop <- list(
  Y0   = max(dat_pop$Y, na.rm = TRUE),  # near weed-free population
  Lmax = max(dat_pop$Y, na.rm = TRUE) - min(dat_pop$Y, na.rm = TRUE),
  K    = median(dat_pop$B[dat_pop$B > 0], na.rm = TRUE)
)

hyper_mod_pop <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_pop,
  start = start_vals_pop
)

summary(hyper_mod_pop)

# 3) AIC and R2 comparison -----------------------------------------------

# AIC
aic_lin_pop   <- AIC(lin_mod_pop)
aic_hyper_pop <- AIC(hyper_mod_pop)

# R2 for linear model
gl_lin_pop <- broom::glance(lin_mod_pop)
R2_lin_pop     <- gl_lin_pop$r.squared
R2_lin_pop_adj <- gl_lin_pop$adj.r.squared

# R2 for nls (hyperbolic)
get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res  # reconstruct observed response
  
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper_pop <- get_r2_nls(hyper_mod_pop)

# Residual df (more meaningful than n)
df_lin_pop   <- df.residual(lin_mod_pop)
df_hyper_pop <- df.residual(hyper_mod_pop)

mod_comp_pop <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin_pop, df_hyper_pop),
  AIC   = c(aic_lin_pop, aic_hyper_pop),
  R2    = c(R2_lin_pop, R2_hyper_pop)
)

mod_comp_pop |>
  kable(
    digits  = 3,
    caption = "Model comparison for bean population vs total weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Diagnostics
```{r}
# --- Bean population vs weed biomass ---------------------------------

# Mixed model diagnostics (chosen population model)
check_model(pop_mod_best)

# Simple linear model diagnostics (population)
check_model(lin_mod_pop)

# Optional specific checks for the population linear model
check_heteroscedasticity(lin_mod_pop)
check_normality(lin_mod_pop)
check_outliers(lin_mod_pop)
```

#### Visualisation
```{r}
ggplot(dat_pop, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: uses palette color and appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: always black, not tied to color scale
  stat_function(
    fun = function(x) coef(lin_mod_pop)[1] + coef(lin_mod_pop)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: uses palette color and appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod_pop)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  # <-- new: y-axis as whole numbers, no scientific notation
  scale_y_continuous(
    labels = scales::label_comma(accuracy = 1)
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("dashed", "dotdash"),
        shape    = c(NA, NA),
        alpha    = c(1, 1)
      )
    )
  ) +
  labs(
    x = bquote("Total weed biomass (kg ha"^-1*")"),
    y = bquote("Bean population (plants ha"^-1*")"),
    title = "Bean population vs total weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )


```

### Seed weight
#### Dataset
```{r}
# ----------------------------------------------------------------------
# Model testing: total weed biomass vs 100-seed weight (g)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_100sw <- weed_biomass_clean |>
  filter(
    !is.na(weed_biomass_kg_ha),
    !is.na(seed_weight),   # <-- replace with your actual 100-seed weight column
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_100sw[, c("site_year", "block",
                  "weed_biomass_kg_ha", "seed_weight")])

```

#### Mixed model selection (site-year effects)
```{r}
### Mixed model selection (site-year effects): 100-seed weight ----------
## Mixed model selection: additive vs interaction ----------------------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-year (site_year as fixed effect)
sw100_add <- lmer(
  seed_weight ~ weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_100sw
)

# Interaction: slope allowed to differ by site-year
sw100_int <- lmer(
  seed_weight ~ weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_100sw
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_sw100   <- anova(sw100_add, sw100_int)
p_int_sw100 <- lrt_sw100$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_sw100 <- tibble(
  model = c(
    "Additive: weed_biomass_kg_ha + site_year",
    "Interaction: weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(sw100_add), AIC(sw100_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_sw100)
  )

# Selection rule: prefer simpler model unless interaction is clearly better
keep_interaction_sw100 <- (p_int_sw100 < 0.05) &&
  (aic_sw100$deltaAIC[aic_sw100$model ==
                        "Interaction: weed_biomass_kg_ha * site_year"] < 2)

sw100_mod_best <- if (keep_interaction_sw100) sw100_int else sw100_add

chosen_model_name_sw100 <- if (keep_interaction_sw100) {
  "Interaction: weed_biomass_kg_ha * site_year"
} else {
  "Additive: weed_biomass_kg_ha + site_year"
}

aic_sw100_out <- aic_sw100 |>
  mutate(
    Selected = if_else(model == chosen_model_name_sw100, "Yes", "")
  )

aic_sw100_out |>
  kable(
    digits  = 2,
    caption = "100-seed weight vs total weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for 100-seed weight vs total weed biomass:\n  ",
  chosen_model_name_sw100,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_sw100)
)

```

#### Slope summary via emtrends
```{r}
### Slope summary: 100-seed weight vs total weed biomass ---------------

if (chosen_model_name_sw100 == "Additive: weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope_sw100 <- emtrends(
    sw100_mod_best, ~ 1,
    var = "weed_biomass_kg_ha"
  )
  
  emm_slope_sw100_df <- summary(emm_slope_sw100, infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_sw100 <- get_trend_col(emm_slope_sw100_df, "weed_biomass_kg_ha")
  stat_col_sw100  <- get_stat_col(emm_slope_sw100_df)
  
  has_df_sw100 <- "df" %in% names(emm_slope_sw100_df)
  has_p_sw100  <- "p.value" %in% names(emm_slope_sw100_df)
  
  slope_tbl_sw100 <- emm_slope_sw100_df |>
    transmute(
      Scope                              = "Pooled across site-years",
      slope_g_per_kgweed                 = .data[[trend_col_sw100]],
      slope_g_per_100kgweed              = .data[[trend_col_sw100]] * 100,
      SE,
      df         = if (has_df_sw100) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_sw100)) .data[[stat_col_sw100]] else NA_real_,
      p_value    = if (has_p_sw100) .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope_sw100 <- emtrends(
    sw100_mod_best, ~ site_year,
    var = "weed_biomass_kg_ha"
  )
  
  emm_slope_sw100_df <- summary(emm_slope_sw100, infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_sw100 <- get_trend_col(emm_slope_sw100_df, "weed_biomass_kg_ha")
  stat_col_sw100  <- get_stat_col(emm_slope_sw100_df)
  
  has_df_sw100 <- "df" %in% names(emm_slope_sw100_df)
  has_p_sw100  <- "p.value" %in% names(emm_slope_sw100_df)
  
  slope_tbl_sw100 <- emm_slope_sw100_df |>
    transmute(
      Scope                              = as.character(site_year),
      slope_g_per_kgweed                 = .data[[trend_col_sw100]],
      slope_g_per_100kgweed              = .data[[trend_col_sw100]] * 100,
      SE,
      df         = if (has_df_sw100) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_sw100)) .data[[stat_col_sw100]] else NA_real_,
      p_value    = if (has_p_sw100) .data[["p.value"]] else NA_real_
    )
}

slope_tbl_sw100 |>
  mutate(
    across(
      c(slope_g_per_kgweed,
        slope_g_per_100kgweed,
        SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of 100-seed weight vs total weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (g per kg weed ha^-1)",
      "Slope (g per 100 kg weed ha^-1)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Linear vs rectangular hyperbolic response
```{r}
## Linear vs rectangular hyperbolic response of 100-seed weight to weed biomass ----

# Use the same plot-level data (no random effects here, just a marginal comparison)
dat_sw100 <- reg_100sw |>
  select(weed_biomass_kg_ha, seed_weight) |>
  rename(
    B = weed_biomass_kg_ha,
    Y = seed_weight
  )

# Quick sanity check (optional)
names(dat_sw100)    # should be "B" and "Y"

# 1) Linear model -------------------------------------------------------

lin_mod_sw100 <- lm(Y ~ B, data = dat_sw100)

# 2) Rectangular hyperbolic model:
#    Y = Y0 - (Lmax * B) / (K + B)
#    Y0   = weed-free 100-seed weight (g)
#    Lmax = maximum loss in 100-seed weight
#    K    = weed biomass at half of Lmax

start_vals_sw100 <- list(
  Y0   = max(dat_sw100$Y, na.rm = TRUE),  # near weed-free 100-seed weight
  Lmax = max(dat_sw100$Y, na.rm = TRUE) - min(dat_sw100$Y, na.rm = TRUE),
  K    = median(dat_sw100$B[dat_sw100$B > 0], na.rm = TRUE)
)

hyper_mod_sw100 <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_sw100,
  start = start_vals_sw100
)

summary(hyper_mod_sw100)

# 3) AIC and R2 comparison -----------------------------------------------

# AIC
aic_lin_sw100   <- AIC(lin_mod_sw100)
aic_hyper_sw100 <- AIC(hyper_mod_sw100)

# R2 for linear model
gl_lin_sw100 <- broom::glance(lin_mod_sw100)
R2_lin_sw100     <- gl_lin_sw100$r.squared
R2_lin_sw100_adj <- gl_lin_sw100$adj.r.squared

# R2 for nls (hyperbolic)
get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res  # reconstruct observed response
  
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper_sw100 <- get_r2_nls(hyper_mod_sw100)

# Residual df (more meaningful than n)
df_lin_sw100   <- df.residual(lin_mod_sw100)
df_hyper_sw100 <- df.residual(hyper_mod_sw100)

mod_comp_sw100 <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin_sw100, df_hyper_sw100),
  AIC   = c(aic_lin_sw100, aic_hyper_sw100),
  R2    = c(R2_lin_sw100, R2_hyper_sw100)
)

mod_comp_sw100 |>
  kable(
    digits  = 3,
    caption = "Model comparison for 100-seed weight vs total weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Diagnostics
```{r}
# --- 100-seed weight vs weed biomass ---------------------------------

# Mixed model diagnostics (chosen 100-seed weight model)
check_model(sw100_mod_best)

# Simple linear model diagnostics (100-seed weight)
check_model(lin_mod_sw100)

# Optional specific checks for the 100-seed weight linear model
check_heteroscedasticity(lin_mod_sw100)
check_normality(lin_mod_sw100)
check_outliers(lin_mod_sw100)

```

#### Visualisation
```{r}
ggplot(dat_sw100, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: uses palette color and appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: always black, not tied to color scale
  stat_function(
    fun = function(x) coef(lin_mod_sw100)[1] + coef(lin_mod_sw100)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: uses palette color and appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod_sw100)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("dashed", "dotdash"),
        shape    = c(NA, NA),
        alpha    = c(1, 1)
      )
    )
  ) +
  labs(
    x = bquote("Total weed biomass (kg ha"^-1*")"),
    y = "100-seed weight (g)",
    title = "100-seed weight vs total weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )

```

## In-row weed biomass
### Yield
#### Dataset
```{r}
# ----------------------------------------------------------------------
# Model testing: in-row weed biomass vs adjusted soybean yield (kg ha^-1)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_yield_inrow <- weed_biomass_clean |>
  filter(
    !is.na(inrow_weed_biomass_kg_ha),
    !is.na(bean_yield_adj_kg_ha),
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_yield_inrow[, c("site_year", "block",
                        "inrow_weed_biomass_kg_ha",
                        "bean_yield_adj_kg_ha")])

```

#### Mixed model selection (site-year effects)
```{r}
## Mixed model selection: in-row weed biomass vs adjusted yield ---------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-year (site_year as fixed effect)
yield_inrow_add <- lmer(
  bean_yield_adj_kg_ha ~ inrow_weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_yield_inrow
)

# Interaction: slope allowed to differ by site-year
yield_inrow_int <- lmer(
  bean_yield_adj_kg_ha ~ inrow_weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_yield_inrow
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_yield_inrow   <- anova(yield_inrow_add, yield_inrow_int)
p_int_yield_inrow <- lrt_yield_inrow$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_yield_inrow <- tibble(
  model = c(
    "Additive: inrow_weed_biomass_kg_ha + site_year",
    "Interaction: inrow_weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(yield_inrow_add), AIC(yield_inrow_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_yield_inrow)
  )

# Selection rule: prefer simpler model unless interaction is clearly better
keep_interaction_inrow <- (p_int_yield_inrow < 0.05) &&
  (aic_yield_inrow$deltaAIC[
    aic_yield_inrow$model ==
      "Interaction: inrow_weed_biomass_kg_ha * site_year"
  ] < 2)

yield_inrow_mod_best <- if (keep_interaction_inrow) yield_inrow_int else yield_inrow_add

chosen_model_name_yield_inrow <- if (keep_interaction_inrow) {
  "Interaction: inrow_weed_biomass_kg_ha * site_year"
} else {
  "Additive: inrow_weed_biomass_kg_ha + site_year"
}

aic_yield_inrow_out <- aic_yield_inrow |>
  mutate(
    Selected = if_else(model == chosen_model_name_yield_inrow, "Yes", "")
  )

aic_yield_inrow_out |>
  kable(
    digits  = 2,
    caption = "Adjusted soybean yield vs in-row weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for adjusted soybean yield vs in-row weed biomass:\n  ",
  chosen_model_name_yield_inrow,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_yield_inrow)
)

```

#### Slope summary via emtrends
```{r}
## Slope summary: emtrends for yield vs in-row weed biomass ------------

if (chosen_model_name_yield_inrow ==
    "Additive: inrow_weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope_inrow <- emtrends(
    yield_inrow_mod_best, ~ 1,
    var = "inrow_weed_biomass_kg_ha"
  )
  
  emm_slope_inrow_df <- summary(emm_slope_inrow, infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_inrow <- get_trend_col(
    emm_slope_inrow_df,
    "inrow_weed_biomass_kg_ha"
  )
  stat_col_inrow  <- get_stat_col(emm_slope_inrow_df)
  
  has_df_inrow <- "df" %in% names(emm_slope_inrow_df)
  has_p_inrow  <- "p.value" %in% names(emm_slope_inrow_df)
  
  slope_tbl_inrow <- emm_slope_inrow_df |>
    transmute(
      Scope              = "Pooled across site-years",
      slope_kg_per_kg    = .data[[trend_col_inrow]],
      slope_kg_per_100kg = .data[[trend_col_inrow]] * 100,
      SE,
      df         = if (has_df_inrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_inrow)) .data[[stat_col_inrow]] else NA_real_,
      p_value    = if (has_p_inrow) .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope_inrow <- emtrends(
    yield_inrow_mod_best, ~ site_year,
    var = "inrow_weed_biomass_kg_ha"
  )
  
  emm_slope_inrow_df <- summary(emm_slope_inrow, infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_inrow <- get_trend_col(
    emm_slope_inrow_df,
    "inrow_weed_biomass_kg_ha"
  )
  stat_col_inrow  <- get_stat_col(emm_slope_inrow_df)
  
  has_df_inrow <- "df" %in% names(emm_slope_inrow_df)
  has_p_inrow  <- "p.value" %in% names(emm_slope_inrow_df)
  
  slope_tbl_inrow <- emm_slope_inrow_df |>
    transmute(
      Scope              = as.character(site_year),
      slope_kg_per_kg    = .data[[trend_col_inrow]],
      slope_kg_per_100kg = .data[[trend_col_inrow]] * 100,
      SE,
      df         = if (has_df_inrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_inrow)) .data[[stat_col_inrow]] else NA_real_,
      p_value    = if (has_p_inrow) .data[["p.value"]] else NA_real_
    )
}

slope_tbl_inrow |>
  mutate(
    across(
      c(slope_kg_per_kg, slope_kg_per_100kg,
        SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of adjusted soybean yield vs in-row weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (kg yield per kg in-row weed)",
      "Slope (kg yield per 100 kg in-row weed)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Linear vs rectangular hyperbolic response
```{r}
## Linear vs rectangular hyperbolic response of yield to in-row weed biomass ----

# Use the same plot-level data (no random effects here, just a marginal comparison)
dat_yield_inrow <- reg_yield_inrow |>
  select(inrow_weed_biomass_kg_ha, bean_yield_adj_kg_ha) |>
  rename(
    B = inrow_weed_biomass_kg_ha,
    Y = bean_yield_adj_kg_ha
  )

# Quick sanity check (optional)
names(dat_yield_inrow)    # should be "B" and "Y"

# 1) Linear model -------------------------------------------------------

lin_mod_inrow <- lm(Y ~ B, data = dat_yield_inrow)

# 2) Rectangular hyperbolic model:
#    Y = Y0 - (Lmax * B) / (K + B)
#    Y0   = weed-free yield
#    Lmax = maximum yield loss
#    K    = in-row biomass at half of Lmax

start_vals_inrow <- list(
  Y0   = max(dat_yield_inrow$Y, na.rm = TRUE),  # near weed-free yield
  Lmax = max(dat_yield_inrow$Y, na.rm = TRUE) - min(dat_yield_inrow$Y, na.rm = TRUE),
  K    = median(dat_yield_inrow$B[dat_yield_inrow$B > 0], na.rm = TRUE)
)

hyper_mod_inrow <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_yield_inrow,
  start = start_vals_inrow
)

summary(hyper_mod_inrow)

# 3) AIC and R2 comparison -----------------------------------------------

# AIC
aic_lin_inrow   <- AIC(lin_mod_inrow)
aic_hyper_inrow <- AIC(hyper_mod_inrow)

# R2 for linear model
gl_lin_inrow <- broom::glance(lin_mod_inrow)
R2_lin_inrow     <- gl_lin_inrow$r.squared
R2_lin_inrow_adj <- gl_lin_inrow$adj.r.squared

# R2 for nls (hyperbolic)
get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res  # reconstruct observed response
  
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper_inrow <- get_r2_nls(hyper_mod_inrow)

# Residual df (more meaningful than n)
df_lin_inrow   <- df.residual(lin_mod_inrow)
df_hyper_inrow <- df.residual(hyper_mod_inrow)

mod_comp_inrow <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin_inrow, df_hyper_inrow),
  AIC   = c(aic_lin_inrow, aic_hyper_inrow),
  R2    = c(R2_lin_inrow, R2_hyper_inrow)
)

mod_comp_inrow |>
  kable(
    digits  = 3,
    caption = "Model comparison for adjusted soybean yield vs in-row weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Diagnostics
```{r}
## Diagnostics: yield vs in-row weed biomass ---------------------------

library(performance)

# Mixed model diagnostics (chosen in-row model)
check_model(yield_inrow_mod_best)

# Simple linear model diagnostics (in-row)
check_model(lin_mod_inrow)

# Optional specific checks for the linear model
check_heteroscedasticity(lin_mod_inrow)
check_normality(lin_mod_inrow)
check_outliers(lin_mod_inrow)

```

#### Visualisation
```{r}
ggplot(dat_yield_inrow, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: uses palette color and appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: always black, not tied to color scale
  stat_function(
    fun = function(x) coef(lin_mod_inrow)[1] + coef(lin_mod_inrow)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: uses palette color and appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod_inrow)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("dashed", "dotdash"),
        shape    = c(NA, NA),
        alpha    = c(1, 1)
      )
    )
  ) +
  labs(
    x = bquote("In-row weed biomass (kg ha"^-1*")"),
    y = bquote("Adjusted soybean yield (kg ha"^-1*")"),
    title = "Yield vs in-row weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )

```

### Biomass
#### Dataset
```{r}
# ----------------------------------------------------------------------
# Model testing: in-row weed biomass vs soybean biomass (kg ha^-1)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_biomass_inrow <- weed_biomass_clean |>
  filter(
    !is.na(inrow_weed_biomass_kg_ha),
    !is.na(bean_biomass_kg_ha),
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_biomass_inrow[, c("site_year", "block",
                          "inrow_weed_biomass_kg_ha",
                          "bean_biomass_kg_ha")])

```

#### Mixed model selection (site-year effects)
```{r}
### Mixed model selection (site-year effects): soybean biomass, in-row ---
## Mixed model selection: additive vs interaction ----------------------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-year (site_year as fixed effect)
biomass_inrow_add <- lmer(
  bean_biomass_kg_ha ~ inrow_weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_biomass_inrow
)

# Interaction: slope allowed to differ by site-year
biomass_inrow_int <- lmer(
  bean_biomass_kg_ha ~ inrow_weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_biomass_inrow
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_biomass_inrow   <- anova(biomass_inrow_add, biomass_inrow_int)
p_int_biomass_inrow <- lrt_biomass_inrow$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_biomass_inrow <- tibble(
  model = c(
    "Additive: inrow_weed_biomass_kg_ha + site_year",
    "Interaction: inrow_weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(biomass_inrow_add), AIC(biomass_inrow_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_biomass_inrow)
  )

# Selection rule: prefer simpler model unless interaction is clearly better
keep_interaction_biomass_inrow <- (p_int_biomass_inrow < 0.05) &&
  (aic_biomass_inrow$deltaAIC[
    aic_biomass_inrow$model ==
      "Interaction: inrow_weed_biomass_kg_ha * site_year"
  ] < 2)

biomass_inrow_mod_best <- if (keep_interaction_biomass_inrow)
  biomass_inrow_int else biomass_inrow_add

chosen_model_name_biomass_inrow <- if (keep_interaction_biomass_inrow) {
  "Interaction: inrow_weed_biomass_kg_ha * site_year"
} else {
  "Additive: inrow_weed_biomass_kg_ha + site_year"
}

aic_biomass_inrow_out <- aic_biomass_inrow |>
  mutate(
    Selected = if_else(model == chosen_model_name_biomass_inrow, "Yes", "")
  )

aic_biomass_inrow_out |>
  kable(
    digits  = 2,
    caption = "Soybean biomass vs in-row weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for soybean biomass vs in-row weed biomass:\n  ",
  chosen_model_name_biomass_inrow,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_biomass_inrow)
)

```

#### Slope summary via emtrends
```{r}
### Slope summary: soybean biomass vs in-row weed biomass ---------------

if (chosen_model_name_biomass_inrow ==
    "Additive: inrow_weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope_biomass_inrow <- emtrends(
    biomass_inrow_mod_best, ~ 1,
    var = "inrow_weed_biomass_kg_ha"
  )
  
  emm_slope_biomass_inrow_df <- summary(emm_slope_biomass_inrow,
                                        infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_biomass_inrow <- get_trend_col(
    emm_slope_biomass_inrow_df,
    "inrow_weed_biomass_kg_ha"
  )
  stat_col_biomass_inrow  <- get_stat_col(emm_slope_biomass_inrow_df)
  
  has_df_biomass_inrow <- "df" %in% names(emm_slope_biomass_inrow_df)
  has_p_biomass_inrow  <- "p.value" %in% names(emm_slope_biomass_inrow_df)
  
  slope_tbl_biomass_inrow <- emm_slope_biomass_inrow_df |>
    transmute(
      Scope                      = "Pooled across site-years",
      slope_kgbean_per_kgweed    = .data[[trend_col_biomass_inrow]],
      slope_kgbean_per_100kgweed = .data[[trend_col_biomass_inrow]] * 100,
      SE,
      df         = if (has_df_biomass_inrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_biomass_inrow))
        .data[[stat_col_biomass_inrow]] else NA_real_,
      p_value    = if (has_p_biomass_inrow) .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope_biomass_inrow <- emtrends(
    biomass_inrow_mod_best, ~ site_year,
    var = "inrow_weed_biomass_kg_ha"
  )
  
  emm_slope_biomass_inrow_df <- summary(emm_slope_biomass_inrow,
                                        infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_biomass_inrow <- get_trend_col(
    emm_slope_biomass_inrow_df,
    "inrow_weed_biomass_kg_ha"
  )
  stat_col_biomass_inrow  <- get_stat_col(emm_slope_biomass_inrow_df)
  
  has_df_biomass_inrow <- "df" %in% names(emm_slope_biomass_inrow_df)
  has_p_biomass_inrow  <- "p.value" %in% names(emm_slope_biomass_inrow_df)
  
  slope_tbl_biomass_inrow <- emm_slope_biomass_inrow_df |>
    transmute(
      Scope                      = as.character(site_year),
      slope_kgbean_per_kgweed    = .data[[trend_col_biomass_inrow]],
      slope_kgbean_per_100kgweed = .data[[trend_col_biomass_inrow]] * 100,
      SE,
      df         = if (has_df_biomass_inrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_biomass_inrow))
        .data[[stat_col_biomass_inrow]] else NA_real_,
      p_value    = if (has_p_biomass_inrow) .data[["p.value"]] else NA_real_
    )
}

slope_tbl_biomass_inrow |>
  mutate(
    across(
      c(slope_kgbean_per_kgweed,
        slope_kgbean_per_100kgweed,
        SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of soybean biomass vs in-row weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (kg bean per kg in-row weed)",
      "Slope (kg bean per 100 kg in-row weed)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


```

#### Linear vs rectangular hyperbolic response
```{r}
## Linear vs rectangular hyperbolic response of soybean biomass to in-row weed biomass ----

# Use the same plot-level data (no random effects here, just a marginal comparison)
dat_biomass_inrow <- reg_biomass_inrow |>
  select(inrow_weed_biomass_kg_ha, bean_biomass_kg_ha) |>
  rename(
    B = inrow_weed_biomass_kg_ha,
    Y = bean_biomass_kg_ha
  )

# Quick sanity check (optional)
names(dat_biomass_inrow)    # should be "B" and "Y"

# 1) Linear model -------------------------------------------------------

lin_mod_biomass_inrow <- lm(Y ~ B, data = dat_biomass_inrow)

# 2) Rectangular hyperbolic model:
#    Y = Y0 - (Lmax * B) / (K + B)
#    Y0   = weed-free soybean biomass
#    Lmax = maximum biomass loss
#    K    = in-row weed biomass at half of Lmax

start_vals_biomass_inrow <- list(
  Y0   = max(dat_biomass_inrow$Y, na.rm = TRUE),  # near weed-free biomass
  Lmax = max(dat_biomass_inrow$Y, na.rm = TRUE) - min(dat_biomass_inrow$Y, na.rm = TRUE),
  K    = median(dat_biomass_inrow$B[dat_biomass_inrow$B > 0], na.rm = TRUE)
)

hyper_mod_biomass_inrow <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_biomass_inrow,
  start = start_vals_biomass_inrow
)

summary(hyper_mod_biomass_inrow)

# 3) AIC and R2 comparison -----------------------------------------------

# AIC
aic_lin_biomass_inrow   <- AIC(lin_mod_biomass_inrow)
aic_hyper_biomass_inrow <- AIC(hyper_mod_biomass_inrow)

# R2 for linear model
gl_lin_biomass_inrow <- broom::glance(lin_mod_biomass_inrow)
R2_lin_biomass_inrow     <- gl_lin_biomass_inrow$r.squared
R2_lin_biomass_inrow_adj <- gl_lin_biomass_inrow$adj.r.squared

# R2 for nls (hyperbolic)
get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res  # reconstruct observed response
  
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper_biomass_inrow <- get_r2_nls(hyper_mod_biomass_inrow)

# Residual df (more meaningful than n)
df_lin_biomass_inrow   <- df.residual(lin_mod_biomass_inrow)
df_hyper_biomass_inrow <- df.residual(hyper_mod_biomass_inrow)

mod_comp_biomass_inrow <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin_biomass_inrow, df_hyper_biomass_inrow),
  AIC   = c(aic_lin_biomass_inrow, aic_hyper_biomass_inrow),
  R2    = c(R2_lin_biomass_inrow, R2_hyper_biomass_inrow)
)

mod_comp_biomass_inrow |>
  kable(
    digits  = 3,
    caption = "Model comparison for soybean biomass vs in-row weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Diagnostics
```{r}
## Diagnostics: soybean biomass vs in-row weed biomass ------------------

# Mixed model diagnostics (chosen in-row biomass model)
check_model(biomass_inrow_mod_best)

# Simple linear model diagnostics (biomass ~ in-row biomass)
check_model(lin_mod_biomass_inrow)

# Optional specific checks for the biomass linear model
check_heteroscedasticity(lin_mod_biomass_inrow)
check_normality(lin_mod_biomass_inrow)
check_outliers(lin_mod_biomass_inrow)

```

#### Visualisation
```{r}
ggplot(dat_biomass_inrow, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: uses palette color and appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: always black, not tied to color scale
  stat_function(
    fun = function(x) coef(lin_mod_biomass_inrow)[1] +
                      coef(lin_mod_biomass_inrow)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: uses palette color and appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod_biomass_inrow)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("dashed", "dotdash"),
        shape    = c(NA, NA),
        alpha    = c(1, 1)
      )
    )
  ) +
  labs(
    x = bquote("In-row weed biomass (kg ha"^-1*")"),
    y = bquote("Soybean biomass (kg ha"^-1*")"),
    title = "Soybean biomass vs in-row weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )

```

### Population
#### Dataset
```{r}
# ----------------------------------------------------------------------
# Model testing: in-row weed biomass vs bean population (plants ha^-1)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_pop_inrow <- weed_biomass_clean |>
  filter(
    !is.na(inrow_weed_biomass_kg_ha),
    !is.na(bean_population_hectare),
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_pop_inrow[, c("site_year", "block",
                      "inrow_weed_biomass_kg_ha",
                      "bean_population_hectare")])

```

#### Mixed model selection (site-year effects)
```{r}
### Mixed model selection (site-year effects): bean population, in-row ----
## Mixed model selection: additive vs interaction ----------------------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-year (site_year as fixed effect)
pop_inrow_add <- lmer(
  bean_population_hectare ~ inrow_weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_pop_inrow
)

# Interaction: slope allowed to differ by site-year
pop_inrow_int <- lmer(
  bean_population_hectare ~ inrow_weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_pop_inrow
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_pop_inrow   <- anova(pop_inrow_add, pop_inrow_int)
p_int_pop_inrow <- lrt_pop_inrow$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_pop_inrow <- tibble(
  model = c(
    "Additive: inrow_weed_biomass_kg_ha + site_year",
    "Interaction: inrow_weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(pop_inrow_add), AIC(pop_inrow_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_pop_inrow)
  )

# Selection rule: prefer simpler model unless interaction is clearly better
keep_interaction_pop_inrow <- (p_int_pop_inrow < 0.05) &&
  (aic_pop_inrow$deltaAIC[
    aic_pop_inrow$model ==
      "Interaction: inrow_weed_biomass_kg_ha * site_year"
  ] < 2)

pop_inrow_mod_best <- if (keep_interaction_pop_inrow)
  pop_inrow_int else pop_inrow_add

chosen_model_name_pop_inrow <- if (keep_interaction_pop_inrow) {
  "Interaction: inrow_weed_biomass_kg_ha * site_year"
} else {
  "Additive: inrow_weed_biomass_kg_ha + site_year"
}

aic_pop_inrow_out <- aic_pop_inrow |>
  mutate(
    Selected = if_else(model == chosen_model_name_pop_inrow, "Yes", "")
  )

aic_pop_inrow_out |>
  kable(
    digits  = 2,
    caption = "Bean population vs in-row weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for bean population vs in-row weed biomass:\n  ",
  chosen_model_name_pop_inrow,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_pop_inrow)
)

```

#### Slope summary via emtrends
```{r}
### Slope summary: bean population vs in-row weed biomass ---------------

if (chosen_model_name_pop_inrow ==
    "Additive: inrow_weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope_pop_inrow <- emtrends(
    pop_inrow_mod_best, ~ 1,
    var = "inrow_weed_biomass_kg_ha"
  )
  
  emm_slope_pop_inrow_df <- summary(emm_slope_pop_inrow,
                                    infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_pop_inrow <- get_trend_col(
    emm_slope_pop_inrow_df,
    "inrow_weed_biomass_kg_ha"
  )
  stat_col_pop_inrow  <- get_stat_col(emm_slope_pop_inrow_df)
  
  has_df_pop_inrow <- "df" %in% names(emm_slope_pop_inrow_df)
  has_p_pop_inrow  <- "p.value" %in% names(emm_slope_pop_inrow_df)
  
  slope_tbl_pop_inrow <- emm_slope_pop_inrow_df |>
    transmute(
      Scope                               = "Pooled across site-years",
      slope_plants_per_kg_inrowweed       = .data[[trend_col_pop_inrow]],
      slope_plants_per_100kg_inrowweed    = .data[[trend_col_pop_inrow]] * 100,
      SE,
      df         = if (has_df_pop_inrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_pop_inrow))
        .data[[stat_col_pop_inrow]] else NA_real_,
      p_value    = if (has_p_pop_inrow) .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope_pop_inrow <- emtrends(
    pop_inrow_mod_best, ~ site_year,
    var = "inrow_weed_biomass_kg_ha"
  )
  
  emm_slope_pop_inrow_df <- summary(emm_slope_pop_inrow,
                                    infer = c(TRUE, TRUE)) |>
    as_tibble()
  
  trend_col_pop_inrow <- get_trend_col(
    emm_slope_pop_inrow_df,
    "inrow_weed_biomass_kg_ha"
  )
  stat_col_pop_inrow  <- get_stat_col(emm_slope_pop_inrow_df)
  
  has_df_pop_inrow <- "df" %in% names(emm_slope_pop_inrow_df)
  has_p_pop_inrow  <- "p.value" %in% names(emm_slope_pop_inrow_df)
  
  slope_tbl_pop_inrow <- emm_slope_pop_inrow_df |>
    transmute(
      Scope                               = as.character(site_year),
      slope_plants_per_kg_inrowweed       = .data[[trend_col_pop_inrow]],
      slope_plants_per_100kg_inrowweed    = .data[[trend_col_pop_inrow]] * 100,
      SE,
      df         = if (has_df_pop_inrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_pop_inrow))
        .data[[stat_col_pop_inrow]] else NA_real_,
      p_value    = if (has_p_pop_inrow) .data[["p.value"]] else NA_real_
    )
}

slope_tbl_pop_inrow |>
  mutate(
    across(
      c(slope_plants_per_kg_inrowweed,
        slope_plants_per_100kg_inrowweed,
        SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of bean population vs in-row weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (plants ha^-1 per kg in-row weed ha^-1)",
      "Slope (plants ha^-1 per 100 kg in-row weed ha^-1)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Linear vs rectangular hyperbolic response
```{r}
## Linear vs rectangular hyperbolic response of bean population to in-row weed biomass ----

# Use the same plot-level data (no random effects here, just a marginal comparison)
dat_pop_inrow <- reg_pop_inrow |>
  select(inrow_weed_biomass_kg_ha, bean_population_hectare) |>
  rename(
    B = inrow_weed_biomass_kg_ha,
    Y = bean_population_hectare
  )

# Quick sanity check (optional)
names(dat_pop_inrow)    # should be "B" and "Y"

# 1) Linear model -------------------------------------------------------

lin_mod_pop_inrow <- lm(Y ~ B, data = dat_pop_inrow)

# 2) Rectangular hyperbolic model:
#    Y = Y0 - (Lmax * B) / (K + B)
#    Y0   = weed-free bean population (plants ha^-1)
#    Lmax = maximum population loss
#    K    = in-row weed biomass at half of Lmax

start_vals_pop_inrow <- list(
  Y0   = max(dat_pop_inrow$Y, na.rm = TRUE),  # near weed-free population
  Lmax = max(dat_pop_inrow$Y, na.rm = TRUE) - min(dat_pop_inrow$Y, na.rm = TRUE),
  K    = median(dat_pop_inrow$B[dat_pop_inrow$B > 0], na.rm = TRUE)
)

hyper_mod_pop_inrow <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_pop_inrow,
  start = start_vals_pop_inrow
)

summary(hyper_mod_pop_inrow)

# 3) AIC and R2 comparison -----------------------------------------------

# AIC
aic_lin_pop_inrow   <- AIC(lin_mod_pop_inrow)
aic_hyper_pop_inrow <- AIC(hyper_mod_pop_inrow)

# R2 for linear model
gl_lin_pop_inrow <- broom::glance(lin_mod_pop_inrow)
R2_lin_pop_inrow     <- gl_lin_pop_inrow$r.squared
R2_lin_pop_inrow_adj <- gl_lin_pop_inrow$adj.r.squared

# R2 for nls (hyperbolic)
get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res  # reconstruct observed response
  
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper_pop_inrow <- get_r2_nls(hyper_mod_pop_inrow)

# Residual df (more meaningful than n)
df_lin_pop_inrow   <- df.residual(lin_mod_pop_inrow)
df_hyper_pop_inrow <- df.residual(hyper_mod_pop_inrow)

mod_comp_pop_inrow <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin_pop_inrow, df_hyper_pop_inrow),
  AIC   = c(aic_lin_pop_inrow, aic_hyper_pop_inrow),
  R2    = c(R2_lin_pop_inrow, R2_hyper_pop_inrow)
)

mod_comp_pop_inrow |>
  kable(
    digits  = 3,
    caption = "Model comparison for bean population vs in-row weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```


#### Diagnostics
```{r}
# --- Bean population vs in-row weed biomass -----------------------------

# Mixed model diagnostics (chosen in-row population model)
check_model(pop_inrow_mod_best)

# Simple linear model diagnostics (population ~ in-row biomass)
check_model(lin_mod_pop_inrow)

# Optional specific checks for the in-row population linear model
check_heteroscedasticity(lin_mod_pop_inrow)
check_normality(lin_mod_pop_inrow)
check_outliers(lin_mod_pop_inrow)

```

#### Visualisation
```{r}
ggplot(dat_pop_inrow, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: uses palette color and appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: always black, not tied to color scale
  stat_function(
    fun = function(x) coef(lin_mod_pop_inrow)[1] +
                      coef(lin_mod_pop_inrow)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: uses palette color and appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod_pop_inrow)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  # y-axis as whole numbers, no scientific notation
  scale_y_continuous(
    labels = scales::label_comma(accuracy = 1)
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("dashed", "dotdash"),
        shape    = c(NA, NA),
        alpha    = c(1, 1)
      )
    )
  ) +
  labs(
    x = bquote("In-row weed biomass (kg ha"^-1*")"),
    y = bquote("Bean population (plants ha"^-1*")"),
    title = "Bean population vs in-row weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )

```
### Seed weight
#### Dataset
```{r}
# ----------------------------------------------------------------------
# Model testing: in-row weed biomass vs 100-seed weight (g)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_100sw_inrow <- weed_biomass_clean |>
  filter(
    !is.na(inrow_weed_biomass_kg_ha),
    !is.na(seed_weight),   # 100-seed weight column
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_100sw_inrow[, c("site_year", "block",
                        "inrow_weed_biomass_kg_ha",
                        "seed_weight")])

```


#### Mixed model selection (site-year effects)
```{r}
### Mixed model selection (site-year effects): 100-seed weight, in-row ----
## Mixed model selection: additive vs interaction ----------------------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-year (site_year as fixed effect)
sw100_inrow_add <- lmer(
  seed_weight ~ inrow_weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_100sw_inrow
)

# Interaction: slope allowed to differ by site-year
sw100_inrow_int <- lmer(
  seed_weight ~ inrow_weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_100sw_inrow
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_sw100_inrow   <- anova(sw100_inrow_add, sw100_inrow_int)
p_int_sw100_inrow <- lrt_sw100_inrow$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_sw100_inrow <- tibble(
  model = c(
    "Additive: inrow_weed_biomass_kg_ha + site_year",
    "Interaction: inrow_weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(sw100_inrow_add), AIC(sw100_inrow_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_sw100_inrow)
  )

# Selection rule: prefer simpler model unless interaction is clearly better
keep_interaction_sw100_inrow <- (p_int_sw100_inrow < 0.05) &&
  (aic_sw100_inrow$deltaAIC[
    aic_sw100_inrow$model ==
      "Interaction: inrow_weed_biomass_kg_ha * site_year"
  ] < 2)

sw100_inrow_mod_best <- if (keep_interaction_sw100_inrow)
  sw100_inrow_int else sw100_inrow_add

chosen_model_name_sw100_inrow <- if (keep_interaction_sw100_inrow) {
  "Interaction: inrow_weed_biomass_kg_ha * site_year"
} else {
  "Additive: inrow_weed_biomass_kg_ha + site_year"
}

aic_sw100_inrow_out <- aic_sw100_inrow |>
  mutate(
    Selected = if_else(model == chosen_model_name_sw100_inrow, "Yes", "")
  )

aic_sw100_inrow_out |>
  kable(
    digits  = 2,
    caption = "100-seed weight vs in-row weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for 100-seed weight vs in-row weed biomass:\n  ",
  chosen_model_name_sw100_inrow,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_sw100_inrow)
)

```

#### Slope summary via emtrends
```{r}
### Slope summary: 100-seed weight vs in-row weed biomass ---------------

if (chosen_model_name_sw100_inrow ==
    "Additive: inrow_weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope_sw100_inrow <- emtrends(
    sw100_inrow_mod_best, ~ 1,
    var = "inrow_weed_biomass_kg_ha"
  )
  
  emm_slope_sw100_inrow_df <- summary(
    emm_slope_sw100_inrow,
    infer = c(TRUE, TRUE)
  ) |>
    as_tibble()
  
  trend_col_sw100_inrow <- get_trend_col(
    emm_slope_sw100_inrow_df,
    "inrow_weed_biomass_kg_ha"
  )
  stat_col_sw100_inrow  <- get_stat_col(emm_slope_sw100_inrow_df)
  
  has_df_sw100_inrow <- "df" %in% names(emm_slope_sw100_inrow_df)
  has_p_sw100_inrow  <- "p.value" %in% names(emm_slope_sw100_inrow_df)
  
  slope_tbl_sw100_inrow <- emm_slope_sw100_inrow_df |>
    transmute(
      Scope                               = "Pooled across site-years",
      slope_g_per_kg_inrowweed            = .data[[trend_col_sw100_inrow]],
      slope_g_per_100kg_inrowweed         = .data[[trend_col_sw100_inrow]] * 100,
      SE,
      df         = if (has_df_sw100_inrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_sw100_inrow))
        .data[[stat_col_sw100_inrow]] else NA_real_,
      p_value    = if (has_p_sw100_inrow)
        .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope_sw100_inrow <- emtrends(
    sw100_inrow_mod_best, ~ site_year,
    var = "inrow_weed_biomass_kg_ha"
  )
  
  emm_slope_sw100_inrow_df <- summary(
    emm_slope_sw100_inrow,
    infer = c(TRUE, TRUE)
  ) |>
    as_tibble()
  
  trend_col_sw100_inrow <- get_trend_col(
    emm_slope_sw100_inrow_df,
    "inrow_weed_biomass_kg_ha"
  )
  stat_col_sw100_inrow  <- get_stat_col(emm_slope_sw100_inrow_df)
  
  has_df_sw100_inrow <- "df" %in% names(emm_slope_sw100_inrow_df)
  has_p_sw100_inrow  <- "p.value" %in% names(emm_slope_sw100_inrow_df)
  
  slope_tbl_sw100_inrow <- emm_slope_sw100_inrow_df |>
    transmute(
      Scope                               = as.character(site_year),
      slope_g_per_kg_inrowweed            = .data[[trend_col_sw100_inrow]],
      slope_g_per_100kg_inrowweed         = .data[[trend_col_sw100_inrow]] * 100,
      SE,
      df         = if (has_df_sw100_inrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_sw100_inrow))
        .data[[stat_col_sw100_inrow]] else NA_real_,
      p_value    = if (has_p_sw100_inrow)
        .data[["p.value"]] else NA_real_
    )
}

slope_tbl_sw100_inrow |>
  mutate(
    across(
      c(slope_g_per_kg_inrowweed,
        slope_g_per_100kg_inrowweed,
        SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of 100-seed weight vs in-row weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (g per kg in-row weed ha^-1)",
      "Slope (g per 100 kg in-row weed ha^-1)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Linear vs rectangular hyperbolic response
```{r}
## Linear vs rectangular hyperbolic response of 100-seed weight to in-row weed biomass ----

# Use the same plot-level data (no random effects here, just a marginal comparison)
dat_sw100_inrow <- reg_100sw_inrow |>
  select(inrow_weed_biomass_kg_ha, seed_weight) |>
  rename(
    B = inrow_weed_biomass_kg_ha,
    Y = seed_weight
  )

# Quick sanity check (optional)
names(dat_sw100_inrow)    # should be "B" and "Y"

# 1) Linear model -------------------------------------------------------

lin_mod_sw100_inrow <- lm(Y ~ B, data = dat_sw100_inrow)

# 2) Rectangular hyperbolic model:
#    Y = Y0 - (Lmax * B) / (K + B)
#    Y0   = weed-free 100-seed weight (g)
#    Lmax = maximum loss in 100-seed weight
#    K    = in-row weed biomass at half of Lmax

start_vals_sw100_inrow <- list(
  Y0   = max(dat_sw100_inrow$Y, na.rm = TRUE),  # near weed-free 100-seed weight
  Lmax = max(dat_sw100_inrow$Y, na.rm = TRUE) - min(dat_sw100_inrow$Y, na.rm = TRUE),
  K    = median(dat_sw100_inrow$B[dat_sw100_inrow$B > 0], na.rm = TRUE)
)

hyper_mod_sw100_inrow <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_sw100_inrow,
  start = start_vals_sw100_inrow
)

summary(hyper_mod_sw100_inrow)

# 3) AIC and R2 comparison -----------------------------------------------

# AIC
aic_lin_sw100_inrow   <- AIC(lin_mod_sw100_inrow)
aic_hyper_sw100_inrow <- AIC(hyper_mod_sw100_inrow)

# R2 for linear model
gl_lin_sw100_inrow  <- broom::glance(lin_mod_sw100_inrow)
R2_lin_sw100_inrow  <- gl_lin_sw100_inrow$r.squared
R2_lin_sw100_inrow_adj <- gl_lin_sw100_inrow$adj.r.squared

# R2 for nls (hyperbolic)
get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res  # reconstruct observed response
  
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper_sw100_inrow <- get_r2_nls(hyper_mod_sw100_inrow)

# Residual df (more meaningful than n)
df_lin_sw100_inrow   <- df.residual(lin_mod_sw100_inrow)
df_hyper_sw100_inrow <- df.residual(hyper_mod_sw100_inrow)

mod_comp_sw100_inrow <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin_sw100_inrow, df_hyper_sw100_inrow),
  AIC   = c(aic_lin_sw100_inrow, aic_hyper_sw100_inrow),
  R2    = c(R2_lin_sw100_inrow, R2_hyper_sw100_inrow)
)

mod_comp_sw100_inrow |>
  kable(
    digits  = 3,
    caption = "Model comparison for 100-seed weight vs in-row weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Diagnostics
```{r}
# --- 100-seed weight vs in-row weed biomass ---------------------------

# Mixed model diagnostics (chosen 100-seed weight in-row model)
check_model(sw100_inrow_mod_best)

# Simple linear model diagnostics (100-seed weight ~ in-row biomass)
check_model(lin_mod_sw100_inrow)

# Optional specific checks for the 100-seed weight linear model (in-row)
check_heteroscedasticity(lin_mod_sw100_inrow)
check_normality(lin_mod_sw100_inrow)
check_outliers(lin_mod_sw100_inrow)

```

#### Visualisation
```{r}
ggplot(dat_sw100_inrow, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: uses palette color and appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: always black, not tied to color scale
  stat_function(
    fun = function(x) coef(lin_mod_sw100_inrow)[1] +
                      coef(lin_mod_sw100_inrow)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: uses palette color and appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod_sw100_inrow)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("dashed", "dotdash"),
        shape    = c(NA, NA),
        alpha    = c(1, 1)
      )
    )
  ) +
  labs(
    x = bquote("In-row weed biomass (kg ha"^-1*")"),
    y = "100-seed weight (g)",
    title = "100-seed weight vs in-row weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )

```
## Inter-row weed biomass
### Yield
#### Dataset

```{r}
# ----------------------------------------------------------------------
# Model testing: interrow weed biomass vs adjusted soybean yield (kg ha^-1)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_yield_interrow <- weed_biomass_clean |>
  filter(
    !is.na(interrow_weed_biomass_kg_ha),
    !is.na(bean_yield_adj_kg_ha),
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_yield_interrow[, c("site_year", "block",
                           "interrow_weed_biomass_kg_ha",
                           "bean_yield_adj_kg_ha")])

```

#### Mixed model selection (site-year effects)

```{r}
## Mixed model selection: interrow weed biomass vs adjusted yield ---------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-year (site_year as fixed effect)
yield_interrow_add <- lmer(
  bean_yield_adj_kg_ha ~ interrow_weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_yield_interrow
)

# Interaction: slope allowed to differ by site-year
yield_interrow_int <- lmer(
  bean_yield_adj_kg_ha ~ interrow_weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_yield_interrow
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_yield_interrow   <- anova(yield_interrow_add, yield_interrow_int)
p_int_yield_interrow <- lrt_yield_interrow$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_yield_interrow <- tibble(
  model = c(
    "Additive: interrow_weed_biomass_kg_ha + site_year",
    "Interaction: interrow_weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(yield_interrow_add), AIC(yield_interrow_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_yield_interrow)
  )

# Selection rule: prefer simpler model unless interaction is clearly better
keep_interaction_interrow <- (p_int_yield_interrow < 0.05) &&
  (aic_yield_interrow$deltaAIC[
    aic_yield_interrow$model ==
      "Interaction: interrow_weed_biomass_kg_ha * site_year"
  ] < 2)

yield_interrow_mod_best <- if (keep_interaction_interrow)
  yield_interrow_int else yield_interrow_add

chosen_model_name_yield_interrow <- if (keep_interaction_interrow) {
  "Interaction: interrow_weed_biomass_kg_ha * site_year"
} else {
  "Additive: interrow_weed_biomass_kg_ha + site_year"
}

aic_yield_interrow_out <- aic_yield_interrow |>
  mutate(
    Selected = if_else(model == chosen_model_name_yield_interrow, "Yes", "")
  )

aic_yield_interrow_out |>
  kable(
    digits  = 2,
    caption = "Adjusted soybean yield vs interrow weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for adjusted soybean yield vs interrow weed biomass:\n  ",
  chosen_model_name_yield_interrow,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_yield_interrow)
)

```

#### Slope summary via emtrends
```{r}
## Slope summary: emtrends for yield vs interrow weed biomass ------------

if (chosen_model_name_yield_interrow ==
    "Additive: interrow_weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope_interrow <- emtrends(
    yield_interrow_mod_best, ~ 1,
    var = "interrow_weed_biomass_kg_ha"
  )
  
  emm_slope_interrow_df <- summary(
    emm_slope_interrow,
    infer = c(TRUE, TRUE)
  ) |>
    as_tibble()
  
  trend_col_interrow <- get_trend_col(
    emm_slope_interrow_df,
    "interrow_weed_biomass_kg_ha"
  )
  stat_col_interrow  <- get_stat_col(emm_slope_interrow_df)
  
  has_df_interrow <- "df" %in% names(emm_slope_interrow_df)
  has_p_interrow  <- "p.value" %in% names(emm_slope_interrow_df)
  
  slope_tbl_interrow <- emm_slope_interrow_df |>
    transmute(
      Scope              = "Pooled across site-years",
      slope_kg_per_kg    = .data[[trend_col_interrow]],
      slope_kg_per_100kg = .data[[trend_col_interrow]] * 100,
      SE,
      df         = if (has_df_interrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_interrow))
        .data[[stat_col_interrow]] else NA_real_,
      p_value    = if (has_p_interrow)
        .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope_interrow <- emtrends(
    yield_interrow_mod_best, ~ site_year,
    var = "interrow_weed_biomass_kg_ha"
  )
  
  emm_slope_interrow_df <- summary(
    emm_slope_interrow,
    infer = c(TRUE, TRUE)
  ) |>
    as_tibble()
  
  trend_col_interrow <- get_trend_col(
    emm_slope_interrow_df,
    "interrow_weed_biomass_kg_ha"
  )
  stat_col_interrow  <- get_stat_col(emm_slope_interrow_df)
  
  has_df_interrow <- "df" %in% names(emm_slope_interrow_df)
  has_p_interrow  <- "p.value" %in% names(emm_slope_interrow_df)
  
  slope_tbl_interrow <- emm_slope_interrow_df |>
    transmute(
      Scope              = as.character(site_year),
      slope_kg_per_kg    = .data[[trend_col_interrow]],
      slope_kg_per_100kg = .data[[trend_col_interrow]] * 100,
      SE,
      df         = if (has_df_interrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_interrow))
        .data[[stat_col_interrow]] else NA_real_,
      p_value    = if (has_p_interrow)
        .data[["p.value"]] else NA_real_
    )
}

slope_tbl_interrow |>
  mutate(
    across(
      c(slope_kg_per_kg, slope_kg_per_100kg,
        SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of adjusted soybean yield vs interrow weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (kg yield per kg interrow weed)",
      "Slope (kg yield per 100 kg interrow weed)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Linear vs rectangular hyperbolic response
```{r}
## Linear vs rectangular hyperbolic response of yield to interrow weed biomass ----

# Use the same plot-level data (no random effects here, just a marginal comparison)
dat_yield_interrow <- reg_yield_interrow |>
  select(interrow_weed_biomass_kg_ha, bean_yield_adj_kg_ha) |>
  rename(
    B = interrow_weed_biomass_kg_ha,
    Y = bean_yield_adj_kg_ha
  )

# Quick sanity check (optional)
names(dat_yield_interrow)    # should be "B" and "Y"

# 1) Linear model -------------------------------------------------------

lin_mod_interrow <- lm(Y ~ B, data = dat_yield_interrow)

# 2) Rectangular hyperbolic model:
#    Y = Y0 - (Lmax * B) / (K + B)
#    Y0   = weed-free yield
#    Lmax = maximum yield loss
#    K    = interrow biomass at half of Lmax

start_vals_interrow <- list(
  Y0   = max(dat_yield_interrow$Y, na.rm = TRUE),  # near weed-free yield
  Lmax = max(dat_yield_interrow$Y, na.rm = TRUE) -
         min(dat_yield_interrow$Y, na.rm = TRUE),
  K    = median(dat_yield_interrow$B[dat_yield_interrow$B > 0], na.rm = TRUE)
)

hyper_mod_interrow <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_yield_interrow,
  start = start_vals_interrow
)

summary(hyper_mod_interrow)

# 3) AIC and R2 comparison -----------------------------------------------

# AIC
aic_lin_interrow   <- AIC(lin_mod_interrow)
aic_hyper_interrow <- AIC(hyper_mod_interrow)

# R2 for linear model
gl_lin_interrow     <- broom::glance(lin_mod_interrow)
R2_lin_interrow     <- gl_lin_interrow$r.squared
R2_lin_interrow_adj <- gl_lin_interrow$adj.r.squared

# R2 for nls (hyperbolic)
get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res  # reconstruct observed response
  
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper_interrow <- get_r2_nls(hyper_mod_interrow)

# Residual df (more meaningful than n)
df_lin_interrow   <- df.residual(lin_mod_interrow)
df_hyper_interrow <- df.residual(hyper_mod_interrow)

mod_comp_interrow <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin_interrow, df_hyper_interrow),
  AIC   = c(aic_lin_interrow, aic_hyper_interrow),
  R2    = c(R2_lin_interrow, R2_hyper_interrow)
)

mod_comp_interrow |>
  kable(
    digits  = 3,
    caption = "Model comparison for adjusted soybean yield vs interrow weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


```

#### Diagnostics
```{r}
## Diagnostics & visualization: yield vs interrow weed biomass ---------

library(performance)

# Mixed model diagnostics (chosen interrow model)
check_model(yield_interrow_mod_best)

# Simple linear model diagnostics (interrow)
check_model(lin_mod_interrow)

# Optional specific checks for the interrow linear model
check_heteroscedasticity(lin_mod_interrow)
check_normality(lin_mod_interrow)
check_outliers(lin_mod_interrow)

```

#### Visualisation
```{r}
ggplot(dat_yield_interrow, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: uses palette color and appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: always black, not tied to color scale
  stat_function(
    fun = function(x) coef(lin_mod_interrow)[1] + coef(lin_mod_interrow)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: uses palette color and appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod_interrow)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("dashed", "dotdash"),
        shape    = c(NA, NA),
        alpha    = c(1, 1)
      )
    )
  ) +
  labs(
    x = bquote("Interrow weed biomass (kg ha"^-1*")"),
    y = bquote("Adjusted soybean yield (kg ha"^-1*")"),
    title = "Yield vs interrow weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )

```

### Biomass
#### Dataset
```{r}
# ----------------------------------------------------------------------
# Model testing: interrow weed biomass vs soybean biomass (kg ha^-1)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_biomass_interrow <- weed_biomass_clean |>
  filter(
    !is.na(interrow_weed_biomass_kg_ha),
    !is.na(bean_biomass_kg_ha),
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_biomass_interrow[, c("site_year", "block",
                             "interrow_weed_biomass_kg_ha",
                             "bean_biomass_kg_ha")])

```

#### Mixed model selection (site-year effects)
```{r}
### Mixed model selection (site-year effects): soybean biomass ---------
## Mixed model selection: additive vs interaction ----------------------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-year (site_year as fixed effect)
biomass_interrow_add <- lmer(
  bean_biomass_kg_ha ~ interrow_weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_biomass_interrow
)

# Interaction: slope allowed to differ by site-year
biomass_interrow_int <- lmer(
  bean_biomass_kg_ha ~ interrow_weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_biomass_interrow
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_biomass_interrow   <- anova(biomass_interrow_add, biomass_interrow_int)
p_int_biomass_interrow <- lrt_biomass_interrow$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_biomass_interrow <- tibble(
  model = c(
    "Additive: interrow_weed_biomass_kg_ha + site_year",
    "Interaction: interrow_weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(biomass_interrow_add), AIC(biomass_interrow_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_biomass_interrow)
  )

# Selection rule: prefer simpler model unless interaction is clearly better
keep_interaction_biomass_interrow <- (p_int_biomass_interrow < 0.05) &&
  (aic_biomass_interrow$deltaAIC[
    aic_biomass_interrow$model ==
      "Interaction: interrow_weed_biomass_kg_ha * site_year"
  ] < 2)

biomass_interrow_mod_best <-
  if (keep_interaction_biomass_interrow) biomass_interrow_int else biomass_interrow_add

chosen_model_name_biomass_interrow <- if (keep_interaction_biomass_interrow) {
  "Interaction: interrow_weed_biomass_kg_ha * site_year"
} else {
  "Additive: interrow_weed_biomass_kg_ha + site_year"
}

aic_biomass_interrow_out <- aic_biomass_interrow |>
  mutate(
    Selected = if_else(model == chosen_model_name_biomass_interrow, "Yes", "")
  )

aic_biomass_interrow_out |>
  kable(
    digits  = 2,
    caption = "Soybean biomass vs interrow weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for soybean biomass vs interrow weed biomass:\n  ",
  chosen_model_name_biomass_interrow,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_biomass_interrow)
)

```


#### Slope summary via emtrends
```{r}
### Slope summary: soybean biomass vs interrow weed biomass ---------------

if (chosen_model_name_biomass_interrow ==
    "Additive: interrow_weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope_biomass_interrow <- emtrends(
    biomass_interrow_mod_best, ~ 1,
    var = "interrow_weed_biomass_kg_ha"
  )
  
  emm_slope_biomass_interrow_df <- summary(
    emm_slope_biomass_interrow, infer = c(TRUE, TRUE)
  ) |>
    as_tibble()
  
  trend_col_biomass_interrow <- get_trend_col(
    emm_slope_biomass_interrow_df,
    "interrow_weed_biomass_kg_ha"
  )
  stat_col_biomass_interrow  <- get_stat_col(emm_slope_biomass_interrow_df)
  
  has_df_biomass_interrow <- "df" %in% names(emm_slope_biomass_interrow_df)
  has_p_biomass_interrow  <- "p.value" %in% names(emm_slope_biomass_interrow_df)
  
  slope_tbl_biomass_interrow <- emm_slope_biomass_interrow_df |>
    transmute(
      Scope                        = "Pooled across site-years",
      slope_kgbean_per_kginterrow  = .data[[trend_col_biomass_interrow]],
      slope_kgbean_per_100kginterrow =
        .data[[trend_col_biomass_interrow]] * 100,
      SE,
      df         = if (has_df_biomass_interrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_biomass_interrow))
                     .data[[stat_col_biomass_interrow]] else NA_real_,
      p_value    = if (has_p_biomass_interrow)
                     .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope_biomass_interrow <- emtrends(
    biomass_interrow_mod_best, ~ site_year,
    var = "interrow_weed_biomass_kg_ha"
  )
  
  emm_slope_biomass_interrow_df <- summary(
    emm_slope_biomass_interrow, infer = c(TRUE, TRUE)
  ) |>
    as_tibble()
  
  trend_col_biomass_interrow <- get_trend_col(
    emm_slope_biomass_interrow_df,
    "interrow_weed_biomass_kg_ha"
  )
  stat_col_biomass_interrow  <- get_stat_col(emm_slope_biomass_interrow_df)
  
  has_df_biomass_interrow <- "df" %in% names(emm_slope_biomass_interrow_df)
  has_p_biomass_interrow  <- "p.value" %in% names(emm_slope_biomass_interrow_df)
  
  slope_tbl_biomass_interrow <- emm_slope_biomass_interrow_df |>
    transmute(
      Scope                        = as.character(site_year),
      slope_kgbean_per_kginterrow  = .data[[trend_col_biomass_interrow]],
      slope_kgbean_per_100kginterrow =
        .data[[trend_col_biomass_interrow]] * 100,
      SE,
      df         = if (has_df_biomass_interrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_biomass_interrow))
                     .data[[stat_col_biomass_interrow]] else NA_real_,
      p_value    = if (has_p_biomass_interrow)
                     .data[["p.value"]] else NA_real_
    )
}

slope_tbl_biomass_interrow |>
  mutate(
    across(
      c(slope_kgbean_per_kginterrow,
        slope_kgbean_per_100kginterrow,
        SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of soybean biomass vs interrow weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (kg biomass per kg interrow weed)",
      "Slope (kg biomass per 100 kg interrow weed)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Linear vs rectangular hyperbolic response
```{r}
## Linear vs rectangular hyperbolic response of soybean biomass
## to interrow weed biomass ----

# Plot-level data (no random effects here)
dat_biomass_interrow <- reg_biomass_interrow |>
  select(interrow_weed_biomass_kg_ha, bean_biomass_kg_ha) |>
  rename(
    B = interrow_weed_biomass_kg_ha,
    Y = bean_biomass_kg_ha
  )

names(dat_biomass_interrow)  # should be "B" and "Y"

# 1) Linear model -------------------------------------------------------

lin_mod_biomass_interrow <- lm(Y ~ B, data = dat_biomass_interrow)

# 2) Rectangular hyperbolic model:
#    Y = Y0 - (Lmax * B) / (K + B)

start_vals_biomass_interrow <- list(
  Y0   = max(dat_biomass_interrow$Y, na.rm = TRUE),
  Lmax = max(dat_biomass_interrow$Y, na.rm = TRUE) -
         min(dat_biomass_interrow$Y, na.rm = TRUE),
  K    = median(dat_biomass_interrow$B[dat_biomass_interrow$B > 0],
                na.rm = TRUE)
)

hyper_mod_biomass_interrow <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_biomass_interrow,
  start = start_vals_biomass_interrow
)

summary(hyper_mod_biomass_interrow)

# 3) AIC and R² comparison ----------------------------------------------

aic_lin_biomass_interrow   <- AIC(lin_mod_biomass_interrow)
aic_hyper_biomass_interrow <- AIC(hyper_mod_biomass_interrow)

gl_lin_biomass_interrow <- broom::glance(lin_mod_biomass_interrow)
R2_lin_biomass_interrow <- gl_lin_biomass_interrow$r.squared

get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper_biomass_interrow <- get_r2_nls(hyper_mod_biomass_interrow)

df_lin_biomass_interrow   <- df.residual(lin_mod_biomass_interrow)
df_hyper_biomass_interrow <- df.residual(hyper_mod_biomass_interrow)

mod_comp_biomass_interrow <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin_biomass_interrow, df_hyper_biomass_interrow),
  AIC   = c(aic_lin_biomass_interrow, aic_hyper_biomass_interrow),
  R2    = c(R2_lin_biomass_interrow, R2_hyper_biomass_interrow)
)

mod_comp_biomass_interrow |>
  kable(
    digits  = 3,
    caption = "Model comparison for soybean biomass vs interrow weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))


```

#### Diagnostics
```{r}
# --- Soybean biomass vs interrow weed biomass -------------------------

# Mixed model diagnostics (chosen biomass model)
check_model(biomass_interrow_mod_best)

# Simple linear model diagnostics (biomass ~ interrow biomass)
check_model(lin_mod_biomass_interrow)

# Optional specific checks for the biomass linear model
check_heteroscedasticity(lin_mod_biomass_interrow)
check_normality(lin_mod_biomass_interrow)
check_outliers(lin_mod_biomass_interrow)

```

#### Visualisation
```{r}
ggplot(dat_biomass_interrow, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: uses palette color and appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: always black, not tied to color scale
  stat_function(
    fun = function(x) coef(lin_mod_biomass_interrow)[1] +
                     coef(lin_mod_biomass_interrow)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: uses palette color and appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod_biomass_interrow)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  labs(
    x = bquote("Interrow weed biomass (kg ha"^-1*")"),
    y = bquote("Soybean biomass (kg ha"^-1*")"),
    title = "Soybean biomass vs interrow weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )

```
### Population
#### Dataset
```{r}
# ----------------------------------------------------------------------
# Model testing: interrow weed biomass vs bean population (plants ha^-1)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_pop_interrow <- weed_biomass_clean |>
  filter(
    !is.na(interrow_weed_biomass_kg_ha),
    !is.na(bean_population_hectare),
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_pop_interrow[, c("site_year", "block",
                         "interrow_weed_biomass_kg_ha",
                         "bean_population_hectare")])

```

#### Mixed model selection (site-year effects)
```{r}
## Mixed model selection: site-year effects (bean population) ----------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-years
pop_interrow_add <- lmer(
  bean_population_hectare ~ interrow_weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_pop_interrow
)

# Interaction: slope allowed to differ by site-year
pop_interrow_int <- lmer(
  bean_population_hectare ~ interrow_weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_pop_interrow
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_pop_interrow   <- anova(pop_interrow_add, pop_interrow_int)
p_int_pop_interrow <- lrt_pop_interrow$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_pop_interrow <- tibble(
  model = c(
    "Additive: interrow_weed_biomass_kg_ha + site_year",
    "Interaction: interrow_weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(pop_interrow_add), AIC(pop_interrow_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_pop_interrow)
  )

# Selection rule: prefer simpler model unless interaction clearly better
keep_interaction_pop_interrow <- (p_int_pop_interrow < 0.05) &&
  (aic_pop_interrow$deltaAIC[
    aic_pop_interrow$model ==
      "Interaction: interrow_weed_biomass_kg_ha * site_year"
  ] < 2)

pop_interrow_mod_best <- if (keep_interaction_pop_interrow)
  pop_interrow_int else pop_interrow_add

chosen_model_name_pop_interrow <- if (keep_interaction_pop_interrow) {
  "Interaction: interrow_weed_biomass_kg_ha * site_year"
} else {
  "Additive: interrow_weed_biomass_kg_ha + site_year"
}

aic_pop_interrow_out <- aic_pop_interrow |>
  mutate(
    Selected = if_else(model == chosen_model_name_pop_interrow, "Yes", "")
  )

aic_pop_interrow_out |>
  kable(
    digits  = 2,
    caption = "Bean population vs interrow weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for bean population vs interrow weed biomass:\n  ",
  chosen_model_name_pop_interrow,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_pop_interrow)
)

```

#### Slope summary via emtrends
```{r}
### Slope summary: bean population vs interrow weed biomass ------------

if (chosen_model_name_pop_interrow ==
    "Additive: interrow_weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope_pop_interrow <- emtrends(
    pop_interrow_mod_best, ~ 1,
    var = "interrow_weed_biomass_kg_ha"
  )
  
  emm_slope_pop_interrow_df <- summary(
    emm_slope_pop_interrow,
    infer = c(TRUE, TRUE)
  ) |>
    as_tibble()
  
  trend_col_pop_interrow <- get_trend_col(
    emm_slope_pop_interrow_df,
    "interrow_weed_biomass_kg_ha"
  )
  stat_col_pop_interrow  <- get_stat_col(emm_slope_pop_interrow_df)
  
  has_df_pop_interrow <- "df" %in% names(emm_slope_pop_interrow_df)
  has_p_pop_interrow  <- "p.value" %in% names(emm_slope_pop_interrow_df)
  
  slope_tbl_pop_interrow <- emm_slope_pop_interrow_df |>
    transmute(
      Scope                      = "Pooled across site-years",
      slope_plants_per_kgweed    = .data[[trend_col_pop_interrow]],
      slope_plants_per_100kgweed = .data[[trend_col_pop_interrow]] * 100,
      SE,
      df         = if (has_df_pop_interrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_pop_interrow))
                     .data[[stat_col_pop_interrow]] else NA_real_,
      p_value    = if (has_p_pop_interrow)
                     .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope_pop_interrow <- emtrends(
    pop_interrow_mod_best, ~ site_year,
    var = "interrow_weed_biomass_kg_ha"
  )
  
  emm_slope_pop_interrow_df <- summary(
    emm_slope_pop_interrow,
    infer = c(TRUE, TRUE)
  ) |>
    as_tibble()
  
  trend_col_pop_interrow <- get_trend_col(
    emm_slope_pop_interrow_df,
    "interrow_weed_biomass_kg_ha"
  )
  stat_col_pop_interrow  <- get_stat_col(emm_slope_pop_interrow_df)
  
  has_df_pop_interrow <- "df" %in% names(emm_slope_pop_interrow_df)
  has_p_pop_interrow  <- "p.value" %in% names(emm_slope_pop_interrow_df)
  
  slope_tbl_pop_interrow <- emm_slope_pop_interrow_df |>
    transmute(
      Scope                      = as.character(site_year),
      slope_plants_per_kgweed    = .data[[trend_col_pop_interrow]],
      slope_plants_per_100kgweed = .data[[trend_col_pop_interrow]] * 100,
      SE,
      df         = if (has_df_pop_interrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_pop_interrow))
                     .data[[stat_col_pop_interrow]] else NA_real_,
      p_value    = if (has_p_pop_interrow)
                     .data[["p.value"]] else NA_real_
    )
}

slope_tbl_pop_interrow |>
  mutate(
    across(
      c(slope_plants_per_kgweed,
        slope_plants_per_100kgweed,
        SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of bean population vs interrow weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (plants ha^-1 per kg interrow weed ha^-1)",
      "Slope (plants ha^-1 per 100 kg interrow weed ha^-1)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Linear vs rectangular hyperbolic response
```{r}
## Linear vs rectangular hyperbolic response of bean population 
## to interrow weed biomass --------------------------------------------

# Use the same plot-level data (no random effects here)
dat_pop_interrow <- reg_pop_interrow |>
  select(interrow_weed_biomass_kg_ha, bean_population_hectare) |>
  rename(
    B = interrow_weed_biomass_kg_ha,
    Y = bean_population_hectare
  )

# Quick sanity check
names(dat_pop_interrow)  # "B", "Y"

# 1) Linear model ------------------------------------------------------

lin_mod_pop_interrow <- lm(Y ~ B, data = dat_pop_interrow)

# 2) Rectangular hyperbolic model:
#    Y = Y0 - (Lmax * B) / (K + B)

start_vals_pop_interrow <- list(
  Y0   = max(dat_pop_interrow$Y, na.rm = TRUE),
  Lmax = max(dat_pop_interrow$Y, na.rm = TRUE) -
         min(dat_pop_interrow$Y, na.rm = TRUE),
  K    = median(dat_pop_interrow$B[dat_pop_interrow$B > 0], na.rm = TRUE)
)

hyper_mod_pop_interrow <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_pop_interrow,
  start = start_vals_pop_interrow
)

summary(hyper_mod_pop_interrow)

# 3) AIC and R2 comparison ---------------------------------------------

aic_lin_pop_interrow   <- AIC(lin_mod_pop_interrow)
aic_hyper_pop_interrow <- AIC(hyper_mod_pop_interrow)

gl_lin_pop_interrow <- broom::glance(lin_mod_pop_interrow)
R2_lin_pop_interrow <- gl_lin_pop_interrow$r.squared

get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res
  
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper_pop_interrow <- get_r2_nls(hyper_mod_pop_interrow)

df_lin_pop_interrow   <- df.residual(lin_mod_pop_interrow)
df_hyper_pop_interrow <- df.residual(hyper_mod_pop_interrow)

mod_comp_pop_interrow <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin_pop_interrow, df_hyper_pop_interrow),
  AIC   = c(aic_lin_pop_interrow, aic_hyper_pop_interrow),
  R2    = c(R2_lin_pop_interrow, R2_hyper_pop_interrow)
)

mod_comp_pop_interrow |>
  kable(
    digits  = 3,
    caption = "Model comparison for bean population vs interrow weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Diagnostics
```{r}
# --- Bean population vs interrow weed biomass -------------------------

# Mixed model diagnostics (chosen population model)
check_model(pop_interrow_mod_best)

# Simple linear model diagnostics (population)
check_model(lin_mod_pop_interrow)

# Optional specific checks for the population linear model
check_heteroscedasticity(lin_mod_pop_interrow)
check_normality(lin_mod_pop_interrow)
check_outliers(lin_mod_pop_interrow)

```

#### Visualisation
```{r}
ggplot(dat_pop_interrow, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: black, not in legend
  stat_function(
    fun = function(x) coef(lin_mod_pop_interrow)[1] +
                      coef(lin_mod_pop_interrow)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod_pop_interrow)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  scale_y_continuous(
    labels = scales::label_comma(accuracy = 1)
  ) +
  guides(
    color = guide_legend(
      override.aes = list(
        linetype = c("dashed", "dotdash"),
        shape    = c(NA, NA),
        alpha    = c(1, 1)
      )
    )
  ) +
  labs(
    x = bquote("Interrow weed biomass (kg ha"^-1*")"),
    y = bquote("Bean population (plants ha"^-1*")"),
    title = "Bean population vs interrow weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )

```
### Seed weight
#### Dataset
```{r}
# ----------------------------------------------------------------------
# Model testing: interrow weed biomass vs 100-seed weight (g)
# ----------------------------------------------------------------------

## Exploratory: regression dataset -------------------------------------

reg_100sw_interrow <- weed_biomass_clean |>
  filter(
    !is.na(interrow_weed_biomass_kg_ha),
    !is.na(seed_weight),   # 100-seed weight column
    !is.na(site_year),
    !is.na(block)
  )

# Quick structure check (optional)
str(reg_100sw_interrow[, c("site_year", "block",
                           "interrow_weed_biomass_kg_ha",
                           "seed_weight")])

```

#### Mixed model selection (site-year effects)
```{r}
### Mixed model selection (site-year effects): 100-seed weight ----------

options(contrasts = c("contr.sum", "contr.poly"))

# 1) Candidate mixed models --------------------------------------------

# Additive: common slope across site-year
sw100_interrow_add <- lmer(
  seed_weight ~ interrow_weed_biomass_kg_ha + site_year +
    (1 | site_year:block),
  data = reg_100sw_interrow
)

# Interaction: slope allowed to differ by site-year
sw100_interrow_int <- lmer(
  seed_weight ~ interrow_weed_biomass_kg_ha * site_year +
    (1 | site_year:block),
  data = reg_100sw_interrow
)

# 2) Likelihood ratio test ---------------------------------------------

lrt_sw100_interrow   <- anova(sw100_interrow_add, sw100_interrow_int)
p_int_sw100_interrow <- lrt_sw100_interrow$`Pr(>Chisq)`[2]

# 3) AIC table with LRT p-value only for interaction model -------------

aic_sw100_interrow <- tibble(
  model = c(
    "Additive: interrow_weed_biomass_kg_ha + site_year",
    "Interaction: interrow_weed_biomass_kg_ha * site_year"
  ),
  AIC = c(AIC(sw100_interrow_add), AIC(sw100_interrow_int))
) |>
  mutate(
    deltaAIC = AIC - min(AIC),
    `LRT p (interaction vs additive)` = c(NA_real_, p_int_sw100_interrow)
  )

# Selection rule: prefer simpler model unless interaction clearly better
keep_interaction_sw100_interrow <-
  (p_int_sw100_interrow < 0.05) &&
  (aic_sw100_interrow$deltaAIC[
    aic_sw100_interrow$model ==
      "Interaction: interrow_weed_biomass_kg_ha * site_year"
  ] < 2)

sw100_interrow_mod_best <- if (keep_interaction_sw100_interrow)
  sw100_interrow_int else sw100_interrow_add

chosen_model_name_sw100_interrow <- if (keep_interaction_sw100_interrow) {
  "Interaction: interrow_weed_biomass_kg_ha * site_year"
} else {
  "Additive: interrow_weed_biomass_kg_ha + site_year"
}

aic_sw100_interrow_out <- aic_sw100_interrow |>
  mutate(
    Selected = if_else(model == chosen_model_name_sw100_interrow, "Yes", "")
  )

aic_sw100_interrow_out |>
  kable(
    digits  = 2,
    caption = "100-seed weight vs interrow weed biomass: model comparison (additive vs interaction). LRT p-value compares the interaction model vs the additive model."
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

cat(
  "\nSelected model for 100-seed weight vs interrow weed biomass:\n  ",
  chosen_model_name_sw100_interrow,
  sprintf("  [LRT p (interaction) = %.3f]\n", p_int_sw100_interrow)
)

```


#### Slope summary via emtrends
```{r}
### Slope summary: 100-seed weight vs interrow weed biomass ------------

if (chosen_model_name_sw100_interrow ==
    "Additive: interrow_weed_biomass_kg_ha + site_year") {
  
  # One common slope across site-years
  emm_slope_sw100_interrow <- emtrends(
    sw100_interrow_mod_best, ~ 1,
    var = "interrow_weed_biomass_kg_ha"
  )
  
  emm_slope_sw100_interrow_df <- summary(
    emm_slope_sw100_interrow,
    infer = c(TRUE, TRUE)
  ) |>
    as_tibble()
  
  trend_col_sw100_interrow <- get_trend_col(
    emm_slope_sw100_interrow_df,
    "interrow_weed_biomass_kg_ha"
  )
  stat_col_sw100_interrow  <- get_stat_col(emm_slope_sw100_interrow_df)
  
  has_df_sw100_interrow <- "df" %in% names(emm_slope_sw100_interrow_df)
  has_p_sw100_interrow  <- "p.value" %in% names(emm_slope_sw100_interrow_df)
  
  slope_tbl_sw100_interrow <- emm_slope_sw100_interrow_df |>
    transmute(
      Scope                         = "Pooled across site-years",
      slope_g_per_kgweed            = .data[[trend_col_sw100_interrow]],
      slope_g_per_100kgweed         = .data[[trend_col_sw100_interrow]] * 100,
      SE,
      df         = if (has_df_sw100_interrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_sw100_interrow))
                     .data[[stat_col_sw100_interrow]] else NA_real_,
      p_value    = if (has_p_sw100_interrow)
                     .data[["p.value"]] else NA_real_
    )
  
} else {
  
  # Slopes vary by site-year
  emm_slope_sw100_interrow <- emtrends(
    sw100_interrow_mod_best, ~ site_year,
    var = "interrow_weed_biomass_kg_ha"
  )
  
  emm_slope_sw100_interrow_df <- summary(
    emm_slope_sw100_interrow,
    infer = c(TRUE, TRUE)
  ) |>
    as_tibble()
  
  trend_col_sw100_interrow <- get_trend_col(
    emm_slope_sw100_interrow_df,
    "interrow_weed_biomass_kg_ha"
  )
  stat_col_sw100_interrow  <- get_stat_col(emm_slope_sw100_interrow_df)
  
  has_df_sw100_interrow <- "df" %in% names(emm_slope_sw100_interrow_df)
  has_p_sw100_interrow  <- "p.value" %in% names(emm_slope_sw100_interrow_df)
  
  slope_tbl_sw100_interrow <- emm_slope_sw100_interrow_df |>
    transmute(
      Scope                         = as.character(site_year),
      slope_g_per_kgweed            = .data[[trend_col_sw100_interrow]],
      slope_g_per_100kgweed         = .data[[trend_col_sw100_interrow]] * 100,
      SE,
      df         = if (has_df_sw100_interrow) df else NA_real_,
      lower_95CL = lower.CL,
      upper_95CL = upper.CL,
      stat       = if (!is.na(stat_col_sw100_interrow))
                     .data[[stat_col_sw100_interrow]] else NA_real_,
      p_value    = if (has_p_sw100_interrow)
                     .data[["p.value"]] else NA_real_
    )
}

slope_tbl_sw100_interrow |>
  mutate(
    across(
      c(slope_g_per_kgweed,
        slope_g_per_100kgweed,
        SE, lower_95CL, upper_95CL),
      ~ round(.x, 3)
    )
  ) |>
  kable(
    caption = "Slopes of 100-seed weight vs interrow weed biomass (kg ha^-1): mixed-model emtrends",
    col.names = c(
      "Scope (site-year / pooled)",
      "Slope (g per kg interrow weed ha^-1)",
      "Slope (g per 100 kg interrow weed ha^-1)",
      "SE", "df",
      "Lower 95% CL", "Upper 95% CL",
      "Test statistic", "p-value"
    )
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Linear vs rectangular hyperbolic response
```{r}
## Linear vs rectangular hyperbolic response of 100-seed weight 
## to interrow weed biomass --------------------------------------------

dat_sw100_interrow <- reg_100sw_interrow |>
  select(interrow_weed_biomass_kg_ha, seed_weight) |>
  rename(
    B = interrow_weed_biomass_kg_ha,
    Y = seed_weight
  )

names(dat_sw100_interrow)  # "B", "Y"

# 1) Linear model ------------------------------------------------------

lin_mod_sw100_interrow <- lm(Y ~ B, data = dat_sw100_interrow)

# 2) Rectangular hyperbolic model -------------------------------------

start_vals_sw100_interrow <- list(
  Y0   = max(dat_sw100_interrow$Y, na.rm = TRUE),
  Lmax = max(dat_sw100_interrow$Y, na.rm = TRUE) -
         min(dat_sw100_interrow$Y, na.rm = TRUE),
  K    = median(dat_sw100_interrow$B[dat_sw100_interrow$B > 0], na.rm = TRUE)
)

hyper_mod_sw100_interrow <- nls(
  Y ~ Y0 - (Lmax * B) / (K + B),
  data  = dat_sw100_interrow,
  start = start_vals_sw100_interrow
)

summary(hyper_mod_sw100_interrow)

# 3) AIC and R2 comparison ---------------------------------------------

aic_lin_sw100_interrow   <- AIC(lin_mod_sw100_interrow)
aic_hyper_sw100_interrow <- AIC(hyper_mod_sw100_interrow)

gl_lin_sw100_interrow <- broom::glance(lin_mod_sw100_interrow)
R2_lin_sw100_interrow <- gl_lin_sw100_interrow$r.squared

get_r2_nls <- function(mod) {
  y_hat <- fitted(mod)
  res   <- resid(mod)
  y     <- y_hat + res
  
  rss <- sum(res^2)
  tss <- sum((y - mean(y))^2)
  1 - rss / tss
}

R2_hyper_sw100_interrow <- get_r2_nls(hyper_mod_sw100_interrow)

df_lin_sw100_interrow   <- df.residual(lin_mod_sw100_interrow)
df_hyper_sw100_interrow <- df.residual(hyper_mod_sw100_interrow)

mod_comp_sw100_interrow <- tibble(
  model = c("Linear", "Hyperbolic"),
  df    = c(df_lin_sw100_interrow, df_hyper_sw100_interrow),
  AIC   = c(aic_lin_sw100_interrow, aic_hyper_sw100_interrow),
  R2    = c(R2_lin_sw100_interrow, R2_hyper_sw100_interrow)
)

mod_comp_sw100_interrow |>
  kable(
    digits  = 3,
    caption = "Model comparison for 100-seed weight vs interrow weed biomass: linear vs rectangular hyperbolic"
  ) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

```

#### Diagnostics
```{r}
# --- 100-seed weight vs interrow weed biomass ------------------------

# Mixed model diagnostics (chosen 100-seed weight model)
check_model(sw100_interrow_mod_best)

# Simple linear model diagnostics (100-seed weight)
check_model(lin_mod_sw100_interrow)

# Optional specific checks for the 100-seed weight linear model
check_heteroscedasticity(lin_mod_sw100_interrow)
check_normality(lin_mod_sw100_interrow)
check_outliers(lin_mod_sw100_interrow)

```

#### Visualisation
```{r}
ggplot(dat_sw100_interrow, aes(x = B, y = Y)) +
  # points: always black
  geom_point(
    color = "black",
    alpha = 1
  ) +
  # LOESS: uses palette color and appears in legend
  geom_smooth(
    aes(color = "LOESS"),
    method = "loess",
    se = FALSE,
    linewidth = 1,
    linetype = "dashed"
  ) +
  # Linear: always black, not tied to color scale
  stat_function(
    fun = function(x) coef(lin_mod_sw100_interrow)[1] +
                      coef(lin_mod_sw100_interrow)[2] * x,
    linewidth = 1,
    color = "black"
  ) +
  # Hyperbolic: uses palette color and appears in legend
  stat_function(
    aes(color = "Hyperbolic"),
    fun = function(x) {
      pars <- coef(hyper_mod_sw100_interrow)
      pars["Y0"] - (pars["Lmax"] * x) / (pars["K"] + x)
    },
    linewidth = 1,
    linetype = "dotdash"
  ) +
  scale_color_manual(
    name   = NULL,
    breaks = c("LOESS", "Hyperbolic"),
    labels = c(
      "LOESS smooth",
      "Hyperbolic fit"
    ),
    values = c(
      "LOESS"      = line_cols[2],
      "Hyperbolic" = line_cols[4]
    )
  ) +
  labs(
    x = bquote("Interrow weed biomass (kg ha"^-1*")"),
    y = "100-seed weight (g)",
    title = "100-seed weight vs interrow weed biomass",
    subtitle = "Data (black), LOESS, and linear vs hyperbolic fits"
  ) +
  theme_classic() +
  theme(
    legend.position = "right"
  )

```

