---
title: "Bean Yield"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries


```{r message=FALSE, warning=FALSE}
#Set work directory
setwd("/Users/ey239/Github/Mowtivation/rmarkdowns")

#Load packages 
library(tidyverse) ##install.packages("tidyverse")
library(knitr)
library(patchwork) ##install.packages("patchwork")
library(skimr)     ##install.packages("skimr")
library(readxl)
library(janitor) ##install.packages("janitor")
library(kableExtra) ##install.packages("kableExtra")
library(webshot) ##install.packages("webshot")
webshot::install_phantomjs()
library(viridis) ##install.packages("viridis")
library(lme4) ##install.packages("lme4")
library(lmerTest) ##install.packages("lmerTest")
library(emmeans) ##install.packages("emmeans")
library(rstatix) ##install.packages("rstatix")
#library(Matrix) ##install.packages("Matrix")
library(multcomp) ##install.packages("multcomp")
library(multcompView) ##install.packages("multcompView")
library(ggResidpanel) ##install.packages("ggResidpanel")
#library(car)
#library(TMB)  ##install.packages("TMB")
#library(glmmTMB)  ##install.packages("glmmTMB")
library(DHARMa)  ##install.packages("DHARMa")
library(performance) ##install.packages("performance")
#Load Functions
MeanPlusSe<-function(x) mean(x)+plotrix::std.error(x)

find_logw0=function(x){c=trunc(log(min(x[x>0],na.rm=T)))
d=exp(c)
return(d)}
```

<br>

# Load and clean data

 
## Load data
``` {r}
combined_raw <- read_excel("~/Github/Mowtivation/raw-data/All Treatments/combined_raw.xlsx")
kable(head(combined_raw))

```

<br>

## Clean data
```{r}


#Standardaze column names, convert to factors, check for outliers of variable**
clean_combined <- clean_names(combined_raw) |>  
  rename ('weed_control'= treatment) |> 
  mutate(across(c(weed_control, block, plot, location, year), as.factor)) #|> 
  #mutate(is_outlier = totwbm < (quantile(totwbm, 0.25) - 1.5 * IQR(totwbm)) |
                       #wbm > (quantile(totwbm, 0.75) + 1.5 * IQR(totwbm)))

#select and convert data for wbm analysis
  bean_yield_clean <- clean_combined |>  
    filter(!is.na(bean_yield)) |>
  mutate(bean_yield = as.numeric(bean_yield)) |>  # Convert beanyd to numeric
    # Exclude rows with NA in beanyd
  mutate(
    bean_yield_adj_bu_acre = (((bean_yield / 454) / (16.4 / 43560)) / 60) * ((100 - 0.00001) / (100 - 14)),
    bean_yield_adj_lbs_acre = ((bean_yield / 454) / (16.4 / 43560)) * ((100 - 0.00001) / (100 - 14)),
    bean_yield_adj_kg_ha = ((bean_yield / 454) / (16.4 / 43560)) * 1.12085 * ((100 - 0.00001) / (100 - 14))
  )
kable(head(bean_yield_clean)) 
```
<br>


# Model testing
## Lmer
Block is random
Tyler is under the impression that block should always be random and that post-hoc comparisons should use TUKEY rather the Fischer. Fisher is bogus apparently. 

```{r}
random <- lmer( bean_yield_adj_kg_ha  ~ weed_control + (1|location) + (1|location:block) , data =  bean_yield_clean)

random <- lmer( bean_yield_adj_kg_ha  ~ weed_control*location + (1|location:block) , data =  bean_yield_clean)


resid_panel(random)

simulateResiduals(random,plot = TRUE) # Residuals and normality look good
check_model(random)
```


<br>

## Joint test (anova)
```{r}
 random |> 
  joint_tests() |> 
  kable()  
```

<br>

## Means comparison
### Weed-control (Not significant)
```{r message=FALSE}
means_weed_control <- emmeans(random, ~  weed_control)
pairwise_comparisons_weed_control<- pairs(means_weed_control) 
kable(head(pairwise_comparisons_weed_control))
```


<br>

## Tukey compact letter display
### Weed-control (Not significant)
```{r}

cld_weed_control_tukey <-cld(emmeans(random, ~  weed_control, type = "response"), Letters = letters, sort = TRUE, reversed=TRUE)
cld_weed_control_tukey
```
<br>


## Fisher compact letter display
### Weed-control (No significant)

```{r}
cld_weed_control_fisher <-cld(emmeans(random, ~  weed_control , type = "response"), Letters = letters, adjust = "none",sort = TRUE, reversed=TRUE)
cld_weed_control_fisher
```

# Figures using fisher
###Bu/a
#### Weed-control (NS)

```{r message=FALSE, warning=FALSE}
bean_yield_clean |> 
  left_join(cld_weed_control_fisher) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")), y = bean_yield_adj_bu_acre, fill = weed_control)) +
  stat_summary(geom = "bar", fun = "mean", width = 0.7) +
  stat_summary(geom = "errorbar", fun.data = "mean_se", width = 0.2) +
  #stat_summary(geom="text", fun = "MeanPlusSe", aes(label= trimws(.group)),size=6.5,vjust=-0.5) +
  labs(
    x = "",
    y = expression(paste("Soybean yield (", bu, "/", a, " at 13% moisture)")),
    #title = str_c("The influence of interrow weed control on soybean yield"),
    subtitle = expression(italic("Not signficant"))) +
  
  scale_x_discrete(labels = c("Rolled,\nno additional\nweed control",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nhigh-residue\ncultivation",
                              "Tilled,\ninterrow\nmowing",
                          "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12),
    axis.title = element_text(size = 24),  # Increase font size of axis titles
    axis.text = element_text(size = 24),   # Increase font size of axis labels
    plot.title = element_text(size = 24, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 24, face = "italic")  # Increase font size of subtitle
  )
ggsave("bean_yield_weed_control_bua.png", width = 10, height = 8, dpi = 300)
```



###Kg/hectare
#### Weed-control (NS)

```{r message=FALSE, warning=FALSE}
bean_yield_clean |> 
  left_join(cld_weed_control_fisher) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")), y =  bean_yield_adj_kg_ha, fill = weed_control)) +
  stat_summary(geom = "bar", fun = "mean", width = 0.7) +
  stat_summary(geom = "errorbar", fun.data = "mean_se", width = 0.2) +
  #stat_summary(geom="text", fun = "MeanPlusSe", aes(label= trimws(.group)),size=6.5,vjust=-0.5) +
  labs(
    x = "",
    y = expression(paste("Soybean yield (", kg~ha^{-1}, " at 13% moisture)")),

    #title = str_c("The influence of interrow weed control on soybean yield"),
    subtitle = expression(italic("Not signficant"))) +
  
  scale_x_discrete(labels = c("Rolled,\nno additional\nweed control",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nhigh-residue\ncultivation",
                              "Tilled,\ninterrow\nmowing",
                          "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12),
    axis.title = element_text(size = 24),  # Increase font size of axis titles
    axis.text = element_text(size = 24),   # Increase font size of axis labels
    plot.title = element_text(size = 24, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 24, face = "italic")  # Increase font size of subtitle
  )
ggsave("bean_yield_weed_control_kgha.png", width = 10, height = 8, dpi = 300)
```