---
title: "Bean Yield"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Load libraries**


```{r message=FALSE, warning=FALSE}
#Set work directory
setwd("/Users/ey239/Github/Mowtivation/rmarkdowns")

#Load packages 
library(tidyverse) ##install.packages("tidyverse")
library(knitr)
library(patchwork) ##install.packages("patchwork")
library(skimr)     ##install.packages("skimr")
library(readxl)
library(janitor) ##install.packages("janitor")
library(kableExtra) ##install.packages("kableExtra")
library(webshot) ##install.packages("webshot")
webshot::install_phantomjs()
library(viridis) ##install.packages("viridis")
library(lme4) ##install.packages("lme4")
library(lmerTest) ##install.packages("lmerTest")
library(emmeans) ##install.packages("emmeans")
library(rstatix) ##install.packages("rstatix")
#library(Matrix) ##install.packages("Matrix")
library(multcomp) ##install.packages("multcomp")
library(multcompView) ##install.packages("multcompView")
library(ggResidpanel) ##install.packages("ggResidpanel")
#library(car)
#library(TMB)  ##install.packages("TMB")
#library(glmmTMB)  ##install.packages("glmmTMB")
#library(DHARMa)  ##install.packages("DHARMa")

#Load Functions
MeanPlusSe<-function(x) mean(x)+plotrix::std.error(x)

find_logw0=function(x){c=trunc(log(min(x[x>0],na.rm=T)))
d=exp(c)
return(d)}
```

<br>

# **Load and Clean Data**

### **Load individual datasets**

``` {r}
combined_raw <- read_excel("~/Github/Mowtivation/raw-data/All Treatments/combined_raw.xlsx")
kable(head(combined_raw))

```



```{r}


#Standardaze column names, convert to factors, check for outliers of variable**
clean_combined <- clean_names(combined_raw) |>  
  rename ('weed_control'= treatment) |> 
  mutate(across(c(weed_control, block, plot, location, year), as.factor)) #|> 
  #mutate(is_outlier = totwbm < (quantile(totwbm, 0.25) - 1.5 * IQR(totwbm)) |
                       #wbm > (quantile(totwbm, 0.75) + 1.5 * IQR(totwbm)))

#select and convert data for wbm analysis
  bean_yield_clean <- clean_combined |>  
    filter(!is.na(bean_yield)) |>
  mutate(bean_yield = as.numeric(bean_yield)) |>  # Convert beanyd to numeric
    # Exclude rows with NA in beanyd
  mutate(
    bean_yield_adj_bu_acre = (((bean_yield / 454) / (16.4 / 43560)) / 60) * ((100 - 0.00001) / (100 - 14)),
    bean_yield_adj_lbs_acre = ((bean_yield / 454) / (16.4 / 43560)) * ((100 - 0.00001) / (100 - 14)),
    bean_yield_adj_kg_ha = ((bean_yield / 454) / (16.4 / 43560)) * 1.12085 * ((100 - 0.00001) / (100 - 14))
  )
kable(head(bean_yield_clean)) 
```
<br>


## **Model testing**

### **block is random**
#Ask tyler about model format, should block always be fixed, etc. should location be nested in year? Do i need to transform data if zeros are present? How to separate year from location? 

Last meeting we updated the model and determined that for most of mhy work, block should be random using model below and that post hoc comparisons should use TUKEY ratther the Fischer. 

```{r}
random <- lmer( bean_yield_adj_bu_acre  ~ location+weed_control + location:weed_control +(1|location:block) , data =  bean_yield_clean)

resid_panel(random)
```

<br>

##**Joint test**
```{r}
 random |> 
  joint_tests() |> 
  kable()  
```

<br>

# **Means comparison**
```{r message=FALSE}
means <- emmeans(random, ~  weed_control)
# Optional: Adjust for multiple comparisons (e.g., using Tukey's method)

pairwise_comparisons<- pairs(means) 
kable(head(pairwise_comparisons))
```

### **Fisher's method for comparing means**

```{r}
#weed_control
cld_weed_control_tukey <-cld(emmeans(random, ~  weed_control, type = "response"), Letters = letters, sort = TRUE, reversed=TRUE)
cld_weed_control_tukey
```


```{r}
#location
cld_location_tukey <-cld(emmeans(random, ~  location, type = "response"), Letters = letters, sort = TRUE, reversed=TRUE)
cld_location_tukey
```
```{r}
#weed_control|location
cld_weed_control_location_tukey <-cld(emmeans(random, ~  weed_control|location, type = "response"), Letters = letters, sort = TRUE, reversed=TRUE)
cld_weed_control_location_tukey
```
# **FIGURES** only location was significant

## **weed_control**

```{r message=FALSE}
bean_yield_clean |> 
  left_join(cld_weed_control_tukey) |> 
  ggplot(aes(x = weed_control, y = bean_yield_adj_bu_acre, fill = weed_control)) +
  stat_summary(geom = "bar", fun = "mean", width = 0.7) +
  stat_summary(geom = "errorbar", fun.data = "mean_se", width = 0.2) +
  #stat_summary(geom="text", fun = "MeanPlusSe", aes(label= trimws(.group)),size=6.5,vjust=-0.5) +
  labs(
    x = "Method of interrow weed control",
    y = expression("Soybean yield" ~ (bu ~ acre^{-1})),
    title = str_c("The influence of the method of interrow weed control on soybean yield"),
    subtitle = expression(italic("Not signficant"))) +
  
  scale_x_discrete(labels = c("Rolled,\nhigh-residue\ncultivation",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nno additional\nweed control",
                          "Tilled,\nstandard\ncultivation",
                              "Tilled,\ninterrow\nmowing")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12)
  )
ggsave("beanydall_plot_cutivation.png", width = 8, height = 6, dpi = 300)
```

## **location**

```{r message=FALSE}
bean_yield_clean |> 
  left_join(cld_location_tukey) |> 
  ggplot(aes(x = location, y = bean_yield_adj_bu_acre, fill = location)) +
  stat_summary(geom = "bar", fun = "mean", width = 0.7) +
  stat_summary(geom = "errorbar", fun.data = "mean_se", width = 0.2) +
  #stat_summary(geom="text", fun = "MeanPlusSe", aes(label= trimws(.group)),size=6.5,vjust=-0.5) +
  labs(
    x = "Location",
    y = expression("Soybean yield" ~ (bu ~ A^{-1})),
    title = str_c("The influence of location on soybean yield"),
    subtitle = expression(italic("P < 0.005"))) +
   scale_x_discrete(labels = c("Field O2 East ",
                              "Field O2 West",
                              "Field X")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12)
  )
ggsave("beanydall_location.png", width = 8, height = 6, dpi = 300)
```