---
title: "Soybean biomass"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r message=FALSE, warning=FALSE}
#Set work directory
setwd("/Users/ey239/Github/Mowtivation/rmarkdowns")

#Load packages 
library(tidyverse) ##install.packages("tidyverse")
library(knitr)
library(patchwork) ##install.packages("patchwork")
library(skimr)     ##install.packages("skimr")
library(readxl)
library(janitor) ##install.packages("janitor")

library(kableExtra) ##install.packages("kableExtra")
library(webshot) ##install.packages("webshot")
webshot::install_phantomjs()
library(viridis) ##install.packages("viridis")
library(lme4) ##install.packages("lme4")
library(lmerTest) ##install.packages("lmerTest")
library(emmeans) ##install.packages("emmeans")
library(rstatix) ##install.packages("rstatix")
#library(Matrix) ##install.packages("Matrix")
library(multcomp) ##install.packages("multcomp")
library(multcompView) ##install.packages("multcompView")
library(ggResidpanel) ##install.packages("ggResidpanel")
#library(car)
#library(TMB)  ##install.packages("TMB")
#library(glmmTMB)  ##install.packages("glmmTMB")
library(DHARMa)  ##install.packages("DHARMa")
library(performance) 
#Load Functions
MeanPlusSe<-function(x) mean(x)+plotrix::std.error(x)

find_logw0=function(x){c=trunc(log(min(x[x>0],na.rm=T)))
d=exp(c)
return(d)}
```

<br>

# Load and clean data

## Load data

``` {r}
combined_raw <- read_excel("~/Github/Mowtivation/raw-data/All Treatments/combined_raw.xlsx")
kable(head(combined_raw))

```

#Clean data

```{r}


#Standardaze column names, convert to factors, check for outliers of variable**
clean_combined <- clean_names(combined_raw) |>  
  rename ('weed_control'= treatment) |> 
  mutate(across(c(weed_control, block, plot, location, year), as.factor)) #|> 
  #mutate(is_outlier = totwbm < (quantile(totwbm, 0.25) - 1.5 * IQR(totwbm)) |
                       #wbm > (quantile(totwbm, 0.75) + 1.5 * IQR(totwbm)))

#select and convert data for wbm analysis
bean_biomass_clean <-clean_combined |>              
  mutate(bean_biomass_grams_meter = (bean_biomass * 2)) |> 
  mutate(bean_biomass_kg_ha = ((bean_biomass/0.5) *(10000))/(1000)) |>
  mutate(bean_biomass_lbs_ac = (((bean_biomass/0.5) *(10000))/(1000))* 0.892179)
kable(head(bean_biomass_clean)) 
```
<br>
# Model testing

Block is random
Tyler is under the impression that block should always be random and that post-hoc comparisons should use TUKEY rather the Fischer. Fisher is bogus apparently. 

## Lmer
```{r}
biomass.lmer <- lmer(bean_biomass_kg_ha  ~ weed_control*location + (1|location:block) , data =bean_biomass_clean)

resid_panel(biomass.lmer)
simulateResiduals(biomass.lmer ,plot = TRUE) # Residuals and normality look good
check_model(biomass.lmer ) #Perfect, preditions match real data

```


## Joint test (anova)
```{r}

joint_tests(emmeans(biomass.lmer, ~ weed_control * location))
```
```{r}
options(contrasts = c("contr.sum", "contr.poly"))
biomass.anova <- anova(biomass.lmer, ddf = "Satterthwaite")
print(biomass.anova)
```
<br>

## Tukey means comparisons
### Weed-control (significant)

```{r}
tukey_weed_control <- emmeans(biomass.lmer, pairwise ~ weed_control, adjust = "tukey")
tukey_weed_control

cld_weed_control_tukey <- cld(emmeans(biomass.lmer, ~ weed_control), adjust = "tukey", Letters = letters, sort = TRUE, reversed = TRUE)

cld_weed_control_tukey
```


<br>

### Location (Significant)
```{r}
tukey_location <- emmeans(biomass.lmer, list(pairwise ~ location), adjust = "tukey")
tukey_location

cld_location_tukey <- cld(emmeans(biomass.lmer, ~ location), adjust = "tukey", Letters = letters, sort = TRUE, reversed = TRUE)

cld_location_tukey
```

###weed_control:location (not significant)

```{r}
tukey_weed_control_location <- emmeans(biomass.lmer, list(pairwise ~ weed_control|location), adjust = "tukey")
tukey_weed_control_location

cld_weed_control_location_tukey <- cld(emmeans(biomass.lmer, ~ weed_control|location), adjust = "tukey", Letters = letters, sort = TRUE, reversed = TRUE)

cld_weed_control_location_tukey
```





#

# Figures using Tukey
## Weed-control (Significant)

```{r message=FALSE}
bean_biomass_clean |> 
  left_join(cld_weed_control_tukey) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")),, y = bean_biomass_kg_ha, fill = weed_control)) +
  stat_summary(geom = "bar", fun = "mean", width = 0.7) +
  stat_summary(geom = "errorbar", fun.data = "mean_se", width = 0.2) +
  stat_summary(geom="text", fun = "MeanPlusSe", aes(label= trimws(.group)),size=6.5,vjust=-0.5) +
  labs(
    x = "",
    y = expression(paste("Soybean biomass (", kg~ha^{-1}, ")")),
    title = str_c(""),
    subtitle = expression(italic("P < 0.001"))) +
  
  scale_x_discrete(labels =  c("Rolled,\nno additional\nweed control",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nhigh-residue\ncultivation",
                              "Tilled,\ninterrow\nmowing",
                          "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12),
    axis.title = element_text(size = 20),  # Increase font size of axis titles
    axis.text = element_text(size = 16),   # Increase font size of axis labels
    plot.title = element_text(size = 20, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 16, face = "italic")  # Increase font size of subtitle
  )
  
ggsave("bean_biomass_weed_control_tukey_lbA.png", width = 10, height = 6, dpi = 300)
```

## Location (Significant)
```{r message=FALSE}
bean_biomass_clean |> 
  left_join(cld_location_tukey) |> 
  ggplot(aes(x = location, y = bean_biomass_kg_ha, fill = location)) +
  stat_summary(geom = "bar", fun = "mean", width = 0.7) +
  stat_summary(geom = "errorbar", fun.data = "mean_se", width = 0.2) +
  stat_summary(geom="text", fun = "MeanPlusSe", aes(label= trimws(.group)),size=6.5,vjust=-0.5) +
  labs(
    x = "",
    y = expression(paste("Soybean biomass (", kg~ha^{-1}, ")")),
    title = str_c(""),
    subtitle = expression(italic("P < 0.001"))) +
  
   scale_x_discrete(labels = c("Field O2 East ",
                              "Field O2 West",
                              "Field X")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12),
    axis.title = element_text(size = 20),  # Increase font size of axis titles
    axis.text = element_text(size = 16),   # Increase font size of axis labels
    plot.title = element_text(size = 20, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 16, face = "italic")  # Increase font size of subtitle
  
  )
ggsave("bean_biomass_location_tukey_lbA.png", width = 12, height = 6, dpi = 300)
```

###Kg/hectare
#### weed_control:location (not significant)

```{r message=FALSE, warning=FALSE}
bean_biomass_clean |> 
  left_join(cld_weed_control_location_tukey ) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")), y =  bean_biomass_kg_ha, fill = weed_control)) +
  facet_wrap( ~location, labeller = labeller(
    location = c("field O2 east" = "Field O2 East", "field O2 west" = "Field O2 West","field v" = "Field V" )))+
  stat_summary(geom = "bar", fun = "mean", width = 0.7) +
  stat_summary(geom = "errorbar", fun.data = "mean_se", width = 0.2) +
  #stat_summary(geom="text", fun = "MeanPlusSe", aes(label= trimws(.group)),size=6.5,vjust=-0.5) +
  labs(
    x = "",
    y = expression(paste("Soybean biomass (", kg~ha^{-1}, ")")),

    #title = str_c("The influence of interrow weed control on soybean yield"),
    subtitle = expression(italic("Not significant"))) +
  
  scale_x_discrete(labels = c("Rolled,\nno additional\nweed control",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nhigh-residue\ncultivation",
                              "Tilled,\ninterrow\nmowing",
                          "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 20),
    axis.title = element_text(size = 20),  # Increase font size of axis titles
    axis.text = element_text(size = 20),   # Increase font size of axis labels
    plot.title = element_text(size = 24, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 24, face = "italic")  # Increase font size of subtitle
  )
ggsave("bean_biomass_weed_control_location_kgha.png", width = 24, height = 8, dpi = 300)
```


