---
title: "interrow weed biomass"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r message=FALSE, warning=FALSE}
#Set work directory
setwd("/Users/ey239/Github/Mowtivation/rmarkdowns")

#Load packages 
library(tidyverse) ##install.packages("tidyverse")
library(knitr)
library(patchwork) ##install.packages("patchwork")
library(skimr)     ##install.packages("skimr")
library(readxl)
library(janitor) ##install.packages("janitor")

library(kableExtra) ##install.packages("kableExtra")
library(webshot) ##install.packages("webshot")
webshot::install_phantomjs()
library(viridis) ##install.packages("viridis")
library(lme4) ##install.packages("lme4")
library(lmerTest) ##install.packages("lmerTest")
library(emmeans) ##install.packages("emmeans")
library(rstatix) ##install.packages("rstatix")
#library(Matrix) ##install.packages("Matrix")
library(multcomp) ##install.packages("multcomp")
library(multcompView) ##install.packages("multcompView")
library(ggResidpanel) ##install.packages("ggResidpanel")
#library(car)
#library(TMB)  ##install.packages("TMB")
library(glmmTMB)  ##install.packages("glmmTMB")
library(DHARMa)  ##install.packages("DHARMa")
library(performance) ##install.packages("performance")

#Load Functions
MeanPlusSe<-function(x) mean(x)+plotrix::std.error(x)

find_logw0=function(x){c=trunc(log(min(x[x>0],na.rm=T)))
d=exp(c)
return(d)}
```
<br>

# Load and Clean Data

## Load data

``` {r}
combined_raw <- read_excel("~/Github/Mowtivation/raw-data/All Treatments/combined_raw.xlsx")
kable(head(combined_raw))

```

## Clean data
```{r}


#Standardaze column names, convert to factors, check for outliers of variable**
clean_combined <- clean_names(combined_raw) |>  
  rename ('weed_control'= treatment) |> 
  mutate(across(c(weed_control, block, plot, location, year), as.factor)) #|> 
  #mutate(is_outlier = totwbm < (quantile(totwbm, 0.25) - 1.5 * IQR(totwbm)) |
                       #wbm > (quantile(totwbm, 0.75) + 1.5 * IQR(totwbm)))

#select and convert data for wbm analysis
interrow_weed_biomass_clean <-clean_combined |>              
  mutate(log_interrow_weed_biomass_grams_meter=  (log((interrow_weed_biomass*2)+1)))|>
  mutate(interrow_weed_biomass_grams_meter = (interrow_weed_biomass /0.4026))|> # 
 
  mutate(interrow_weed_biomass_kg_ha = (interrow_weed_biomass_grams_meter *(10000))/(1000)) |>
  mutate(interrow_weed_biomass_lbs_ac = ((interrow_weed_biomass_grams_meter *(10000))/(1000))* 0.892179)
kable(head(interrow_weed_biomass_clean)) 

# Model testing
```
Block is random
Tyler is under the impression that block should always be random and that post-hoc comparisons should use TUKEY rather the Fischer. Fisher is bogus apparently. 

## Lmer (no transformation)
```{r}
random <- lmer(interrow_weed_biomass_lbs_ac  ~ location+weed_control + location:weed_control +(1|location:block) , data = interrow_weed_biomass_clean)

resid_panel(random)
### Two checks specifically for a generalize linear approach
simulateResiduals(random,plot = TRUE) # Residuals and normality look good
check_model(random) #Perfect, preditions match real data

```

###Joint test (anova)
```{r}
random |> 
  joint_tests() |> 
  kable()  
```
### Means comparison
#### Weed-control (Significant)
```{r message=FALSE}
means_weed_control <- emmeans(random, ~  weed_control)
pairwise_comparisons_weed_control<- pairs(means_weed_control) 
kable(pairwise_comparisons_weed_control)
```
br>

#### Location (Not significant)
```{r message=FALSE}
means_location <- emmeans(random, ~  location)
pairwise_comparisons_location<- pairs(means_location) 
kable(pairwise_comparisons_location)
```
#### Location|weed-control (Not significant)
```{r message=FALSE}
means_weed_control_location <- emmeans(random, ~  weed_control|location)
pairwise_comparisons_weed_control_location<- pairs(means_weed_control_location) 
kable(pairwise_comparisons_weed_control_location)
```
### Tukey compact letter display
#### Weed-control (Significant)

```{r}
#weed control
cld_weed_control_tukey <-cld(emmeans(random, ~  weed_control , type = "response"), Letters = letters, sort = TRUE, reversed=TRUE)
cld_weed_control_tukey

```
### Fisher compact letter display
#### Weed-control (Significant)

```{r}
 cld_weed_control_fisher <-cld(emmeans(random,~  weed_control,type = "response"), Letters = letters, sort = TRUE, adjust="none", reversed=TRUE)
 cld_weed_control_fisher
```

## Lmer (log transformation)
```{r}
random_log <- lmer(log_interrow_weed_biomass_grams_meter  ~ location+weed_control + location:weed_control +(1|location:block) , data = interrow_weed_biomass_clean)

resid_panel(random_log)
```
###Joint test (anova)
```{r}
random_log |> 
  joint_tests() |> 
  kable()  
```

### Means comparison
#### Weed-control (Significant)
```{r message=FALSE}
means_weed_control_log <- emmeans(random_log, ~  weed_control)
pairwise_comparisons_weed_control_log<- pairs(means_weed_control) 
kable(pairwise_comparisons_weed_control_log)
```
#### Location (Significant)
```{r message=FALSE}
means_location_log <- emmeans(random_log, ~  location)
pairwise_comparisons_location_log<- pairs(means_location_log) 
kable(pairwise_comparisons_location_log)
```

#### Location|weed-control (Not significant)
```{r message=FALSE}
means_weed_control_location_log <- emmeans(random_log, ~  weed_control|location)
pairwise_comparisons_weed_control_location_log<- pairs(means_weed_control_location_log) 
kable(pairwise_comparisons_weed_control_location_log)
```

### Tukey compact letter display
#### Weed-control (Significant)

```{r}
#weed control
cld_weed_control_tukey_log <-cld(emmeans(random_log, ~  weed_control , type = "response"), Letters = letters, sort = TRUE, reversed=TRUE)
cld_weed_control_tukey_log

```
#### Location (Significant)
```{r}
#location
cld_location_tukey_log <-cld(emmeans(random_log, ~  location , type = "response"), Letters = letters, sort = TRUE, reversed=TRUE)
cld_location_tukey_log

```
for weed seed look into quasi or negative binomial, hurdle model, or 0 inflated

### Fisher compact letter display
#### Weed-control (Significant)

```{r}
 cld_weed_control_fisher_log <-cld(emmeans(random_log,~  weed_control,type = "response"), Letters = letters, sort = TRUE, adjust="none", reversed=TRUE)
 cld_weed_control_fisher_log
```

#### Location (Significant)

```{r}
 cld_weed_control_fisher_log <-cld(emmeans(random_log,~  location,type = "response"), Letters = letters, sort = TRUE, adjust="none", reversed=TRUE)
cld_weed_control_fisher_log
```
#GLMM
Redo this after changing block to 12 blocks (note that )
```{r}

model_tweedie_log <- glmmTMB(
interrow_weed_biomass_lbs_ac ~  weed_control + (1|location) + (1|block)+
  (1|location:block), 
  data = interrow_weed_biomass_clean, 
  family = tweedie(link = "log")

)
### Two checks specifically for a generalize linear approach
simulateResiduals(model_tweedie_log,plot = TRUE) # Residuals and normality look good
check_model(model_tweedie_log) #Perfect, preditions match real data

summary(model_tweedie_log )
VarCorr(model_tweedie_log )


```
### Joint test (anova)
```{r}
model_tweedie_log |> 
  joint_tests() |> 
  kable()  
```

## Fisher compact letter display
### Weed-control (S)

```{r}
cld_weed_control_fisher <-cld(emmeans(model_tweedie_log, ~  weed_control, type = "response"), Letters = letters,adjust = "none", sort = TRUE, reversed=TRUE)
cld_weed_control_fisher


```

# Figures
## Weed-control




```{r message=FALSE}
interrow_weed_biomass_clean |> 
  left_join(cld_weed_control_fisher) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")), y = response, fill = weed_control)) +
  #stat_summary(geom = "bar", fun = mean, width = 0.7) +  # Ensuring mean calculation
  #stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2) +  # Explicit mean_se
  #stat_summary(geom = "text", fun = mean, aes(label = trimws(.group)), 
               #size = 6.5, vjust = -0.5) +  # Adjusting labels) 
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), width=.2,
                 position=position_dodge(.9))+
geom_text(aes(label = trimws(.group), y = response + (SE + 20)), size = 7) +
  labs(
    x = "",
    y = expression("Interrow weed biomass" ~ (lbs * "/" * a)),
    #title = str_c("Influence of interrow weed control on interrow weed biomass"),
    subtitle = expression(italic("P < 0.005"))) +
  
 scale_x_discrete(labels = c("Rolled,\nno additional\nweed control",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nhigh-residue\ncultivation",
                              "Tilled,\ninterrow\nmowing",
                          "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12),
    axis.title = element_text(size = 20),  # Increase font size of axis titles
    axis.text = element_text(size = 20),   # Increase font size of axis labels
    plot.title = element_text(size = 22, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 18, face = "italic")  # Increase font size of subtitle
  )
ggsave("interrow_weed_biomass_weed_control_lbac.png", width = 10, height = 6, dpi = 300)
```

```{r}
interrow_weed_biomass_clean |> 
  left_join(cld_weed_control_fisher) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")), 
             y = response, 
             fill = weed_control)) +
  geom_bar(stat = "identity", width = 0.7) +
  geom_errorbar(aes(ymin = response - SE, ymax = response + SE), width = 0.2) +
  geom_text(aes(y = response + SE, label = trimws(.group)), size = 6.5, vjust = -0.5) +
  labs(
    x = "",
    y = expression("Interrow weed biomass" ~ (lbs * "/" * a)),
    subtitle = expression(italic("P < 0.005"))
  ) +
  scale_x_discrete(labels = c("Rolled,\nno additional\nweed control",
                               "Rolled,\ninterrow\nmowing",
                               "Rolled,\nhigh-residue\ncultivation",
                               "Tilled,\ninterrow\nmowing",
                               "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12),
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 16),
    plot.title = element_text(size = 22, face = "bold"),
    plot.subtitle = element_text(size = 18, face = "italic")
  )
ggsave("interrow_weed_biomass_weed_control_lbac.png", width = 10, height = 6, dpi = 300)

```
