---
title: "Weed biomass"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# **Load libraries**

```{r message=FALSE, warning=FALSE}
#Set work directory
setwd("/Users/ey239/Github/Mowtivation/rmarkdowns")

#Load packages 
library(tidyverse) ##install.packages("tidyverse")
library(knitr)
library(patchwork) ##install.packages("patchwork")
library(skimr)     ##install.packages("skimr")
library(readxl)
library(janitor) ##install.packages("janitor")

library(kableExtra) ##install.packages("kableExtra")
library(webshot) ##install.packages("webshot")
webshot::install_phantomjs()
library(viridis) ##install.packages("viridis")
library(lme4) ##install.packages("lme4")
library(lmerTest) ##install.packages("lmerTest")
library(emmeans) ##install.packages("emmeans")
library(rstatix) ##install.packages("rstatix")
#library(Matrix) ##install.packages("Matrix")
library(multcomp) ##install.packages("multcomp")
library(multcompView) ##install.packages("multcompView")
library(ggResidpanel) ##install.packages("ggResidpanel")
#library(car)
#library(TMB)  ##install.packages("TMB")
#library(glmmTMB)  ##install.packages("glmmTMB")
#library(DHARMa)  ##install.packages("DHARMa")

#Load Functions
MeanPlusSe<-function(x) mean(x)+plotrix::std.error(x)

find_logw0=function(x){c=trunc(log(min(x[x>0],na.rm=T)))
d=exp(c)
return(d)}
```
<br>

# **Load and Clean Data**

### **Load individual datasets**

``` {r}
combined_raw <- read_excel("~/Github/Mowtivation/raw-data/All Treatments/combined_raw.xlsx")
kable(head(combined_raw))

```


```{r}


#Standardaze column names, convert to factors, check for outliers of variable**
clean_combined <- clean_names(combined_raw) |>  
  rename ('weed_control'= treatment) |> 
  mutate(across(c(weed_control, block, plot, location, year), as.factor)) #|> 
  #mutate(is_outlier = totwbm < (quantile(totwbm, 0.25) - 1.5 * IQR(totwbm)) |
                       #wbm > (quantile(totwbm, 0.75) + 1.5 * IQR(totwbm)))

#select and convert data for wbm analysis
weed_biomass_clean <-clean_combined |>              
  mutate(weed_biomass_grams_meter = (weed_biomass * 2)) |> 
  mutate(weed_biomass_kg_ha = ((weed_biomass/0.5) *(10000))/(1000)) |>
  mutate(weed_biomass_lbs_ac = (((weed_biomass/0.5) *(10000))/(1000))* 0.892179)
kable(head(weed_biomass_clean)) 
```
<br>
### **block is random**
#Ask tyler about model format, should block always be fixed, etc. should location be nested in year? Do i need to transform data if zeros are present? How to separate year from location? 

```{r}
random <- lmer(weed_biomass_kg_ha  ~ location+weed_control + location:weed_control +(1|location:block) , data = weed_biomass_clean)

resid_panel(random)
```

```{r}
summary(random)
weed_biomass_clean |> count(year,location
                            ,weed_control)

```

### **block is random**
```{r}
random <- lmer(weed_biomass_kg_ha  ~ weed_control + (1|block)  , data = weed_biomass_clean)

resid_panel(random)
```

##**Joint test**
```{r}
fixed |> 
  joint_tests() |> 
  kable()  
```
<br>

# **Means comparison of totwbm**

```{r message=FALSE}
means <- 
 emmeans(fixed, ~  weed_control)
# Optional: Adjust for multiple comparisons (e.g., using Tukey's method)

pairwise_comparisons<- pairs(means) 
kable(head(pairwise_comparisons))
```
### **Fisher's method for comparing means**

```{r}
#mowing
cld_weed_control_fisher <-cld(emmeans(fixed, ~  weed_control , type = "response"), Letters = letters, sort = TRUE, reversed=TRUE)
cld_weed_control_fisher
```



```{r}
#mowing
cld_weed_control_location_fisher <-cld(emmeans(fixed, ~  weed_control|location , type = "response"), Letters = letters, sort = TRUE, adjust="none", reversed=TRUE)
cld_weed_control_location_fisher
```

# **FIGURES**

## **Cultivation**
#For these figures, why doesn't it let my use weed_biomass_kg_ha for my Y value?

```{r message=FALSE}
clean_combined |> 
  left_join(cld_weed_control_fisher) |> 
  ggplot(aes(x = weed_control, y = emmean, fill = weed_control)) +
  stat_summary(geom = "bar", fun = "mean", width = 0.7) +
  stat_summary(geom = "errorbar", fun.data = "mean_se", width = 0.2) +
  stat_summary(geom="text", fun = "MeanPlusSe", aes(label= trimws(.group)),size=6.5,vjust=-0.5) +
  labs(
    x = "Interrow weed control method",
    y = expression("Weed biomass" ~ (kg ~ ha^{-1})),
    title = str_c("The influence of interrow weed control method on weed biomass"),
    subtitle = expression(italic("P < 0.005"))) +
  
  scale_x_discrete(labels = c("Rolled,\nhigh-residue\ncultivation",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nno additional\nweed control",
                          "Tilled,\nstandard\ncultivation",
                              "Tilled,\ninterrow\nmowing")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12)
  )
ggsave("weed_biomass_weed_control.png", width = 8, height = 6, dpi = 300)
```

## **weed_control:location**

```{r message=FALSE}
clean_combined |> 
  left_join(cld_weed_control_location_fisher) |> 
  ggplot(aes(x = weed_control, y = emmean, fill = weed_control)) +
  facet_wrap(~location )+
  stat_summary(geom = "bar", fun = "mean", width = 0.7) +
  stat_summary(geom = "errorbar", fun.data = "mean_se", width = 0.2) +
  stat_summary(geom="text", fun = "MeanPlusSe", aes(label= trimws(.group)),size=6.5,vjust=-0.5) +
  labs(
    x = "Method of interrow weed control",
    y = expression("Weed biomass" ~ (kg ~ ha^{-1})),
    title = str_c("Influence of the method of interrow weed control on weed biomass"),
    subtitle = expression(italic("P < 0.005"))) +
  
  scale_x_discrete(labels = c("Rolled,\nhigh-residue\ncultivation",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nno additional\nweed control",
                          "Tilled,\nstandard\ncultivation",
                              "Tilled,\ninterrow\nmowing")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 12)
  )
ggsave("weed_biomass_weed_control_location.png", width = 12, height = 6, dpi = 300)
```