---
title: "Weed biomass"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Load libraries

```{r message=FALSE, warning=FALSE}
#Set work directory
setwd("/Users/ey239/Github/Mowtivation/rmarkdowns")

#Load packages 
library(tidyverse) ##install.packages("tidyverse")
library(knitr)
library(patchwork) ##install.packages("patchwork")
library(skimr)     ##install.packages("skimr")
library(readxl)
library(janitor) ##install.packages("janitor")

library(kableExtra) ##install.packages("kableExtra")
library(webshot) ##install.packages("webshot")
webshot::install_phantomjs()
library(viridis) ##install.packages("viridis")
library(lme4) ##install.packages("lme4")
library(lmerTest) ##install.packages("lmerTest")
library(emmeans) ##install.packages("emmeans")
library(rstatix) ##install.packages("rstatix")
#library(Matrix) ##install.packages("Matrix")
library(multcomp) ##install.packages("multcomp")
library(multcompView) ##install.packages("multcompView")
library(ggResidpanel) ##install.packages("ggResidpanel")
#library(car)
#library(TMB)  ##install.packages("TMB")
library(glmmTMB)  ##install.packages("glmmTMB")
library(DHARMa)  ##install.packages("DHARMa")
library(performance) ##install.packages("performance")

#Load Functions
MeanPlusSe<-function(x) mean(x)+plotrix::std.error(x)

find_logw0=function(x){c=trunc(log(min(x[x>0],na.rm=T)))
d=exp(c)
return(d)}
```
<br>

# Load and clean data

## Load data

``` {r}
combined_raw <- read_excel("~/Github/Mowtivation/raw-data/All Treatments/combined_raw.xlsx")
kable(head(combined_raw))

```
##Clean data

```{r}


#Standardaze column names, convert to factors, check for outliers of variable**
clean_combined <- clean_names(combined_raw) |>  
  rename ('weed_control'= treatment) |> 
  mutate(across(c(weed_control, block, plot, location, year), as.factor)) #|> 
  #mutate(is_outlier = totwbm < (quantile(totwbm, 0.25) - 1.5 * IQR(totwbm)) |
                       #wbm > (quantile(totwbm, 0.75) + 1.5 * IQR(totwbm)))

#select and convert data for wbm analysis
weed_biomass_clean <-clean_combined |> 
  mutate(weed_biomass_grams_meter = (weed_biomass /0.5)) |> 
  mutate(weed_biomass_kg_ha = (weed_biomass_grams_meter *(10000))/(1000)) |>
  mutate(weed_biomass_lbs_ac = ((weed_biomass_grams_meter *(10000))/(1000))* 0.892179)
kable(head(weed_biomass_clean)) 
```
<br>
# Model testing

Block is random
Tyler is under the impression that block should always be random and that post-hoc comparisons should use TUKEY rather the Fischer. Fisher is bogus apparently. 



#GLMM

```{r}

model_tweedie_log <- glmmTMB(weed_biomass_lbs_ac ~  weed_control*location + (1|location:block), 
  data = weed_biomass_clean, 
  family = tweedie(link = "log")

)

### Two checks specifically for a generalize linear approach
simulateResiduals(model_tweedie_log,plot = TRUE) # Residuals and normality look good
check_model(model_tweedie_log) #Perfect, preditions match real data

summary(model_tweedie_log )
VarCorr(model_tweedie_log )


```




#### Joint test (anova)
```{r}
model_tweedie_log |> 
  joint_tests() |> 
  kable()  
```
```{r}
options(contrasts = c("contr.sum", "contr.poly"))
Anova(model_tweedie_log, type = 3)
```
<br>

## Fisher compact letter display
### Weed-control (S)

```{r}
cld_weed_control_fisher <-cld(emmeans(model_tweedie_log, ~  weed_control, type = "response"), Letters = letters,adjust = "none", sort = TRUE, reversed=TRUE)
cld_weed_control_fisher


```
## Fisher compact letter display
weed_control|location (Significant)

```{r}
cld_weed_control_location_fisher <-cld(emmeans(model_tweedie_log, ~  weed_control|location, type = "response"), Letters = letters,adjust = "none", sort = TRUE, reversed=TRUE)
cld_weed_control_location_fisher
```


# Figures
###lbs/a
#### Weed_control|Location (Significant)

```{r message=FALSE}
weed_biomass_clean |> 
  left_join(cld_weed_control_location_fisher) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")), y = response, fill = weed_control)) +
facet_wrap( ~location, labeller = labeller(
    location = c("field O2 east" = "Field A", "field O2 west" = "Field B","field x" = "Field C" )))+
  #stat_summary(geom = "bar", fun = "mean", width = 0.7) +
  #stat_summary(geom = "errorbar", fun.data = "mean_se", width = 0.2) +
  #stat_summary(geom="text", fun = "MeanPlusSe", aes(label= trimws(.group)),size=6.5,vjust=-0.5) +
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), width=.2,
                 position=position_dodge(.9))+
geom_text(aes(label = trimws(.group), y = response + (SE + 48)), size = 7) +
  labs(
    x = "",
     y = expression("Weed biomass" ~ (lbs * "/" * a)),
    #title = str_c("Influence of interrow weed control on weed biomass"),
    subtitle = expression(italic("P < 0.005"))) +
  
  scale_x_discrete(labels = c("Rolled,\nno additional\nweed control",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nhigh-residue\ncultivation",
                              "Tilled,\ninterrow\nmowing",
                          "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 20),
    axis.title = element_text(size = 20),  # Increase font size of axis titles
    axis.text = element_text(size = 20),   # Increase font size of axis labels
    plot.title = element_text(size = 24, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 24, face = "italic")  # Increase font size of subtitle
  
  )
ggsave("weed_biomass_weed_control_location_lb_acre.png", width = 24, height = 8, dpi = 300)
```

# Figures
###Kg/h
#### Weed_control|Location (Significant)

```{r message=FALSE}
weed_biomass_clean |> 
  left_join(cld_weed_control_location_fisher) |> 
  ggplot(aes(x = factor(weed_control, levels = c("RNO", "RIM", "RIC", "TIM", "TIC")), y = response, fill = weed_control)) +
facet_wrap( ~location, labeller = labeller(
    location = c("field O2 east" = "Field A", "field O2 west" = "Field B","field x" = "Field C" )))+
  #stat_summary(geom = "bar", fun = "mean", width = 0.7) +
  #stat_summary(geom = "errorbar", fun.data = "mean_se", width = 0.2) +
  #stat_summary(geom="text", fun = "MeanPlusSe", aes(label= trimws(.group)),size=6.5,vjust=-0.5) +
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), width=.2,
                 position=position_dodge(.9))+
geom_text(aes(label = trimws(.group), y = response + (SE + 48)), size = 7) +
  labs(
    x = "",
     y = expression("Weed biomass" ~ (kg~ha^{-1})),
    #title = str_c("Influence of interrow weed control on weed biomass"),
    subtitle = expression(italic("P < 0.005"))) +
  
  scale_x_discrete(labels = c("Rolled,\nno additional\nweed control",
                              "Rolled,\ninterrow\nmowing",
                              "Rolled,\nhigh-residue\ncultivation",
                              "Tilled,\ninterrow\nmowing",
                          "Tilled,\nstandard\ncultivation")) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
  scale_fill_viridis(discrete = TRUE, option = "D", direction = -1, end = 0.9, begin = 0.1) +
   theme_bw() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 20),
    axis.title = element_text(size = 20),  # Increase font size of axis titles
    axis.text = element_text(size = 20),   # Increase font size of axis labels
    plot.title = element_text(size = 24, face = "bold"),  # Increase font size of title
    plot.subtitle = element_text(size = 24, face = "italic")  # Increase font size of subtitle
  
  )
ggsave("weed_biomass_weed_control_location_kg_ha.png", width = 24, height = 8, dpi = 300)
```

